---
title: "IST 707 Homework 3"
output:
  word_document: default
---

Due Date: 4/21/2021     

# Introduction

A personal equity plan (PEP) product offering was introduced by a financial
firm and it was marketed by sending out a direct mail piece to customers. 
Data was collected about customers and whether or not they purchased the PEP,
along with demographic information about the customer.

There might be patterns that indicate which customers are more likely than others
to purchase the PEP depending on their characteristics. The following report goes
through the data and applies data mining to try and find these patterns. The goal
is to find interesting association rules, and based on the discoveries, provide a
business recommendation. 

# Data Preparation

Ensure packages are installed and active.

```{r}

if   ("readr"        %in% installed.packages())   {require(readr)}        else {install.packages('readr');        require(readr)}
if   ("dplyr"        %in% installed.packages())   {require(dplyr)}        else {install.packages('dplyr');        require(dplyr)}
if   ("reshape2"     %in% installed.packages())   {require(reshape2)}     else {install.packages('reshape2');     require(reshape2)}
if   ("sqldf"        %in% installed.packages())   {require(sqldf)}        else {install.packages('sqldf');        require(sqldf)}
if   ("stringr"      %in% installed.packages())   {require(stringr)}      else {install.packages('stringr');      require(stringr)}
if   ("reticulate"   %in% installed.packages())   {require(reticulate)}   else {install.packages('reticulate');   require(reticulate)}
if   ("arules"       %in% installed.packages())   {require(arules)}       else {install.packages('arules');       require(arules)}
if   ("ggplot2"      %in% installed.packages())   {require(ggplot2)}      else {install.packages('ggplot2');      require(ggplot2)}

```

Read the data and store it.

```{r}

# make sure that the data-storyteller.csv file is in the current working directory

bank <- as.data.frame(read_csv("bankdata_csv_all.csv", col_types = cols()))

```

Show a sample of the dataset.

```{r}

head(bank)

```

Remove the ID field.

```{r}

# double check the ID field is unique and if yes then delete. 

bank <- if (length(unique(bank$id)) == dim(bank) [1]) {   #logical test - is every id unique? 
        bank <- bank[,-which(colnames(bank) == 'id')] }   #condition if true - remove the column

bankcopy <- bank

```

Discretize the age variable.

```{r}

# using equal width approach with 5 x 10 bins

`age_18-27` <- c()
`age_28-37` <- c()
`age_38-47` <- c()
`age_48-57` <- c()
`age_58-67` <- c()

# populate the categories accordingly

for (x in bank$age) {
  
if (between(x, 18, 27)) {`temp_18-27` <- 1} else {`temp_18-27` <- 0}
if (between(x, 28, 37)) {`temp_28-37` <- 1} else {`temp_28-37` <- 0}
if (between(x, 38, 47)) {`temp_38-47` <- 1} else {`temp_38-47` <- 0}
if (between(x, 48, 57)) {`temp_48-57` <- 1} else {`temp_48-57` <- 0}
if (between(x, 58, 67)) {`temp_58-67` <- 1} else {`temp_58-67` <- 0}
  
`age_18-27` <- c(`age_18-27`, `temp_18-27`)
`age_28-37` <- c(`age_28-37`, `temp_28-37`)
`age_38-47` <- c(`age_38-47`, `temp_38-47`)
`age_48-57` <- c(`age_48-57`, `temp_48-57`)
`age_58-67` <- c(`age_58-67`, `temp_58-67`)
}

# append the categories to dataframe

bank$`age_18-27` <- `age_18-27`
bank$`age_28-37` <- `age_28-37`
bank$`age_38-47` <- `age_38-47`
bank$`age_48-57` <- `age_48-57`
bank$`age_58-67` <- `age_58-67`

# remove the original age variable

bank <- bank[,-which(colnames(bank) == 'age')]

# clean up the environment

rm(`age_18-27`, `age_28-37`, `age_38-47`, `age_48-57`, `age_58-67`,
   `temp_18-27`, `temp_28-37`, `temp_38-47`, `temp_48-57`, `temp_58-67`, x)

```

Discretize the income variable.

```{r}

# determine splits using equal frequency approach

incomebrackets <- discretize(bank$income, method = "frequency", breaks = 3, onlycuts = TRUE)
incomebrackets

incomesplit1 <- incomebrackets[2]
incomesplit2 <- incomebrackets[3]

# three income brackets - low, medium, high

paste('low income <= $', round(incomesplit1,2))
paste('$', round(incomesplit1,2), '< middle income <= $', round(incomesplit2,2))
paste('high income > $', round(incomesplit2,2))

bank$lowIncome <- ifelse(bank$income <= incomesplit1, 1, 0)
bank$middleIncome <- ifelse(bank$income > incomesplit1 & bank$income <= incomesplit2, 1, 0)
bank$highIncome <- ifelse(bank$income > incomesplit2, 1, 0)

# check that the discretization is correct

cat('count of low income = ', sum(bank$lowIncome),'\n',
    'count of middle income = ', sum(bank$middleIncome), '\n',
    'count of high income = ', sum(bank$highIncome))

# remove the original income variable

bank <- bank[,-which(colnames(bank) == 'income')]

# clean up the environment

rm(incomebrackets, incomesplit1, incomesplit2)

```

Discretize the rest of the variables.

```{r}

# create a function for discretizing the rest

discretizebankdata <- function(dataframe, columnstodiscretize) {
  
  tempdf <- dataframe
  
  for (column in columnstodiscretize) {
    
    .GlobalEnv$tempcolname <- as.character(column)
    
    .GlobalEnv$columnvalues <- tempdf[,which(colnames(tempdf) == column)]
    
    valuenames <- unique(columnvalues)
    
      for (valuename in valuenames) {
        
      discretizedvalue <- ifelse(.GlobalEnv$columnvalues == valuename, 1, 0)

      tempdf <- cbind(tempdf, discretizedvalue)
      
      colnames(tempdf) [dim(tempdf) [2]] <- paste0(.GlobalEnv$tempcolname, as.character(valuename))
      
      }
   }
  
  return(tempdf)
  
}

bank <- discretizebankdata(bank, colnames(bank) [1:9])

# remove the original variables

bank <- bank[,-c(1:9)]

# clean up the environment

rm(columnvalues, tempcolname, discretizebankdata)

```

# Initial Data Exploration

```{r}

# summarize the data.

summary(bankcopy)

```

```{r}

# correlation of age and income.

plot(x = bankcopy$age, y = bankcopy$income,
  main = paste('R = ', round(cor(bankcopy$age, bankcopy$income),2)))
abline(lm(income ~ age, data = bankcopy))

# summarize data for the age variable.

summary(bankcopy$age)
hist(bankcopy$age, main = "Histogram of age")
boxplot(bankcopy$age, main = "Boxplot of age")

# summarize data for the income variable.

summary(bankcopy$income)
hist(bankcopy$income, main = "Histogram of income")
boxplot(bankcopy$income, main = "Boxplot of income")

```

```{r}

# average income grouped by variables.

incomematrix <- data.frame()

for (column in colnames(bankcopy)[which( ! colnames(bankcopy) %in% c('age', 'income'))]) {
tempquery          <- paste0('select ', column, ' as grouped, income from bankcopy')
tempdata           <- as.data.frame(sqldf(tempquery))
tempdata$measure   <- column
incomematrix       <- rbind(incomematrix, tempdata)}

incomematrix <- as.data.frame(sqldf(

'select measure, grouped, avgIncome, recordCount, round(recordCount / 600, 2) as pctTotal from (
select measure, grouped, cast(avg(income) as int) as avgIncome, cast(count(measure || grouped) as float) as recordCount from
incomematrix group by measure, grouped order by measure, grouped asc) sub'))

incomematrix

#clean up the environment

rm(tempdata, column, tempquery)

```

Observations:
-there are a near equal amount of customers in each age group
-the income variable is normally distributed but right skewed
-there is a strong positive correlation with age and income
-males and females are an even 50/50 split in the data
-46% of customers purchased the PEP and 54% did not
-avg income is $6,000 higher for those who purchased the PEP

# Data Analysis

Read transactions and generate rules.

```{r}

# read the transactions

banktransactions <- as(as.matrix(bank), "transactions")

# item frequency plot

itemFrequencyPlot(banktransactions, topN=20,type="absolute")

# generate rules

#bankrules <- apriori(banktransactions, parameter = list(supp = 0.5, conf = 0.8)) #attempt 1 - did not produce any rules
#bankrules <- apriori(banktransactions, parameter = list(supp = 0.01, conf = 0.8)) #attempt 2 - produced too many rules
bankrules <- apriori(banktransactions, parameter = list(supp = 0.15, conf = 0.8)) #attempt 3 - produced 31 rules, all strong

```

```{r}

# top 5 rules sorted by highest support

inspect((sort(bankrules, decreasing = TRUE, by = "support"))[1:5])

# top 5 rules sorted by highest confidence

inspect((sort(bankrules, decreasing = TRUE, by = "confidence"))[1:5])

# top 5 rules sorted by highest lift

inspect((sort(bankrules, decreasing = TRUE, by = "lift"))[1:5])

```

Analyze rules with PEP on the RHS.

```{r}

#create new bank rules with PEP on rhs all possible

newbankrules <- apriori(banktransactions, parameter = list(maxlen = 8, supp = .0001, conf = .0001), appearance = list(rhs = c("pepNO", "pepYES"), default = "lhs"))
newbankrulesdf <- as(newbankrules, "data.frame")

```

```{r}

# top 10 rules sorted by highest support

inspect((sort(newbankrules, decreasing = TRUE, by = "support"))[1:10])

# top 10 rules sorted by highest confidence

inspect((sort(newbankrules, decreasing = TRUE, by = "confidence"))[1:10])

# top 10 rules sorted by highest lift

inspect((sort(newbankrules, decreasing = TRUE, by = "lift"))[1:10])

```

My top five interesting rules.

1. {age_58-67} => {pepYES}

  support = 0.13, confidence = 0.61, lift = 1.34
  
  Customers who are in the age bracket of 58 to 67 are 61% likely to purchase PEP.
  13% of all customers fall into this age bracket. This means that every 
  100 customers will produce about 8 PEP sales. Out of all the other age groups,
  the next highest was the 38 to 47 age bracket with 5 PEP sales per 100 customers.
  
  It makes sense that the older demographic is more likely to purchase the PEP,
  because they are also probably more likely to respond to direct mail in general
  than younger generations are. Also, as shown earlier, there is a positive
  correlation with age, income, and a PEP being purchased, which aligns with this rule.
  Below is a graph illustrating the estimated PEP sales for all of the age groups.
  
  Cumulative expected PEP sales per 100 customers = 8
  Rule 1 = 8
  
```{r}

# rule number 1 supporting evidence

temprules <- head(newbankrulesdf[which(grepl('age_', newbankrulesdf$rules)),], 10)
temprules <- temprules[which(grepl('pepYES', temprules$rules)),]
temprules$PEPsalesPer100 <- round((temprules$support * 100) * temprules$confidence, 0)
temprules <- temprules[,which(colnames(temprules) %in% c('rules', 'PEPsalesPer100'))]
temprules$rules <- substr(temprules$rules, 2, (unique(unlist(str_locate_all(temprules$rules, '\\}'))) [1]) -1)

ggplot(temprules, aes(rules, PEPsalesPer100)) + geom_col(aes(width = 0.5)) +
  ggtitle("Expected PEP Sales Per 100 Customers by Age Group")

```
  
2. {highIncome,children1}   => {pepYES}    support = 0.08, confidence = 0.96, lift = 2.10  
   {highIncome,children2}   => {pepYES}    support = 0.07, confidence = 0.89, lift = 1.96  
   {middleIncome,children1} => {pepYES}    support = 0.06, confidence = 0.95, lift = 2.08  
   
  My second rule includes the combination of of three rules. If the customer has either one or
  two children and and is in the high income bracket, or the customer has one child and is in
  the middle income bracket, there are highly likely to purhase PEP. About 21% of all transactions
  constitute this rule, and so out of every 100 customers, this should produce a total of 20 PEP sales.
  
  Some of the 58 to 67 age bracket will cross over with this subsegment. Since these
  customers are already accounted for via the previous rule, I will be
  excluding them from the cumulative totals. After the deduction, the expected PEP sales for
  this rule is (20 - 7.72) = 12 expected PEP sales.
  
  highIncome, children1, and age 58-67     -----> ((23 / 600) x 100) x .96 = 3.68  
  highIncome, children2, and age 58-67     -----> ((23 / 600) x 100) x .89 = 3.41  
  middleIncome, children1, and age 58-67   -----> (( 4 / 600) x 100) x .95 = 0.63  
  
  The graph below shows the expected PEP sales for each of the
  income brackets and for how many children the customers have in each of those income
  brackets. Although the low income bracket with 1 child does produce an estimated 2 PEP
  sales per 100 customers, this not nearly as much as the other demographics.
  
  I am recommending not to market to the low income group altogether. The data
  for low income in aggregate shows that for every 100 customers, there will be 4 PEP sales. 
  This is a much smaller percentage. I am proposing that this group is not worth marketing 
  to and the resources should be prioritized for middle and high income bracket. The support
  for this group and confidence is not high enough.
  
  Cumulative expected PEP sales per 100 customers = 20  
  Rule 1 = 8, Rule 2 = 12
  
```{r}

# rule number 2 supporting evidence

temprules                  <- newbankrulesdf[which(grepl('Income', newbankrulesdf$rules)),]
temprules                  <- temprules[which(grepl('children', temprules$rules)),]
temprules                  <- temprules[which(grepl('pepYES', temprules$rules)),]
temprules                  <- head(temprules, 12)
temprules$PEPsalesPer100   <- round((temprules$support * 100) * temprules$confidence, 0)

# parse through the rules column text

cutpoints <- c(); for (rule in temprules$rules) {
  rulecutpoint <- (unique(unlist(str_locate_all(rule, '\\}'))) [1]) -1
  cutpoints <- c(cutpoints, rulecutpoint)}
temprules$cutpoints <- cutpoints
temprules$rules <- substr(temprules$rules, 2, temprules$cutpoints)

cutpoints <- c(); for (rule in temprules$rules) {
  rulecutpoint <- (unique(unlist(str_locate_all(rule, ','))) [1]) -1
  cutpoints <- c(cutpoints, rulecutpoint)}
temprules$cutpoints <- cutpoints
temprules$IncomeBracket <- substr(temprules$rules, 1, temprules$cutpoints)

cutpoints <- c(); for (rule in temprules$rules) {
  rulecutpoint <- nchar(rule)
  cutpoints <- c(cutpoints, rulecutpoint)}
temprules$cutpoints <- cutpoints
temprules$NumChildren <- substr(temprules$rules, temprules$cutpoints, temprules$cutpoints)

temprules <- temprules[,which(colnames(temprules) %in% c('IncomeBracket', 'NumChildren', 'PEPsalesPer100'))]

# plot each of the income brackets

highincomebracket <- sqldf('select * from temprules where IncomeBracket = \'highIncome\'')
ggplot(highincomebracket, aes(x = NumChildren, y = PEPsalesPer100)) + geom_col(width = 0.5,) + ylim(0, 10) +
  ggtitle("High Income Bracket - Number of Children and Expected PEP Sales per 100 Customers")

middleincomebracket <- sqldf('select * from temprules where IncomeBracket = \'middleIncome\'')
ggplot(middleincomebracket, aes(x = NumChildren, y = PEPsalesPer100)) + geom_col(width = 0.5) + ylim(0, 10) +
  ggtitle("Middle Income Bracket - Number of Children and Expected PEP Sales per 100 Customers")

lowincomebracket <- sqldf('select * from temprules where IncomeBracket = \'lowIncome\'')
ggplot(lowincomebracket, aes(x = NumChildren, y = PEPsalesPer100)) + geom_col(width = 0.5) + ylim(0, 10) +
  ggtitle("Low Income Bracket - Number of Children and Expected PEP Sales per 100 Customers")

# clean up the environment

rm(cutpoints, rule, rulecutpoint, temprules)

```

3. {marriedNO,save_actYES,mortgageNO}      => {pepYES}     support = 0.11, confidence = 0.74, lift = 1.63  
   {marriedNO,current_actYES,mortgageNO}   => {pepYES}     support = 0.12, confidence = 0.72, lift = 1.57  

   I am recommending the combination of these two groups, given that
   they are somewhat similar. There is some overlap between these two rules.
   There is also some overlap with the previous two rules. For example,
   some of these customers may also fall in the age 58-67 bracket.
   To prevent double counting, the deductions are applied, and then an adj.
   support and avg. confidence is used to calculate the estimated PEP sales
   per 100 customers.
   
   After the deductions are applied, the total adj.
   support comes out to be (71 / 600) = 0.12. The avg. confidence is
   (0.74 + 0.72) / 2 = 0.73. Therefore, the expected PEP sales per 100
   customers is = (0.12 x 100) x 0.72 = 8.64 = 8 (rounded down).
   
   As seen in the evidence presented below, when mortgage is not taken into account,
   the percentage of confidence for not being married along with a savings or current
   account is around 56-58%. When not having a mortgage is introduced as a factor,
   that jumps up to 70%, leading to an extra 1-2 PEP sales from the target segment.
   
   Cumulative expected PEP sales per 100 customers = 28  
   Rule 1 = 8, Rule 2 = 12, Rule 3 = 8

```{r}

# rule number 3 supporting evidence

temprules                  <- newbankrulesdf[which(grepl('married', newbankrulesdf$rules)),]
temprules                  <- temprules[which(grepl('pepYES', temprules$rules)),]
temprules                  <- temprules[which(grepl('marriedNO', temprules$rules)),]
temprules                  <- temprules[which(grepl('current_act', temprules$rules) | grepl('save_act', temprules$rule)),]
temprules                  <- sqldf('select * from temprules order by count desc')
temprules                  <- head(temprules, 4)
temprules$PEPsalesPer100   <- round((temprules$support * 100) * temprules$confidence, 0)

temprules

# calculate deductions

rule3deductions            <- bankcopy[which(
  
                                (bankcopy$married == 'NO' &
                                 bankcopy$save_act == 'YES' &
                                 bankcopy$mortgage == 'NO') |
                                  
                                 # OR
                                  
                                (bankcopy$married == 'NO' &
                                 bankcopy$current_act == 'YES' &
                                 bankcopy$mortgage == 'NO')),]

rule3deductions             <- rule3deductions[which( ! rule3deductions$age >= 58),] # 30 deductions
rule3deductions             <- rule3deductions[which( ! (rule3deductions$income > 31132.77 & rule3deductions$children == 1)),]   # 7 deductions
rule3deductions             <- rule3deductions[which( ! (rule3deductions$income > 31132.77 & rule3deductions$children == 2)),]   # 7 deductions
rule3deductions             <- rule3deductions[which( ! ((rule3deductions$income > 20253.80 & rule3deductions$income <= 31132.77) & rule3deductions$children == 1)),]  # 3 deductions

```

4. {children0,save_actNO,mortgageYES} => {pepYES}

   support = 0.06, confidence = 0.92, lift = 2.01

   To find this rule, I first filtered the rules down by removing subsegments that have
   been looked at so far. This rule encompasses a section of the market that has
   not been addressed yet by any of the previous rules.Therefore, there are no deductions
   needed for this rule. The expected PEP sales per 100 customers comes out to 5 (rounded down).
   
   Discuss the support, confidence and lift values and how they are interpreted in this data set.
   
   The support value of 0.06 means that 6% of all of the transactions in the data set contain
   children = 0, savings account = NO, and mortgage = YES. The confidence of 0.92 means that
   in these transactions, 92% of them resulted in a pep = YES. The lift of 2.01 indicates that
   pep = YES is highly likely to occur along with occurences of the LHS. 
   
   Cumulative expected PEP sales per 100 customers = 33  
   Rule 1 = 8, Rule 2 = 12, Rule 3 = 8, Rule 4 = 5

```{r}

# rule number 4 supporting evidence

# remove all of the previous rules

temprules <- newbankrulesdf[which( ! grepl('age_58-67', newbankrulesdf$rules)),]
temprules <- temprules[which( ! (grepl('highIncome', temprules$rules) & grepl('children1', temprules$rules))),]
temprules <- temprules[which( ! (grepl('highIncome', temprules$rules) & grepl('children2', temprules$rules))),]
temprules <- temprules[which( ! (grepl('middleIncome', temprules$rules) & grepl('children1', temprules$rules))),]
temprules <- temprules[which( ! (grepl('marriedNO', temprules$rules) & grepl('save_actYES', temprules$rules) & grepl('mortgageNO', temprules$rules))),]
temprules <- temprules[which( ! (grepl('marriedNO', temprules$rules) & grepl('current_actYES', temprules$rules) & grepl('mortgageNO', temprules$rules))),]

# find the rule

temprules                  <- temprules[which(grepl('pepYES', temprules$rules)),]
temprules$PEPsalesPer100   <- round((temprules$support * 100) * temprules$confidence, 0)
temprules                  <- temprules[which(temprules$PEPsalesPer100 >= 2),]
temprules                  <- temprules[which( ! grepl('children1', temprules$rules)),]

# are there any deductions needed for age 58-67?
# (this is the only potential area of cross over)

length(bankcopy[which(
  
  bankcopy$age >= 58 &
  bankcopy$children == 0 &
  bankcopy$save_act == 'NO' &
  bankcopy$mortgage == 'Yes'), 1])

```

5. {regionINNER_CITY,carYES,save_actNO}    => {pepYES}     support = 0.05, confidence = 0.60, lift = 1.31  
   {regionINNER_CITY,carYES,current_actNO} => {pepYES}     support = 0.03, confidence = 0.57, lift = 1.24  
   
   The last market segment includes customers who live in the inner city region, own a car,
   and do not have either a savings account or a current account. The intent of this rule
   is to try and find a niche in the market that has not been addressed yet by the previous
   rules. This was difficult because the previous rules have already accounted for most
   of the customers.
   
   Deductions are necessary for this category because there is some overlap
   with previous category. After the deductions are applied, the total adj.
   support comes out to be (25 / 600) = 0.04. Given that there is some
   overlap between these two rules, the estimated PEP sales per 100 customers
   is calculated using an avg. confidence.
   
```{r}

# adj. support = 0.04, avg. confidence = 0.585

cat('Expected PEP sales per 100 customers = ', ((0.04 * 100) * 0.585))

```

   Cumulative expected PEP sales per 100 customers = 35
   Rule 1 = 8, Rule 2 = 12, Rule 3 = 8, Rule 4 = 5, Rule 5 = 2
   
```{r}

# rule number 5 supporting evidence

# find the rule

temprules                  <- temprules[which(grepl('regionINNER_CITY', temprules$rules)),]
temprules                  <- temprules[which(grepl('carYES', temprules$rules)),]

# calculate deductions

rule5deductions            <- bankcopy[which(
  
                                (bankcopy$region == 'INNER_CITY' &
                                 bankcopy$car == 'YES' &
                                 bankcopy$current_act == 'NO') |
                                  
                                 # OR
                                  
                                (bankcopy$region == 'INNER_CITY' &
                                 bankcopy$car == 'YES' &
                                 bankcopy$save_act == 'NO')),]

rule5deductions             <- rule5deductions[which( ! rule5deductions$age >= 58),] # 15 deductions
rule5deductions             <- rule5deductions[which( ! (rule5deductions$income > 31132.77 & rule5deductions$children == 1)),]   # 0 deductions
rule5deductions             <- rule5deductions[which( ! (rule5deductions$income > 31132.77 & rule5deductions$children == 2)),]   # 3 deductions
rule5deductions             <- rule5deductions[which( ! ((rule5deductions$income > 20253.80 & rule5deductions$income <= 31132.77) & rule5deductions$children == 1)),]  # 4 deductions
rule5deductions             <- rule5deductions[which( ! (rule5deductions$married == 'NO' & rule5deductions$save_act == 'YES' & rule5deductions$mortgage == 'NO')),]   # 4 deductions
rule5deductions             <- rule5deductions[which( ! (rule5deductions$married == 'NO' & rule5deductions$current_act == 'YES' & rule5deductions$mortgage == 'NO')),]   # 4 deductions
rule5deductions             <- rule5deductions[which( ! (rule5deductions$children == 0 & rule5deductions$save_act == 'NO' & rule5deductions$mortgage == 'YES')),]   # 8 deductions

```

Efficiency = total number of customers marketed to / expected PEP sales per 100 customers

The efficiency is the ratio of the total number of customers marketed to over the
the expected PEP sales per 100 customers. If this number is low, it indicates
good utilization of resources. If this number is high, it indicates poor
utilization of resources. While the target customer segment segment does not achieve
the full 45 PEP sales that it could if it marketed to the entire set of customers,
it gets close and with much less resources. The rules that were chosen
maximize the amount of PEP sales while minimizing the amount of resources used.

Efficiency of marketing to all customers (current state) = 13.3

total number of customers marketed to = 600
expected PEP sales per 100 customers = 45

Efficiency of marketing to target customers (future state) = 9.6

total number of customers marketed to = 335
expected PEP sales per 100 customers = 35

```{r}

# total number of customers marketed to in the target segment

newcustomersegment <- bankcopy[which(
  
  bankcopy$age >= 58 |
  (bankcopy$income > 31132.77 & bankcopy$children == 1) |
  (bankcopy$income > 31132.77 & bankcopy$children == 2) |
  ((bankcopy$income > 20253.80 & bankcopy$income <= 31132.77) & bankcopy$children == 1) |
  (bankcopy$married == 'NO' & bankcopy$save_act == 'YES' & bankcopy$mortgage == 'NO') |
  (bankcopy$married == 'NO' & bankcopy$current_act == 'YES' & bankcopy$mortgage == 'NO') |
  (bankcopy$children == 0 & bankcopy$save_act == 'NO' & bankcopy$mortgage == 'YES') |
  (bankcopy$region == 'INNER_CITY' & bankcopy$car == 'YES' & bankcopy$save_act == 'NO') |
  (bankcopy$region == 'INNER_CITY' & bankcopy$car == 'YES' & bankcopy$current_act == 'NO')),]

```

# Conclusion

From the original data provided, there is a little under a 50/50 shot for 
any given customer to purchase the PEP. So why not just continue to market
the PEP to every customer? Almost half of them will end up purchasing the
PEP anyway. For one, there are expenses that come with marketing. That would be a lot
of money wasted on customers who don't end up purchasing the PEP - money that could
have been used for something else.

In this report, I attempted to identify pockets of customers that are most
likely to purchase the PEP and it turns out that there were several of them. By
narrowing down the amount of customers, the cost of marketing is able to be
minimized while the number if PEP sales per customer is able to maximized.
Therefore, I recommend continuing to use this method for these customers,
as the direct mail advertising is shwown to be effective.

For the customer segments that were not likely to purchase the PEP, I would
recommend collecting more data on. There could be a business opportunity
to introduce a new product. For example, customers on the lower end
of income and who are young are not likely to purchase the PEP.
Was this the outcome that was expected and if not what is the reason that
these customers do not purchase the PEP? Is there something else that they
would purchase? Another factor to consider would be if a
different type of marketing would be more effective. For example, a lot of
marketing is done electronically nowadays, so this might lead to more
purchases of the PEP from younger generations. 

To recap, my final recommendation is to continue using the direct mail advertising
for the PEP for the customer segments that this works best for. This will cut
down the cost of marketing significantly without losing too many sales.
The cost savings from narrowing down the marketing for PEP could be put towards
further researching the market, developing a new product, or trying out an electronic
marketing campaign.