---
output:
    html_document:
        toc: true
        toc_depth: 5
        toc_float:
            collapsed: FALSE
            smooth_scroll: true
        fig_height: 5
        fig_width: 8
        df_print: paged
        code_folding: hide
---

<p>&nbsp;</p>

#### **Introduction**

<p>&nbsp;</p>

<font size = "4">

Up until recently, cryptocurrency has been seen mostly as a speculative asset. 
That all changed when the crypto market popped in early 2020. Suddenly, there
was a huge wave of hype and prices skyrocketed across the board. Since then,
things have cooled off, but there has been a lasting impact. Crypto is starting
to become a widely accepted form of currency.

It is easier than ever to trade crypto nowadays with apps such as Coinbase or
Binance and there are thousands of crypto coins that are available to be traded
on these platforms. Maybe the chance to get rich quick has already passsed, but
I believe that there are still a lot of opportunities to make money. My goal in
this project is to create a prediction of cryptocurrency prices for over 150
different cryptocurrencies.

</font>

<p>&nbsp;</p>

#### **I. Environment Setup**

<p>&nbsp;</p>

##### **i. Load R Packages**

<p>&nbsp;</p>

<font size = "4">

To start off, the necessary R packages will be loaded in order to run the code
in this report. I have added a short comment next to each one below to indicate
its usage.

</font>

```{r Load R packages, message = FALSE, warning = FALSE, class.source = 'fold-show'}

# Load R packages
library(readr)       # Reading in csv files
library(dplyr)       # Pipe operations and joins
library(ggplot2)     # Data visualization
library(plotly)      # Interactive data visualization
library(sqldf)       # Data manipulation with SQL
library(stringr)     # String operations and regex
library(reshape2)    # Casting and melting
library(lubridate)   # Date functions
library(zoo)         # na.locf() function for filling NAs
library(corrplot)    # Correlation matrix
library(kableExtra)  # Rendering nicely formatted tables
library(prophet)     # Time series forecasting

```

<p>&nbsp;</p>

##### **ii. Load the Data**

<p>&nbsp;</p>

<font size = "4">

Next, the data will be loaded. There are 7 csv files, each which correspond with
one table. These files are produced from running the data collection jupyter
notebook which I have done separately. A dataframe that I call the crypto
dataframe is formed by joining the 7 tables together.

</font>

```{r Load the data, message = FALSE, warning = FALSE, class.source = 'fold-show'}

# Load the data
T_Asset <- read_csv("Asset.csv", col_types = cols(.default = "c"))
T_Price <- read_csv("Price.csv", col_types = cols(.default = "c"))
T_CoinMarket <- read_csv("CoinMarket.csv", col_types = cols(.default = "c"))
T_Youtube <- read_csv("Youtube.csv", col_types = cols(.default = "c"))
T_Twitter <- read_csv("Twitter.csv", col_types = cols(.default = "c"))
T_Google <- read_csv("Google.csv", col_types = cols(.default = "c"))
T_Stocks <- read_csv("Stocks.csv", col_types = cols(.default = "c"))

# Convert data types
T_Price[, 3:7] <- T_Price[, 3:7] %>% mutate_all(as.numeric)
T_CoinMarket[, 3:5] <- T_CoinMarket[, 3:5] %>% mutate_all(as.numeric)
T_Youtube[, 3] <- T_Youtube[, 3] %>% mutate_all(as.numeric)
T_Twitter[, 3] <- T_Twitter[, 3] %>% mutate_all(as.numeric)
T_Google[, 3] <- T_Google[, 3] %>% mutate_all(as.numeric)
T_Stocks[, 2:3] <- T_Stocks[, 2:3] %>% mutate_all(as.numeric)

# Crypto dataframe
Crypto <- T_Price %>%
  left_join(T_CoinMarket, by = c("Ticker", "Date")) %>%
  left_join(T_Youtube, by = c("Ticker", "Date")) %>%
  left_join(T_Twitter, by = c("Ticker", "Date")) %>%
  left_join(T_Google, by = c("Ticker", "Date")) %>%
  left_join(T_Stocks, by = c("Date")) %>%
  left_join(T_Asset, by = c("Ticker"))

```

<p>&nbsp;</p>

#### **II. Initial Data Preparation**

<p>&nbsp;</p>

##### **i. Add Date Part Columns**

<p>&nbsp;</p>

<font size = "4">

The following code adds several date part columns to the crypto dataframe. The
columns added are year, month, week, day of week, year month, year week,
last day of month, and last day of week. These columns are useful in time series
analysis and will be used later on.

</font>

```{r Add date part columns, message = FALSE, warning = FALSE, class.source = 'fold-show'}

# Add date part columns
Crypto <- Crypto %>%
  mutate(Year = year(Date)) %>%
  mutate(Month = ifelse(month(Date) < 10, paste0("0", month(Date)), month(Date))) %>%
  mutate(Week = ifelse(week(Date) < 10, paste0("0", week(Date)), week(Date))) %>%
  mutate(DayOfWeek = as.numeric(format(as.Date(Date), format = "%u"))) %>%
  mutate(YearMonth = paste0(Year, Month)) %>%
  mutate(YearWeek = paste0(Year, Week)) %>%
  transform(LastDayMonth = as.numeric(Month != c(Month[-1], NA))) %>%
  mutate(LastDayWeek = ifelse(DayOfWeek == 7, 1, 0))

```

<p>&nbsp;</p>

##### **ii. Handle Some Missing Values**

<p>&nbsp;</p>

<font size = "4">

For the DJIPrice and NASDAQPrice variables, Saturdays, Sundays, and holidays are
non trading days. This is not an issue with crypto since crypto is traded 24/7.
Thus, as a result of joining the data, NAs appear under the DJIPrice and
NASDAQPrice variables on these non trading days. I do not want to drop the NAs
since that would remove important crypto data. Thus, the NAs will be filled with
the last known closing price.

</font>

```{r Fill in NAs in the DJI and NASDAQ variables, message = FALSE, warning = FALSE, class.source = 'fold-show'}

# Fill in NAs in DJI and NASDAQ variables
Crypto$DJIPrice <- na.locf(Crypto$DJIPrice)
Crypto$NASDAQPrice <- na.locf(Crypto$NASDAQPrice)

```

<p>&nbsp;</p>

<font size = "4">

There are some more missing values that need to be handled. I started collecting
the CirculatingSupply, MaximumSupply, CMCRank, and TweetCount variables at the
start of this project. I have been capturing these in daily increments by
running the data collection code once per day, but there have been some days
where I have not been able to run it. I will fill in these NAs in the same way
as I did with the DJIPrice and NASDAQPrice variables.

</font>

```{r Fill in NAs in other variables, message = FALSE, warning = FALSE, class.source = 'fold-show'}

# Fill in NAs in other variables
for (Column in colnames(Crypto)[c(8:10, 12)]){
  for (Cryptocurrency in unique(Crypto$Ticker)) {
    FilteredData <- Crypto[Crypto$Ticker == Cryptocurrency, ]
    BeforeData <- FilteredData[FilteredData$Date < "2022-02-14", Column]
    AfterData <- FilteredData[FilteredData$Date >= "2022-02-14", Column]
    if (sum(na.locf(AfterData)) == 0) {next}
    NewColumn <-  c(BeforeData, na.locf(AfterData, na.rm = FALSE))
    Crypto[Crypto$Ticker == Cryptocurrency, Column] <- NewColumn
  }
}

# Delete the leftover variables
rm(Column, Cryptocurrency, FilteredData, BeforeData, AfterData, NewColumn)

```

<p>&nbsp;</p>

##### **iii. Show the Crypto Dataframe**

<p>&nbsp;</p>

<font size = "4">

After the data transformations that have taken place up to this point, here is
what the crypto dataframe looks like.

</font>

```{r Show the crypto dataframe, message = FALSE, warning = FALSE}

# Show the crypto dataframe
Crypto %>%
  filter(complete.cases(Crypto)) %>%
  arrange(desc(Date), Ticker) %>%
  head(50) %>%
  kbl(digits = 2) %>%
  kable_styling(full_width = TRUE) %>%
  scroll_box(height = "500px", fixed_thead = list(enabled = T))

```

<p>&nbsp;</p>

#### **III. Data Exploration**

<p>&nbsp;</p>

<font size = "4">

Next, I will start to explore the data. First, some basic questions will be
answered. This will help to gain some familiarity with the data. After that, I
will show a histogram of current cryptocurrency prices, a correlation matrix
of the numeric variables in the Crypto dataframe, and an interactive time series
plot.

</font>

<p>&nbsp;</p>

##### **i. Basic Questions**

<p>&nbsp;</p>

<font size = "4">

*How many unique cryptocurrencies are there in the dataset?*

</font>

```{r Unique cryptocurrency coins, message = FALSE, warning = FALSE}

length(unique(Crypto$Ticker))

```

<p>&nbsp;</p>

<font size = "4">

*What are the top 25 most expensive cryptocurrencies?*

</font>

```{r Most expensive cryptocurrency, message = FALSE, warning = FALSE}

sqldf("
select
  Ticker,
  Slug,
  Close
from (
    select
      Ticker,
      Slug,
      Date,
      max(Date) as MaxDate,
      Close
    from
      Crypto
    group by
      Ticker
) sub
where
  Date = MaxDate
order by
 Close desc
") %>%
  
mutate(Close = paste0("$", round(Close, 0))) %>%
rename(`Current Price` = Close) %>%
head(25)

```

<p>&nbsp;</p>

<font size = "4">

*What are the top 25 least expensive cryptocurrencies?*

</font>

```{r Least expensive cryptocurrency, message = FALSE, warning = FALSE}

sqldf("
select
  Ticker,
  Slug,
  Close
from (
    select
      Ticker,
      Slug,
      Date,
      max(Date) as MaxDate,
      Close
    from
      Crypto
    group by
      Ticker
) sub
where
  Date = MaxDate
order by
 Close asc
") %>%
  
mutate(Close = format(Close, scientific = F)) %>%
mutate(Close = paste0("$", trimws(Close))) %>%
rename(`Current Price` = Close) %>%
head(25)

```

<p>&nbsp;</p>

<font size = "4">

*What are the top 25 most traded cryptocurrencies?*

</font>

```{r Most traded cryptocurrency, message = FALSE, warning = FALSE}

aggregate(data = Crypto, Volume ~ Ticker + Slug, mean) %>%
  mutate(Volume = round(Volume, 0)) %>%
  arrange(desc(Volume)) %>%
  mutate(Volume = format(Volume, scientific = F)) %>%
  rename(`Average 24 Hour Trading Volume` = Volume) %>%
  head(25)

```

<p>&nbsp;</p>

<font size = "4">

*What are the top 25 least traded cryptocurrencies?*

</font>

```{r Least traded cryptocurrency, message = FALSE, warning = FALSE}

aggregate(data = Crypto, Volume ~ Ticker + Slug, mean) %>%
  mutate(Volume = round(Volume, 0)) %>%
  arrange(Volume) %>%
  mutate(Volume = format(Volume, scientific = F)) %>%
  rename(`Average 24 Hour Trading Volume` = Volume) %>%
  head(25)

```

<p>&nbsp;</p>

<font size = "4">

*Which 25 cryptocurrencies are closest to their maximum supply?*

</font>

```{r Closest to max supply, message = FALSE, warning = FALSE}

Crypto %>%
  filter(! is.na(CirculatingSupply) & ! is.na(MaximumSupply)) %>%
  select(Ticker, Date) %>%
  aggregate(formula = Date ~ Ticker, FUN = max, data = .) %>%
  left_join(Crypto, by = c("Ticker" = "Ticker", "Date" = "Date")) %>%
  select(Ticker, Slug, CirculatingSupply, MaximumSupply) %>%
  mutate(PercentMaximumSupply = CirculatingSupply / MaximumSupply) %>%
  mutate(CirculatingSupply = format(CirculatingSupply, scientific = F)) %>%
  mutate(MaximumSupply = format(MaximumSupply, scientific = F)) %>%
  arrange(desc(PercentMaximumSupply)) %>%
  mutate(PercentMaximumSupply = paste0(round(PercentMaximumSupply * 100, 2), "%")) %>%
  rename(`Circulating Supply` = CirculatingSupply) %>%
  rename(`Maximum Supply` = MaximumSupply) %>%
  rename(`Percent of Maximum Supply In Circulation` = PercentMaximumSupply) %>%
  head(25)

```

<p>&nbsp;</p>

<font size = "4">

*Which 25 cryptocurrencies are furthest away from their maximum supply?*

</font>

```{r Furthest away from max supply, message = FALSE, warning = FALSE}

Crypto %>%
  filter(! is.na(CirculatingSupply) & ! is.na(MaximumSupply)) %>%
  select(Ticker, Date) %>%
  aggregate(formula = Date ~ Ticker, FUN = max, data = .) %>%
  left_join(Crypto, by = c("Ticker" = "Ticker", "Date" = "Date")) %>%
  select(Ticker, Slug, CirculatingSupply, MaximumSupply) %>%
  mutate(PercentMaximumSupply = CirculatingSupply / MaximumSupply) %>%
  mutate(CirculatingSupply = format(CirculatingSupply, scientific = F)) %>%
  mutate(MaximumSupply = format(MaximumSupply, scientific = F)) %>%
  arrange(PercentMaximumSupply) %>%
  mutate(PercentMaximumSupply = paste0(round(PercentMaximumSupply * 100, 2), "%")) %>%
  rename(`Circulating Supply` = CirculatingSupply) %>%
  rename(`Maximum Supply` = MaximumSupply) %>%
  rename(`Percent of Maximum Supply In Circulation` = PercentMaximumSupply) %>%
  head(25)

```

<p>&nbsp;</p>

<font size = "4">

*Which 25 cryptocurrencies have the largest average daily increase in circulating supply?*

</font>

```{r Largest average daily increase, message = FALSE, warning = FALSE}

Crypto %>%
  filter(! is.na(CirculatingSupply) & ! is.na(MaximumSupply)) %>%
  transform(SupplyIncrease = c(CirculatingSupply[-1], NA) - CirculatingSupply) %>%
  transform(SupplyIncreasePercent = round(SupplyIncrease / c(CirculatingSupply[-1], NA), 2)) %>%
  transform(NewTicker = c(Ticker[-1], NA) != Ticker) %>%
  mutate(SupplyIncrease = ifelse(NewTicker, NA, SupplyIncrease)) %>%
  mutate(SupplyIncreasePercent = ifelse(NewTicker, NA, SupplyIncreasePercent)) %>%
  aggregate(cbind(SupplyIncrease, SupplyIncreasePercent) ~ Ticker + Slug, mean, data = .) %>%
  arrange(desc(SupplyIncreasePercent)) %>%
  mutate(SupplyIncrease = format(round(SupplyIncrease, 0), scientific = F)) %>%
  mutate(SupplyIncreasePercent = paste0(round(SupplyIncreasePercent * 100, 2), "%")) %>%
  rename(`Average Daily Increase In Circulating Supply (Units)` = SupplyIncrease) %>%
  rename(`Average Daily Increase In Circulating Supply (Percent)` = SupplyIncreasePercent) %>%
  head(25)
  

```

<p>&nbsp;</p>

##### **ii. Current Price Histogram**

<p>&nbsp;</p>

<font size = "4">

The following chart shows the distribution of the log 10 of the current prices
for all the cryptocurrencies. The log 10 is used because otherwise the
distribution is way too sparse to fit into a graph. As seen in the top 10 most
expensive and top 10 least expensive cryptocurrencies, the prices can range from
fractions of a cent to tens of thousands of dollars.

</font>

```{r Current price histogram, message = FALSE, warning = FALSE}

# Current prices dataframe
CurrentPrices <- Crypto %>%
  filter(! is.na(Close)) %>%
  filter(Date == max(Date))

# Min and max prices
MinPrice <- round(min(log10(CurrentPrices$Close)), 2)
MaxPrice <- round(max(log10(CurrentPrices$Close)), 2)
PriceRange <- MaxPrice - MinPrice

# X axis adjustment
XAxisBreaks = seq(MinPrice, MaxPrice, PriceRange / 10) %>% round(2)
XAxisTicks = seq(MinPrice, MaxPrice, PriceRange / 10) %>% round(2)

# Current prices histogram
CurrentPrices %>%
  select(Close) %>%
  ggplot(aes(x = log10(Close))) +
  geom_histogram(color = "black", fill = "darkgrey", bins = 50) +
  scale_x_continuous(breaks = XAxisBreaks, labels = XAxisTicks) +
  ggtitle("\nCrypto Price Histogram\n\n") +
  xlab("\nLog 10 of Price\n") +
  ylab("\nCount\n") +
  labs(caption = paste0("Closing prices as of ", max(Crypto$Date), " per Yahoo Finance")) +
  theme(axis.text.x = element_text(size = 9.5, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 9.5),
        axis.title.x = element_text(size = 10, color = "black", face = "bold"),
        axis.title.y = element_text(size = 10, color = "black", face = "bold"),
        title = element_text(size = 11, color = "black", face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        panel.background = element_rect(fill = NA))

# Remove the leftover variables
rm(MaxPrice, MinPrice, XAxisBreaks, XAxisTicks, PriceRange, CurrentPrices)

```

<p>&nbsp;</p>

##### **iii. Correlation Matrix**

<font size = "4">

A couple important things to note about the following correlation matrices. The
correlation displayed is the mean correlation among all cryptocurrencies. Also,
correlations including the CirculatingSupply, CMCRank, VideoCount, and
TweetCount variables are computed using data from Feb 14th or later due to
limited data. The other correlations are computed using data from a max of 5
years ago or the beginning of available data, whichever comes first.

</font>

<p>&nbsp;</p>

```{r Correlation Matrix, fig.height = 8, fig.width = 16, message = FALSE, warning = FALSE}

# Set the display to 1 row 2 columns
par(mfrow = c(1, 2))

# Short term correlation matrix
ShortTerm <- data.frame()
for (Cryptocurrency in unique(Crypto$Ticker)) {
  ShortTerm <- Crypto %>%
    filter(Ticker == Cryptocurrency) %>%
    filter(Date >= "2022-02-14") %>%
    filter(Date <= as.character(Sys.Date())) %>%
    select(Close, CirculatingSupply:TweetCount) %>%
    select(-MaximumSupply) %>%
    cor(method = "pearson") %>%
    as.data.frame() %>%
    mutate(Ticker = Cryptocurrency) %>%
    rbind(ShortTerm)
}

# Further manipulate the dataframe
ShortTerm <- ShortTerm %>%
  mutate(Measure = row.names(.)) %>%
  mutate(Measure = gsub("([0-9]*)", "", Measure)) %>%
  melt(id = c("Ticker", "Measure")) %>%
  aggregate(value ~ Measure + variable, mean, data = .) %>%
  dcast(Measure ~ variable, value.var = "value")  %>%
  mutate(across(where(is.numeric), ~ round(., 2))) %>%
  arrange(Measure) %>%
  `rownames<-`(.[, "Measure"]) %>%
  select(-Measure) %>%
  .[, sort(colnames(.), decreasing = T)] %>%
  .[rev(row.names(.)), ] %>%
  as.matrix() %>%
  
# Feed to correlation plot function
corrplot(
  method = "square",
  type = "lower",
  diag = F,
  addCoef.col = "black",
  col = COL2("BrBG"),
  tl.col = "black",
  title = "\nAverage Correlations 2022-02-14 to Present\n",
  mar = c(1, 1, 1, 1),
  order = "original",
  tl.srt = 45
)

# Long term correlation matrix
LongTerm <- data.frame()
for (Cryptocurrency in unique(Crypto$Ticker)) {
  LongTerm <- Crypto %>%
    filter(Ticker == Cryptocurrency) %>%
    filter(Date <= as.character(Sys.Date())) %>%
    filter(! is.na(InterestLevel)) %>%
    select(Close, Volume, InterestLevel:NASDAQPrice) %>%
    cor(method = "pearson") %>%
    as.data.frame() %>%
    mutate(Ticker = Cryptocurrency) %>%
    rbind(LongTerm)
}

# Further manipulate the dataframe
LongTerm <- LongTerm %>%
  mutate(Measure = row.names(.)) %>%
  mutate(Measure = gsub("([0-9]*)", "", Measure)) %>%
  melt(id = c("Ticker", "Measure")) %>%
  aggregate(value ~ Measure + variable, mean, data = .) %>%
  dcast(Measure ~ variable, value.var = "value")  %>%
  mutate(across(where(is.numeric), ~ round(., 2))) %>%
  arrange(Measure) %>%
  `rownames<-`(.[, "Measure"]) %>%
  select(-Measure) %>%
  .[, sort(colnames(.), decreasing = T)] %>%
  .[rev(row.names(.)), ] %>%
  as.matrix() %>%
  
# Feed to correlation plot function
corrplot(
  method = "square",
  type = "lower",
  diag = F,
  addCoef.col = "black",
  col = COL2("BrBG"),
  tl.col = "black",
  title = "\nAverage Correlations 5 Years Ago to Present\n",
  mar = c(1, 1, 1, 1),
  order = "original",
  tl.srt = 45
)

# Return the display settings to normal
par(mfrow = c(1, 1))

# Remove the leftover variables
rm(LongTerm, ShortTerm, Cryptocurrency)
    
```

<p>&nbsp;</p>

##### **iv. Time Series Plot**

<font size = "4">

The following time series plot displays the price overtime for any of the
cryptocurrencies. One cryptocurrency can be selected at a time by using the
filter.

</font>

<p>&nbsp;</p>

```{r Time series plot, message = FALSE, warning = FALSE}

# Create nested list for the buttons argument
Buttons <- list(mode = "list", length = length(unique(Crypto$Ticker)))
Counter <- 1
for (Name in unique(paste0(Crypto$Slug, " (", Crypto$Ticker, ")"))) {
  Button <- list(
    method = "restyle",
    args = list("transforms[0].value", Name),
    label = Name
    )
  Buttons[[Counter]] <- Button
  Counter <- Counter + 1
}
names(Buttons) <- NULL

# Create the plotly object
Figure <- Crypto %>%
  
  filter(LastDayWeek == 1) %>%
  mutate(CryptoName = paste0(Slug, " (", Ticker, ")")) %>%
  mutate(Date = as.Date(Date, "%Y-%m-%d")) %>%
  
  plot_ly(
    type = "scatter", 
    x = ~Date, 
    y = ~Close,
    text = ~Close,
    hoverinfo = "text",
    mode = "line", 
    transforms = list(
      list(
        type = "filter",
        target = ~CryptoName,
        operation = "=",
        value = unique(paste0(Crypto$Slug, " (", Crypto$Ticker, ")"))[1]
      )
    )
  ) %>%
  
  layout(
    updatemenus = list(
      list(
        type = "dropdown",
        active = 0,
        bordercolor = "black",
        buttons = Buttons,
        label = "Cryptocurrency Selector"
      )
    )
  ) %>%
  
  layout(
    title = "<br>Cryptocurrency Time Series Plot<br>",
    showlegend = FALSE,
    xaxis = list(title = "<br>Week End Date<br>", tickangle = -45),
    yaxis = list(title = "<br>Closing Price<br>"),
    plot_bgcolor = "#e5ecf6"
  )

Figure

# Remove the leftover variables
rm(Button, Buttons, Counter, Name, Figure)

```

<p>&nbsp;</p>

#### **III. Data Modeling**

<p>&nbsp;</p>

<font size = "4">

In the data modeling phase, I will train a forecast for each cryptocurrency
using different techniques and evaluate their performances based on a 3 month
hold out period.

</font>

<p>&nbsp;</p>

##### **i. Split Test / Train Data**

<p>&nbsp;</p>

<font size = "4">

For each cryptocurrency, the test data will consist of the most recent 3 months
and the train data will consist of everything prior to that.

</font>

```{r Split test / train data, message = FALSE, warning = FALSE}

# Set the cutoff to todays date minus 90 days
Cutoff <- as.character(as.Date(Sys.Date(), "%Y-%m-%d") - 90)

# Split test / train data based on the cutoff
TrainAll <- Crypto %>% filter(Date <= Cutoff)
Test <- Crypto %>% filter(Date > Cutoff)

```

<p>&nbsp;</p>

##### **ii. Different History Versions**

<p>&nbsp;</p>

<font size = "4">

Different versions of the training data will be created. Each version has a
pruned amount of historical data There are six versions going from using
all of the history to using just one month of history. Also, each version
will contain outlier adjusted prices, adjusted to 1, 2, and 3 standard
deviations from the mean.

</font>

```{r Different History Versions, message = FALSE, warning = FALSE}

AddSdColumn <- function(Df) {
  Sd <- aggregate(data = Df, Close ~ Ticker, sd)
  colnames(Sd) <- c("Ticker", "Sd")
  left_join(Df, Sd, by = c("Ticker" = "Ticker"))
}

AddMeanColumn <- function(Df) {
  Mean <- aggregate(data = Df, Close ~ Ticker, mean)
  colnames(Mean) <- c("Ticker", "Mean")
  left_join(Df, Mean, by = c("Ticker" = "Ticker"))
}

AddUpperLimit <- function(Df, Sigma) {
  ColumnName <- paste0("UpperLimit", Sigma)
  ColumnVals <- Df$Mean + (Sigma * Df$Sd)
  Df <- Df %>% mutate(Column = ColumnVals)
  colnames(Df)[ncol(Df)] <- ColumnName
  return(Df)
}

AddLowerLimit <- function(Df, Sigma) {
  ColumnName <- paste0("LowerLimit", Sigma)
  ColumnVals <- Df$Mean - (Sigma * Df$Sd)
  Df <- Df %>% mutate(Column = ColumnVals)
  colnames(Df)[ncol(Df)] <- ColumnName
  return(Df)
}

AddAllLimits <- function(Df) {
  for (Sigma in c(1, 2, 3)) {
    Df <- AddUpperLimit(Df, Sigma)
    Df <- AddLowerLimit(Df, Sigma)
  }
  Df
}

AdjustToLimit <- function(Y, Upper, Lower) {
  YAdjusted <- ifelse(Y > Upper, Upper, Y)
  ifelse(YAdjusted < Lower, Lower, YAdjusted)
}

AddAdjusted <- function(Df, Sigma, UpperIx, LowerIx) {
  ColumnName <- paste0("Close", Sigma)
  ColumnVals <- AdjustToLimit(Df[, "Close"], Df[, UpperIx], Df[, LowerIx])
  Df <- Df %>% mutate(Column = ColumnVals)
  colnames(Df)[ncol(Df)] <- ColumnName
  return(Df)
}

AddAllAdjusted <- function(Df) {
  Args <- list(c(1, 2, 3), c(27, 29, 31), c(28, 30, 32))
  for (Ix in c(1, 2, 3)) {
    Df <- AddAdjusted(Df, Args[[1]][Ix], Args[[2]][Ix], Args[[3]][Ix])
  }
  Df
}

OutlierPreprocess <- function(Df) {
  Df <- Df %>% AddSdColumn()
  Df <- Df %>% AddMeanColumn()
  Df <- Df %>% AddAllLimits()
  AddAllAdjusted(Df)
}

TrainAll <- TrainAll %>%
  OutlierPreprocess(.)

Train24 <- Crypto %>%
  filter(Date >= as.character(as.Date(Sys.Date(), "%Y-%m-%d") - 820)) %>%
  filter(Date <= Cutoff) %>%
  OutlierPreprocess(.)

Train12 <- Crypto %>%
  filter(Date >= as.character(as.Date(Sys.Date(), "%Y-%m-%d") - 455)) %>%
  filter(Date <= Cutoff) %>%
  OutlierPreprocess(.)

Train6 <- Crypto %>%
  filter(Date >= as.character(as.Date(Sys.Date(), "%Y-%m-%d") - 275)) %>%
  filter(Date <= Cutoff) %>%
  OutlierPreprocess(.)

Train3 <- Crypto %>%
  filter(Date >= as.character(as.Date(Sys.Date(), "%Y-%m-%d") - 185)) %>%
  filter(Date <= Cutoff) %>%
  OutlierPreprocess(.)

Train1 <- Crypto %>%
  filter(Date >= as.character(as.Date(Sys.Date(), "%Y-%m-%d") - 120)) %>%
  filter(Date <= Cutoff) %>%
  OutlierPreprocess(.)

```

<p>&nbsp;</p>

##### **iii. Prophet Model**

<p>&nbsp;</p>

<font size = "4">

The prophet package is a package developed by Facebook for time series
forecasting. In the following code, I will be producing and evaluating a
forecast using prophet on each version of the crypto data.

</font>

```{r Prophet model, message = FALSE, warning = FALSE}

# Step 1: create historical dataframe
CreateHistoricalDf <- function(CryptoDf, CryptoTicker, YColName) {
  CryptoDf[CryptoDf$Ticker == CryptoTicker, c("Date", YColName)]
}

# Step 2: train a prophet model
TrainProphetModel <- function(HistoricalDf) {
  colnames(HistoricalDf) <- c("ds", "y")
  HistoricalDf %>% prophet(yearly.seasonality = FALSE,
                           weekly.seasonality = FALSE,
                           daily.seasonality = TRUE)
}

# Step 3: create a future dataframe
CreateFutureDf <- function(ProphetModel, NumPeriods) {
  make_future_dataframe(ProphetModel, NumPeriods, include_history = FALSE)
}

# Step 4: make predictions on the future dataframe
MakePredictions <- function(ProphetModel, FutureDf) {
  predict(ProphetModel, FutureDf)[, c("ds", "yhat")] %>%
    mutate(ds = as.character(ds))
}

# Step 5: create a y test dataframe
CreateYTestDf <- function(CryptoDf, CryptoTicker, CutoffDate) {
  CryptoDf <- CryptoDf %>% filter(Date > CutoffDate, Ticker == CryptoTicker)
  CryptoDf %>% select(Date, Close)
}

# Step 6: create a results df
CreateOutputDf <- function(YPredDf, YTestDf) {
  left_join(YPredDf, YTestDf, by = c("ds" = "Date"))
}

# Step 7: measure n day accuracy scores
MeasureNDayAccuracy <- function(OutputDf, N) {
  AbsError <- sum(abs(OutputDf$yhat[1:N] - OutputDf$Close[1:N]), na.rm = TRUE)
  1 - (AbsError / sum(OutputDf$Close[1:N], na.rm = TRUE))
}

# Step 8: create accuracy dataframe
CreateAccuracyDf <- function(OutputDf, CryptoTicker, Version) {
  data.frame(
    "Ticker" = CryptoTicker,
    "Version" = Version,
    "Accuracy7Day" = MeasureNDayAccuracy(OutputDf, 7),
    "Accuracy30Day" = MeasureNDayAccuracy(OutputDf, 30),
    "Accuracy90Day" = MeasureNDayAccuracy(OutputDf, 90)
  )
}

# Step 9: prepare dataframe for union
PrepareDf <- function(Df, TypeName) {
  colnames(Df) <- c("Date", "Quantity")
  Df %>% mutate(Type = TypeName)
}

# Step 10: create all time df
CreateAllTimeDf <- function(HistoricalDf, YPredDf, CryptoTicker, Version) {
  PrepareDf(HistoricalDf, "Historical") %>%
    rbind(PrepareDf(YPredDf, "Forecast")) %>%
    mutate(Version = Version) %>%
    mutate(Ticker = CryptoTicker) %>%
    select(Ticker, Version, Type, Date, Quantity)
}

# Create a list of training versions to iterate through
VersionNames <- list("All", "24", "12", "6", "3", "1")
VersionObjects <- list(TrainAll, Train24, Train12, Train6, Train3, Train1)
VersionList <- list(VersionNames, VersionObjects)

# Create empty dfs for saving all of the results
AccuracyLog <- data.frame()
TimeSeries <- data.frame()

for (VersionNumber in c(1, 2, 3, 4, 5, 6)) {
  for (CryptoTicker in unique(VersionList[[2]][[VersionNumber]]$Ticker)) {
    for (CloseVariant in c("Close", "Close1", "Close2", "Close3")) {
      
      # Save the variables that are needed for this iteration
      CryptoDf <- VersionList[[2]][[VersionNumber]]
      TypeName <- VersionList[[1]][[VersionNumber]]
      TypeName <- paste0(TypeName, "Month", CloseVariant)
      SkipToNext <- FALSE
      
      # Perform all the steps and capture the results
      tryCatch({
        HistoricalDf <- CreateHistoricalDf(CryptoDf, CryptoTicker, CloseVariant)
        ProphetModel <- TrainProphetModel(HistoricalDf)
        FutureDf <- CreateFutureDf(ProphetModel, 365)
        YPredDf <- MakePredictions(ProphetModel, FutureDf)
        YTestDf <- CreateYTestDf(Crypto, CryptoTicker, Cutoff)
        OutputDf <- CreateOutputDf(YPredDf, YTestDf)
        AccuracyDf <- CreateAccuracyDf(OutputDf, CryptoTicker, TypeName)
        AccuracyLog <- rbind(AccuracyLog, AccuracyDf)
        AllTimeDf <- CreateAllTimeDf(HistoricalDf, YPredDf, CryptoTicker, TypeName)
        TimeSeries <- rbind(TimeSeries, AllTimeDf)
      }, error = function(e) {
        SkipToNext <- TRUE
      })
      
      if (SkipToNext) {next}
    }
  }
}

```

<p>&nbsp;</p>

##### **iv. Accuracy Results**

<p>&nbsp;</p>

<font size = "4">

</font>

```{r Accuracy Results, fig.height = 100, fig.width = 12, message = FALSE, warning = FALSE}

print(paste0("Average 7 Day Accuracy: ", mean(AccuracyLog$Accuracy7Day)))
print(paste0("Average 30 Day Accuracy: ", mean(AccuracyLog$Accuracy30Day)))
print(paste0("Average 90 Day Accuracy: ", mean(AccuracyLog$Accuracy90Day)))

AccuracyLog %>%
  melt(id = c("Ticker", "Version")) %>%
  mutate(value = ifelse(value < 0, 0, value)) %>%
  select(Version, variable, value) %>%
  ggplot(aes(x = value)) +
  geom_histogram(color = "black", fill = "darkgrey", bins = 25) +
  xlim(c(0, 1)) +
  xlab("\nAccuracy Score\n") +
  ylab("\nCount\n") +
  facet_wrap(~ Version + variable, nrow = 24, ncol = 3, scales = "free") +
  theme(axis.text.x = element_text(size = 9.5, angle = 45, hjust = 1),
      axis.text.y = element_text(size = 9.5),
      axis.title.x = element_text(size = 10, color = "black", face = "bold"),
      axis.title.y = element_text(size = 10, color = "black", face = "bold"),
      title = element_text(size = 11, color = "black", face = "bold"),
      plot.caption = element_text(size = 9, color = "black", face = "italic"),
      panel.background = element_rect(fill = NA))
  
```

<p>&nbsp;</p>

##### **v. Bad Accuracy Example**

<p>&nbsp;</p>

<font size = "4">

</font>

```{r Bad Accuracy Example, message = FALSE, warning = FALSE, fig.height = 8, fig.width = 12}

ExampleData <- TimeSeries %>%
  filter(Ticker == "BTC") %>%
  filter(Version == "AllMonthClose") %>%
  select(Type, Date, Quantity) %>%
  mutate(Forecast = Quantity) %>%
  rename(Historical = Quantity) %>%
  mutate(Historical = ifelse(Type == "Historical", Historical, NA)) %>%
  mutate(Forecast = ifelse(Type == "Forecast", Forecast, NA))

ggplot(ExampleData, aes(x = as.Date(Date), y = Historical, width = 0.7, fill = Type)) +
geom_bar(stat = "identity", position = "dodge", colour = "dodgerblue4") +
geom_line(data = ExampleData, aes(x = as.Date(Date), y = Forecast, group = 1), size = 1.1, colour = "darkorange2") +
scale_fill_manual("legend", values = c("Historical" = "dodgerblue4", "Forecast" = "darkorange2")) +
ggtitle("\nExample of Cryptocurrency Forecast with Bad Accuracy\n") +
ylab("\nActual / Forecast\n") +
xlab("\nDate\n") +
theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1),
      axis.text.y = element_text(size = 8),
      axis.title.x = element_text(size = 10, color = "black", face = "bold"),
      axis.title.y = element_text(size = 10, color = "black", face = "bold"),
      title = element_text(size = 11, color = "black", face = "bold"),
      plot.caption = element_text(size = 9, color = "black", face = "bold"),
      panel.border = element_rect(color = "grey", fill = NA, size = 1),
      panel.background = element_rect(fill = NA),
      legend.position = c(0.5, .931),
      legend.direction = "horizontal",
      legend.title = element_blank())
  
```

<p>&nbsp;</p>

##### **vi. OK Accuracy Example**

<p>&nbsp;</p>

<font size = "4">

</font>

```{r OK Accuracy Example, message = FALSE, warning = FALSE}

ExampleData <- TimeSeries %>%
  filter(Ticker == "API3") %>%
  filter(Version == "AllMonthClose") %>%
  select(Type, Date, Quantity) %>%
  mutate(Forecast = Quantity) %>%
  rename(Historical = Quantity) %>%
  mutate(Historical = ifelse(Type == "Historical", Historical, NA)) %>%
  mutate(Forecast = ifelse(Type == "Forecast", Forecast, NA))

ggplot(ExampleData, aes(x = as.Date(Date), y = Historical, width = 0.7, fill = Type)) +
geom_bar(stat = "identity", position = "dodge", colour = "dodgerblue4") +
geom_line(data = ExampleData, aes(x = as.Date(Date), y = Forecast, group = 1), size = 1.1, colour = "darkorange2") +
scale_fill_manual("legend", values = c("Historical" = "dodgerblue4", "Forecast" = "darkorange2")) +
ggtitle("\nExample of Cryptocurrency Forecast with OK Accuracy\n") +
ylab("\nActual / Forecast\n") +
xlab("\nDate\n") +
theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1),
      axis.text.y = element_text(size = 8),
      axis.title.x = element_text(size = 10, color = "black", face = "bold"),
      axis.title.y = element_text(size = 10, color = "black", face = "bold"),
      title = element_text(size = 11, color = "black", face = "bold"),
      plot.caption = element_text(size = 9, color = "black", face = "bold"),
      panel.border = element_rect(color = "grey", fill = NA, size = 1),
      panel.background = element_rect(fill = NA),
      legend.position = c(0.5, .931),
      legend.direction = "horizontal",
      legend.title = element_blank())

```

<p>&nbsp;</p>

##### **vii. Good Accuracy Example**

<p>&nbsp;</p>

<font size = "4">

</font>

```{r Good Accuracy Example, message = FALSE, warning = FALSE}

ExampleData <- TimeSeries %>%
  filter(Ticker == "BAL") %>%
  filter(Version == "AllMonthClose") %>%
  select(Type, Date, Quantity) %>%
  mutate(Forecast = Quantity) %>%
  rename(Historical = Quantity) %>%
  mutate(Historical = ifelse(Type == "Historical", Historical, NA)) %>%
  mutate(Forecast = ifelse(Type == "Forecast", Forecast, NA))

ggplot(ExampleData, aes(x = as.Date(Date), y = Historical, width = 0.7, fill = Type)) +
geom_bar(stat = "identity", position = "dodge", colour = "dodgerblue4") +
geom_line(data = ExampleData, aes(x = as.Date(Date), y = Forecast, group = 1), size = 1.1, colour = "darkorange2") +
scale_fill_manual("legend", values = c("Historical" = "dodgerblue4", "Forecast" = "darkorange2")) +
ggtitle("\nExample of Cryptocurrency Forecast with Good Accuracy\n") +
ylab("\nActual / Forecast\n") +
xlab("\nDate\n") +
theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1),
      axis.text.y = element_text(size = 8),
      axis.title.x = element_text(size = 10, color = "black", face = "bold"),
      axis.title.y = element_text(size = 10, color = "black", face = "bold"),
      title = element_text(size = 11, color = "black", face = "bold"),
      plot.caption = element_text(size = 9, color = "black", face = "bold"),
      panel.border = element_rect(color = "grey", fill = NA, size = 1),
      panel.background = element_rect(fill = NA),
      legend.position = c(0.2, .931),
      legend.direction = "horizontal",
      legend.title = element_blank())

```

<p>&nbsp;</p>

#### **Conclusion**

<p>&nbsp;</p>

<font size = "4">

In this project, I took a deep dive into the world of crypto and generated
forecasts for various cryptocurrencies including everything from data
collection, data cleaning, data exploration, and data modeling.

The results show that some cryptocurrencies are more forecastable than others.
I found that one of the biggest challenges was the volatility of the prices.
Most cryptocurrency prices have been subject to extreme fluctuations in the
past, and even though this is beginning to stabalize now that crypto has become
more mainstream, it still makes it difficult to forecast.

I would have liked to fine tune the model more, but unfortunately I ran out of
time. I think it would be interesting to see what would happen if I tried
adjusting some of the parameters that the prophet package has to offer. Maybe
I could a forecast with more ups and downs instead of a straight line.

Nonetheless, I accomplished my goal in this project by learning more about
cryptocurrency and further developed my data science skills in the process. I
feel that the code I wrote was some of the more sophisticated code that I have
wrote, including using python modules and complicated pipe operations in R. I
hope to continue to build on this and develop my own algorithmic trading 
application in the future.

</font>

<p>&nbsp;</p>