---
output:
  word_document: default
---
## Final Project Predicting Song Popularity

By Annie Titus, Tyler Gigot, Daniel Stalica

Syracuse University - IST 707 - Spring 2021

https://www.kaggle.com/edalrami/19000-spotify-songs/discussion/73524
<p>&nbsp;</p>

## Project Introduction

The music industry has evolved a lot over the last several decades and will continue
to evolve. This is in part due to advancements in technology. It is hard to believe
that not too long ago CD players were still commonplace for listening to music,
while nowadays, companies such as Spotify provide on demand streaming of music directly
to any device. Not only has the method of listening to music changed due to technology,
but the sound of the music itself has changed as well. For example, a lot of recordings
are now digitally enhanced and new genres have emerged based on pure digital production
such as electronic dance music and dubstep.

More recent developments in technology allow for attributes about songs to be extracted from
an audio file. For example, the Organize Your Music tool, is an online tool that takes
a song or a selection of songs and gathers information on them. It gathers attributes such as
the song tempo, energy, and danceability, among others. This enables a new perspective
of music to be explored and analyzed, which leads into the purpose of this project.

The following project will explore a Kaggle dataset consisting of thousands of songs with
these attributes. Here are some examples of questions that will attempt to be answered.
What makes a song popular? Are there certain attributes that are correlated with popular
songs? How have popular songs changed overtime? Can the popularity of a song be predicted
based on its attributes? The primary objective will be predicting the popularity of a song.

Song popularity is important because having a popular song can mean many different things to an
artist. One benefit it has is the influence and trajectory it can take an artists' career. One
song can make or break a new artist. Many artists strive to go 'viral' with a song that can bring
the spotlight that they are looking for. Producers of music and labels can also benefit
from understanding what makes a song popular. Having that knowledge can help them create a
hit song, and thus, improve their profitability and chances of breeding successful artists.

In the era of the internet and social media any song could pop off at any time if it has
the right sound. As the music industry and consumer listening preferences continue to change,
it is very important for the creators of music to stay on top of these trends and prepare for them.

\newpage

## Data Preparation

Two original datasets from Kaggle, song_data and song_info, are used along with
two additional datasets, master_song_T and master_artist_T, which were collected
via web scraping wikepedia.

Web scraping was conducted to collect additional data about the songs
and artists which is not included in the original datasets from Kaggle. For example,
when a song was released is a crucial piece of information that was missing.

The code for the web scraping is not provided in the output of this report,
however it is available in the source code. As mentioned before, the data
from the web scraping has been packaged into two csv files for convenience.

The goal of this section is to provide the reader with the necessary background
of the data. In the following section, each of the datasets will be described in more
detail along with any data preprocessing that took place. 

```{r Load Packages, include = FALSE}

# here are the packages needed and what they are used for
# make sure that these packages are installed prior to running

require(readr)               # read in the csv files
require(dplyr)               # mutate_all, pipe operator, joins
require(ggplot2)             # ggplot visualizations
require(stringr)             # string operations
require(arules)              # association rules mining
require(sqldf)               # some sql sauce
require(knitr)               # print out better looking tables
require(reshape2)            # casting and melting
require(Boruta)              # variable significance testing
require(class)               # knn model
require(caret)               # createDataPartition function
require(corrplot)            # correlation plot
require(correlationfunnel)   # binarization function
require(rattle)              # fancy r decision tree plot
require(C50)                 # decision tree modeling
require(randomForest)        # random forest modeling
require(neuralnet)           # neural network modeling
require(rpart)               # decision tree modeling
require(rpart.plot)          # decision tree visualization
require(e1071)               # naive bayes
require(caret)               # create data partition function
require(factoextra)          # cluster visualization

```

```{r View Struc Function, include = FALSE}

# custom function to view dataframe structure
view_struc <- function(df) {
  
  # start off with an empty dataframe
  temp_var <- data.frame()
  
  # store the col names of the dataframe
  df_cols <- colnames(df)
  
  # iterate through each column
  for (x in df_cols) {
    
    # name of the column
    name <- x
    
    # data type of the column
    type <- data.class(df[, which(colnames(df) == x)])
    
    # concatenate name and type
    line <- paste0(name, ': ', type)
    
    # append to the main dataframe
    temp_var <- rbind(temp_var, line)
  }

# print out a line delimiter above
print('~~~~~~~~~~~~~~~~~~~~~~~')

# print out the dataframe structure
for (row in temp_var[1:nrow(temp_var),]) {
  print(row)}

# print out a line delimiter below
print('~~~~~~~~~~~~~~~~~~~~~~~')

}

```

```{r Load Data}

#---------------------------------------------------------------------------#

# load the data
# make sure the files are saved in current working directory

# the first two data sets can be downloaded from kaggle
# https://www.kaggle.com/edalrami/19000-spotify-songs/discussion/73524

song_data         <- read.csv("song_data.csv")
song_info         <- read.csv("song_info.csv")

# the second two datasets are provided seperately
# these two datasets come from web scraping wikipedia

master_artist_T   <- read.csv("master_artist_T.csv")
master_song_T     <- read.csv("master_song_T.csv")

```

```{r Delineator 1}

#---------------------------------------------------------------------------#

```

```{r Song Data Header, echo = FALSE, results = 'asis'}

# print out the dataframe name in boldface

cat(paste("**", 'table: ', deparse(substitute(song_data)), "**", sep=""))
    
```

The song_data table is an original dataset from Kaggle, which consists of a
collection of songs that were parsed via the Organize Your Music tool.

Note: there were many duplicate song names, without a way of uniquely identifying
them. For this reason, duplicate song names were removed.

```{r Delineator 2}

#---------------------------------------------------------------------------#

```

```{r Song Data Original Dim, echo = FALSE}

# print out the dimensions of the original dataframe

cat('Original Table Dimensions \n',
  paste(nrow(song_data), ' rows by ', ncol(song_data), ' cols'))
    
```

```{r Clean Song Data, echo = FALSE}

# sql sauce to get the count of each song

song_counts <- as.data.frame(sqldf('
  select
    song_name,
    count(song_name) as count
  from
    song_data
  group by
    song_name'
))

# join the song data and song counts table

song_data <- left_join(
  song_data,
  song_counts,
  by = c('song_name' = 'song_name')
)

# remove any song with a count > 1

song_data <- song_data[which(
  song_data$count == 1), 1:15]

# print out the dimensions of the cleaned dataframe

cat('Cleaned Table Dimensions \n',
  paste(nrow(song_data), ' rows by ', ncol(song_data), ' cols'), '\n \n',
  'Removed', nrow(read.csv("song_data.csv")) - nrow(song_data), 'Duplicate song_name')

# clean up the environment

rm(song_counts)

```

```{r Song Data Struc, echo = FALSE}

# print out the structure of the dataframe
view_struc(song_data)

```

```{r Delineator 3}

#---------------------------------------------------------------------------#

```

```{r Song Data Metadata}

# meta data from http://organizeyourmusic.playlistmachinery.com/

# song_name         - the name of the song
# song_popularity   - the higher the value the more popular the song is
# song_duration_ms  - the length of the song measured in milliseconds
# acousticness      - the higher the value the more acoustic the song is
# danceability      - the higher the value, the easier it is to dance to
# energy            - the higher the value, the more energtic the song is
# instrumentalness  - the higher the value, the more instrumental the song is
# key               - description not provided
# liveness          - the higher the value, more likely a live recording
# loudness          - the higher the value, the louder, measured in dB
# audio_mode        - description not provided
# speechiness       - the higher the value the more spoken word in the song
# tempo             - the tempo of the song measured in beats per minute
# time_signature    - description not provided
# audio_valence     - the higher the value, the more positive mood

```

```{r Delineator 4}

#---------------------------------------------------------------------------#

```

```{r Song Info Header, echo = FALSE, results = 'asis'}

# print out the dataframe name in boldface

cat(paste("**", 'table: ', deparse(substitute(song_info)), "**", sep=""))
    
```

The song_info table is an original dataset from Kaggle which contains the
corresponding artist, album, and playlist for each song in the song_data table.

Note: as in the song_data table, there were duplication errors in this table
as well. This is resolved in the same manner by removing duplicate song names.

```{r Delineator 5}

#---------------------------------------------------------------------------#

```

```{r Song Info Original Dim, include = FALSE}

#---------------------------------------------------------------------------#

# print out the discovery of how many unique there are of each attribute

cat('Unique artist_name:',    length(unique(song_info$artist_name)), '\n',   # unique artists     = 7,564
    'Unique album_name:',     length(unique(song_info$album_names)), '\n',   # unique albums      = 12,014
    'Unique playlist:',       length(unique(song_info$playlist)))            # unique playlists   = 300

#---------------------------------------------------------------------------#

# what identifies a song as unique?

cat('Unique song_name | artist_name:',                           dim(unique(song_info[,1:2])) [1], '\n',          # unique song-artist                  = 14,006
    'Unique song_name | artist_name | album_name:',              dim(unique(song_info[,1:3])) [1], '\n',          # unique song-artist-album            = 14,615
    'Unique song_name | artist_name | playlist:',                dim(unique(song_info[,c(1, 2, 4)])) [1], '\n',   # unique song-artist-playlist         = 18,813
    'Unique song_name | artist_name | album_name | playlist:',   dim(unique(song_info[,1:4])) [1], '\n')          # unique song-artist-album-playlist   = 18,828

#---------------------------------------------------------------------------#

# Observations
# - there are a number of duplicate song names in the data
# - the way to uniquely identify would be song-artist
# - but the song_data table does not have artist name

```

```{r Song Info Original Dimensions, echo = FALSE}

# print out the dimensions of the original dataframe

cat('\n Original Table Dimensions \n',
  paste(nrow(song_info), ' rows by ', ncol(song_info), ' cols'))
    
```

```{r Song Info Split, echo = FALSE}

# keep only the non duplicate songs

song_info <- song_info[song_info$song_name %in% song_data$song_name, ]

# print out the dimensions of the cleaned dataframe

cat('Cleaned Table Dimensions \n',
  paste(nrow(song_info), ' rows by ', ncol(song_info), ' cols'), '\n \n',
  'Removed', nrow(read.csv("song_info.csv")) - nrow(song_info), 'Duplicate Song Names')

```

```{r Song Info Struc, echo = FALSE}

# print out the structure of the dataframe
view_struc(song_info)

```

```{r Delineator 6}

#---------------------------------------------------------------------------#

```

```{r Song Info Metadata}

# song_name         - the name of the song
# artist_name       - the name of the corresponding artist(s)
# album_names       - the name of the corresponding album(s)
# playlist          - the name of the corresponding playlist(s)

```

```{r Delineator 7}

#---------------------------------------------------------------------------#

```

```{r Master Song T Header, echo=FALSE, results='asis'}

cat(paste("**", 'table: ', deparse(substitute(master_song_T)), "**", sep=""))

```

The master_song_T table was collected by web scraping wikipedia pages
for additional information about the songs.

```{r Delineator 8}

#---------------------------------------------------------------------------#

```

```{r Master Song T Dim, echo=FALSE}

# remove duplicate songs from the master song table
# this table was created prior to removal of duplicate songs

master_song_T <- master_song_T[master_song_T$song_name %in% song_data$song_name, ]
master_song_T <- master_song_T[, -which(colnames(master_song_T) %in% c('song_id', 'artist_name'))]

# print out the dimensions of the master song table

cat(paste(nrow(master_song_T), ' rows by ', ncol(master_song_T), ' cols'))
    
```

```{r Master Song T Struc, echo=FALSE}

view_struc(master_song_T)

```

```{r Delineator 9}

#---------------------------------------------------------------------------#

```

\newpage

```{r Delineator 10}

#---------------------------------------------------------------------------#

```

```{r Master Song T Metadata}

# song_name         - the name of the song
# song_single       - binary whether the song is a single or not
# song_released     - the year that the song was released in
# song_genre        - the corresponding genre(s) for the song
# song_label        - the corresponding label(s) for the song
# song_songwriter   - the corresponding songwriter(s) for the song
# song_producer     - the corresponding producer(s) for the song

```

```{r Delineator 11}

#---------------------------------------------------------------------------#

```

```{r Master Artist T Header, echo=FALSE, results='asis'}

cat(paste("**", 'table: ', deparse(substitute(master_artist_T)), "**", sep=""))

```

The master_artist_T table was constructed by web scraping wikipedia pages
for additional artist information.

```{r Delineator 12}

#---------------------------------------------------------------------------#

```

```{r Master Artist T Dim, echo=FALSE}

# print the dimensiosn of the master artist table

cat(paste(nrow(master_artist_T), ' rows by ', ncol(master_artist_T), ' cols'))
    
```

```{r Master Artist T Struc, echo=FALSE}

view_struc(master_artist_T)

```

```{r Delineator 13}

#---------------------------------------------------------------------------#

```

```{r Master Artist T Metadata}

# artist_name       - the name of the artist
# birthday          - the date that the artist was born
# country           - 2 values - either USA or foreign
# startyear         - the year that the artist started making music

```

```{r Delineator 14}

#---------------------------------------------------------------------------#

```

```{r Plot Database ERD, include = FALSE}
 
# clean up the environment
rm(view_struc)

```

\newpage

##### Song Release Year

The song release year was one of the attributes that was retrieved from
web scraping Wikipedia. Here is how the distribution of the song release years looks.

Note: there are a large number of missing song years due to the data
not being available on Wikipedia which are not displayed on the graph below.

```{r Song Year Frequency, echo=FALSE, fig.height=5, fig.width=10}

plot(data.frame(table(master_song_T$song_released, useNA = c("always"))),
  main = 'Frequency of Number of Songs by Year',
  xlab = 'Song Release Year')
text(x = 85, y = 600, '2018')
text(x = 84, y = 410, '2017')

```

**Observations**  
- there are more newer generation songs than older generation songs  
- there is a particularly larger number of songs between 2010 and 2020  
- there is a large spike in the number of songs released in 2017-2018  
- the bias is probably due to the preferences of the creator of the data  
- there are some outliers, such as year 712, which is not a valid year

<p>&nbsp;</p>

```{r Delineator 15}

#---------------------------------------------------------------------------#

```

```{r Remove Outlier Years}

# years prior to 1970 and later than 2020 thrown out due to lack of data
# this will also take care of any outliers outside of the date range
# NA's will be kept because although the year is not known there a lot of data

cat('Number of Songs Removed:',
    nrow(master_song_T) - nrow(master_song_T[which(
                            (master_song_T$song_released >= 1970  &
                            master_song_T$song_released <= 2020) |
                            is.na(master_song_T$song_released)), ]))

```

```{r Delineator 16}

#---------------------------------------------------------------------------#

```

```{r Min Max Transformation Function, include = FALSE}

# create a function to standardize using a min max transformation
# the dataframe argument is the dataframe containing the columns to transform
# the columns argument is a numerical vector of which columns to transform
min_max_transform <- function(dataframe, columns) {
  
  # create a new version that will be the standardized version
  standardized <- dataframe   # first create a copy
  
  # iterate through each of the columns from the columns argument
  for (column in columns) {
    
    # temporarily store the values for the column
    temp_column <- unlist(as.numeric(standardized[, column]))
    
    # temporarily store the min and max for the column 
    temp_min <- min(temp_column)
    temp_max <- max(temp_column)
    
    # transform the column based on the min max transformation
    new_column <- (temp_column - temp_min) / (temp_max - temp_min)
    
    # change the column to the transformed column
    standardized[, column] <- new_column
  
  }
  
  # return the standardized dataframe as the output of the function
  return(standardized)
  
}

```

##### Discretize Song Release Year into Song Decade

The song years will be discretized into bins containing one decade per bin. This
is based on (a) the assumption that decades of music tend to have similar sounds
and (b) inferring song decade rather than inferring song year will be more accurate.

```{r Discretize Song year, include = FALSE}

#---------------------------------------------------------------------------#

# create a new version with song year discretized
# seperate the ones with year and the ones without
# these tables will be joined back together

song_generations <- master_song_T[which(
  master_song_T$song_released %in% seq(1960, 2020, 1)), c(1, 3)]

song_unknownyear <- master_song_T[which(
  is.na(master_song_T$song_released)), c(1, 3)]

# used custom discretize for graphing purposes

# start off with an empty variable
song_year_discretized <- c()

# iterate through each of the songs
for (song_year in song_generations$song_released) {
  
  # use if else tests to bin the song year
  temp_song_year <- ifelse(
    song_year >= 1970 & song_year < 1980, '1970s', ifelse(
    song_year >= 1980 & song_year < 1990, '1980s', ifelse(
    song_year >= 1990 & song_year < 2000, '1990s', ifelse(
    song_year >= 2000 & song_year < 2010, '2000s', ifelse(
    song_year >= 2010 & song_year < 2020, '2010s', NA)))))
  
  # add the discretized song year to the main
  song_year_discretized <- c(song_year_discretized, temp_song_year)
  
}

# change song year to the discretized song year
song_generations$song_released <- song_year_discretized

# combine the two tables back into one table
new_song_T <- rbind(song_generations, song_unknownyear)

# change the ones with NA's to 'NA' as a character
# this is needed for graphing purposes later on

new_song_T[which(is.na(new_song_T$song_released)), 2] <- 'NA'

# join in the rest of the song characteristics
new_song_T <- left_join(new_song_T, master_song_T[, -3], by = ('song_name' = 'song_name'))

# clean up the environment
rm(song_name, song_year, temp_song_year, song_year_discretized, song_unknownyear, song_generations)

```

```{r Delineator 17}

#---------------------------------------------------------------------------#

```

```{r Show Discretized Years}

# look at the distribution of the new song years after discretizing

table(new_song_T$song_released)

```

```{r Delineator 18}

#---------------------------------------------------------------------------#

```

```{r Song Year Histogram, echo=FALSE, fig.height=5, fig.width=9}

# create a temporary dataframe for showing the histogram
# temporary dataframe is manipulated to make NA's <- 2025
# this makes it so that the NA's can be shown in the histogram
# this is then covered up by manually coding in x axis labels

temp_master_song_T <- master_song_T[which(
  (master_song_T$song_released >= 1970 &           # get rid of outlier years
  master_song_T$song_released <= 2021) |          # get rid of outlier years
  is.na(master_song_T$song_released)), c(1, 3)]   # include the NA's

temp_master_song_T[which(is.na(temp_master_song_T$song_released)), 2] <- 2030

# temporarily store the histogram without plotting it

temp_song_hist <- hist(temp_master_song_T[, 2], plot = FALSE, breaks = c(seq(1970, 2030, 10)))

# generate the histogram plot with some customizations

plot(
     temp_song_hist,
     xaxt = 'n',
     xlab = 'Song Decade',
     ylab = 'Count',
     main = 'Histogram of Number of Songs By Year Discretized',
     col = 'grey'
    )

# remove the x labels

axis(
     1,
     temp_song_hist$mids,
     labels = replicate(6, ''),
     tick = FALSE,
     padj= -1.5
     )

# add custom x labels

text(x = 1975, y = 774, '1970s', cex = 1.2)
text(x = 1985, y = 840, '1980s', cex = 1.2)
text(x = 1995, y = 989, '1990s', cex = 1.2)
text(x = 2005, y = 1090, '2000s', cex = 1.2)
text(x = 2015, y = 2463, '2010s', cex = 1.2)
text(x = 2025, y = 2250, 'NA', cex = 1.3)

# clean up the environment

rm(temp_master_song_T, temp_song_hist)

```

**Observations**  
- as mentioned before, there are a lot of unknown song years, which are shown here  
- the NA values are due to the information not being available on Wikipedia  
- it may be possible to infer the missing song decade 
- having the genre would be helpful for this, so the genre will be explored next

\newpage

##### Song Decade and Song Popularity

Now that the song release years have been discretized into song decades, the
following chart compares the average song popularity for each song decade.

```{r Average Song Popularity by Song Decade, echo=FALSE, fig.height=5, fig.width=9}

# join the new_song_T and song_data tables together

temp_song_T <-
  
  left_join(
    new_song_T,
    song_data,
    by = c('song_name' = 'song_name'))

# query the table to be used for the subsequent plot

temp_song_T <-

  sqldf('
    select
      song_released,
      round(avg(song_popularity), 2) as avg_song_popularity
    from
      temp_song_T
    group by
      song_released')

# show a plot that answers the question

ggplot(temp_song_T, aes(
  x = song_released,
  y = avg_song_popularity)) +
  geom_col(width = .5) +
  ggtitle('\n Average Song Popularity by Song Decade \n') +
  ylab('\n Average Song Popularity \n') +
  xlab('\n Song Decade \n') +

  # some more customizations to the plot
  theme(
        
        # change the font size of x axis and y axis labels to 11
        axis.text.x = element_text(size = 9), 
        axis.text.y = element_text(size = 9),
        
        # change the x axis and y axis titles to size 12, black, and boldface
        axis.title.x = element_text(size = 10, color = "black", face = "bold"),
        axis.title.y = element_text(size = 10, color = "black", face = "bold"),
        
        # get rid of the default background
        panel.background = element_rect(fill = NA)
        
        )

```

**Observations**  
- there is a slight positive correlation between song decade and song popularity  
- the most popular songs are in the 2010s, the least popular songs are in the NA  
- this makes sense because songs that do not have information on Wikipedia  
are probably less popular

<p>&nbsp;</p>

##### Song Genres

The song genre was one of the attributes that was retrieved from web scraping
Wikipedia. Note: there can be multiple genres for one song.

There was a wide range of genres that came out of Wikipedia,
but these were aggregated into higher level parent genres.

```{r Show Genres Sample}

# here is the list of all of the parent genres
# note that hip and hop will get put together into hiphop

parent_genre_list <- c('alt', 'rock', 'metal', 'punk', 'pop', 'hip', 'hop', 'r&b', 'rap', 'jazz', 'blues', 'folk', 'country', 'elect', 'other')

```

```{r Reconcile Tables Function, include = FALSE}

# reconcile tables function
# sets up the table column names to be aligned dynamically for row binding

reconcile_tables <- function(table1, table2) {
  
  colnameslist <- unique(c(colnames(table1), colnames(table2))) # unique list of all columns from both tables
  
  table1rows <- nrow(table1) # num of rows in table 1
  table2rows <- nrow(table2) # num of rows in table 2
  
  table1addcols <- colnames(table2) [which( ! colnames(table2) %in% colnames(table1))] # cols in table 2 and not in table 1
  table2addcols <- colnames(table1) [which( ! colnames(table1) %in% colnames(table2))] # cols in table 1 and not in table 2
  
  table1pad <- as.data.frame(matrix(nrow = table1rows, ncol = length(table1addcols))) # create padded data structure table 1
  table2pad <- as.data.frame(matrix(nrow = table2rows, ncol = length(table2addcols))) # create padded data structure table 1
  
  colnames(table1pad) <- table1addcols # change the col names accordingly table1
  colnames(table2pad) <- table2addcols # change the col names accordingly table1
  
  newtable1 <- cbind(table1, table1pad) # combine the original table 1 and the padded table 1
  newtable2 <- cbind(table2, table2pad) # combine the original table 2 and the padded table 2
  
  newtable1 <- newtable1[, colnameslist] # reorder the columns accordingly
  newtable2 <- newtable2[, colnameslist] # reorder the columns accordingly
  
  return(rbind(newtable1, newtable2)) # return the combined tables

}

```

```{r Unpack Genres, include = FALSE}

# split the data into two dataframes
# one has the known genres one does not

song_known_genres <- new_song_T[which(! is.na(new_song_T$song_genre)), c(1, 4)]

# create a vector of each of the genre counts
# these are all of the genres that are present in the data
# the other genre is anything that is out of the other genres

parent_genre_list <- c('alt', 'rock', 'metal', 'punk', 'pop', 'hip', 'hop', 'r&b',
                       'rap', 'jazz', 'blues', 'folk', 'country', 'elect', 'other')

# create an empty variable for each of the genres
# these will be appended to throughout the for loop

altcount     <- c();   rockcount    <- c();   metalcount     <- c()
punkcount    <- c();   popcount     <- c();   hiphopcount    <- c()
rbcount      <- c();   rapcount     <- c();   jazzcount      <- c()
bluescount   <- c();   folkcount    <- c();   countrycount   <- c()
electcount   <- c();   othercount   <- c()

# iterate through each of the songs and corresponding genres
# if it contains the genre it will be given a 1 for true
# if does not contain the genre it wil be given a 0 for false

for (nestedgenre in song_known_genres$song_genre) {
  
  alttest      <- as.numeric(grepl('alt', nestedgenre)); altcount <- c(altcount, alttest)
  rocktest     <- as.numeric(grepl('rock', nestedgenre)); rockcount <- c(rockcount, rocktest)
  metaltest    <- as.numeric(grepl('metal', nestedgenre)); metalcount <- c(metalcount, metaltest)
  punktest     <- as.numeric(grepl('punk', nestedgenre)); punkcount <- c(punkcount, punktest)
  poptest      <- as.numeric(grepl('pop', nestedgenre)); popcount <- c(popcount, poptest)
  rbtest       <- as.numeric(grepl('r&b', nestedgenre)); rbcount <- c(rbcount, rbtest)
  raptest      <- as.numeric(grepl('rap', nestedgenre)); rapcount <- c(rapcount, raptest)
  jazztest     <- as.numeric(grepl('jazz', nestedgenre)); jazzcount <- c(jazzcount, jazztest)
  bluestest    <- as.numeric(grepl('blues', nestedgenre)); bluescount <- c(bluescount, bluestest)
  folktest     <- as.numeric(grepl('folk', nestedgenre)); folkcount <- c(folkcount, folktest)
  countrytest  <- as.numeric(grepl('country', nestedgenre)); countrycount <- c(countrycount, countrytest)
  electtest    <- as.numeric(grepl('elect', nestedgenre)); electcount <- c(electcount, electtest)
  othertest    <- as.numeric(grepl('other', nestedgenre)); othercount <- c(othercount, othertest)
  hiphoptest   <- as.numeric(grepl('hip', nestedgenre) | grepl('hop', nestedgenre)); hiphopcount <- c(hiphopcount, hiphoptest)

}

# join these vectors into the main dataframe

song_known_genres$alt      <- altcount;     song_known_genres$rock     <- rockcount;   song_known_genres$metal    <- metalcount
song_known_genres$punk     <- punkcount;    song_known_genres$pop      <- popcount;    song_known_genres$hiphop   <- hiphopcount
song_known_genres$rb       <- rbcount;      song_known_genres$rap      <- rapcount;    song_known_genres$jazz     <- jazzcount
song_known_genres$blues    <- bluescount;   song_known_genres$folk     <- folkcount;   song_known_genres$country  <- countrycount
song_known_genres$elect    <- electcount;   song_known_genres$other    <- othercount

# create a dataframe for the genres that are unknown

song_unknown_genres <- new_song_T[which(is.na(new_song_T$song_genre)), c(1, 4)]

# join together the two data frames
# NAs for each of the genres for the unkown genres

song_temp_genres <- reconcile_tables(song_known_genres, song_unknown_genres)

# change the NA into its own column

song_temp_genres$Unknown <- ifelse(is.na(song_temp_genres$song_genre), 1, 0)

#correct any NA's to zeros

for (column in colnames(song_temp_genres[,2:(dim(song_temp_genres)[2])])) {
song_temp_genres[which(is.na(song_temp_genres[,which(colnames(song_temp_genres)==column)])),
which(colnames(song_temp_genres)==column)] <- 0}

# clean up the transient variables from the environment

rm(alttest, rocktest, metaltest, punktest, poptest, hiphoptest, rbtest, raptest,
   jazztest, bluestest, folktest, countrytest, electtest, othertest, nestedgenre,
   altcount, rockcount, metalcount, punkcount, popcount, hiphopcount, rbcount,
   rapcount, jazzcount, bluescount, folkcount, countrycount, electcount, othercount,
   parent_genre_list, column)

```

```{r Prepare Song Genres Plot, include = FALSE}

# join in the song generation with the song genres dataframe

song_temp_genres <- left_join(song_temp_genres, new_song_T[, c(1, 2)], by = c('song_name' = 'song_name'))

# create a temporary variable for the distribution of genres by song generation

temp_analysis <- 
  
sqldf('
select
  song_released,
  sum(alt) as alt,
  sum(rock) as rock,
  sum(metal) as metal,
  sum(punk) as punk,
  sum(pop) as pop,
  sum(hiphop) as hiphop,
  sum(rb) as rb,
  sum(rap) as rap,
  sum(jazz) as jazz,
  sum(blues) as blues,
  sum(folk) as folk,
  sum(elect) as elect,
  sum(other) as other,
  sum(unknown) as unknown
from song_temp_genres
group by song_released')

# melt the data frame for plotting a stacked bar chart

temp_analysis <- melt(temp_analysis, variable.name = 'Genre', value.name = 'Count')

```

Below chart shows the distribution of the song genres and how many of them
there are for each song release year.

\newpage

```{r Plot Song Genres, echo=FALSE, fig.height=5, fig.width=10}

# what is the distribution of the song genres - by generation?
# keep in mind a song can have more than one genre

ggplot(temp_analysis, aes(reorder(Genre, -Count), Count)) +
  geom_col(position = 'stack', aes(fill = song_released), colour = 'black') +
  ggtitle('\n Distribution of Genres by Song Generation\n ') +
  ylab('\n Count \n') + xlab('\n Song Genre \n') +
  scale_fill_grey(start = 0, end = .9) +

  # some more customizations to the plot
  theme(
        
        # change the font size of x axis and y axis labels to 11
        axis.text.x = element_text(size = 9), 
        axis.text.y = element_text(size = 9),
        
        # change the x axis and y axis titles to size 12, black, and boldface
        axis.title.x = element_text(size = 10, color = "black", face = "bold"),
        axis.title.y = element_text(size = 10, color = "black", face = "bold"),
        
        # get rid of the default background
        panel.background = element_rect(fill = NA)
        
        )

```

```{r Song Genres Distribution Table, include = FALSE}

# create a lookup table for the count of all songs in each bin

song_count_lookup <- sqldf('

  select
    song_released,
    count(song_released) as AllCount
  from
    song_temp_genres
  group by
    song_released'
  
)

# join in this information to the temporary analysis table

temp_analysis <- left_join(temp_analysis, song_count_lookup, by = c('song_released' = 'song_released'))

# create a new column and the percentage total

temp_analysis$pctTotal <- paste0(round((temp_analysis$Count / temp_analysis$AllCount) * 100, 0), '%')

# restructure the dataframe into a more readable format

temp_analysis <- t(dcast(temp_analysis[, c(1, 2, 5)], song_released ~ Genre))

# change the column names accordingly
# due to the transpose operation the column names were placed in row 1

colnames(temp_analysis)   <- temp_analysis[1, ]    # use first row as headers
temp_analysis             <- temp_analysis[-1, ]   # get rid of the first row

```

Below table shows the percentage total of the song genres for each song year, 
note that the percentages will not add up to 100 due to being multiple
genres for some songs.

```{r Display Song Genres Distribution Table, echo = FALSE}

# print out a title seperated by 2 lines and then the table

kable(temp_analysis)

# clean up the environment

rm(song_count_lookup, temp_analysis)

```

<p>&nbsp;</p>

**Observations**  
- there are some noticeable correlations with genre and generation  
- for example not many hiphop or electro songs in earlier generations  
- there are more folk and rock songs in earlier gens  
- pop, rock, hiphop are some of the most frequent genres as would be expected
because these genres are mostly newer generation  

##### Song Genre and Song Popularity

The following series of charts displays the average song popularity for
each genre broken down by song decade.

```{r Song Genre and Song Popularity Charts, echo=FALSE, fig.height=12, fig.width=10}

# query the table that has the information
temp_song_T <- left_join(
                left_join(
                  song_temp_genres,
                  song_data,
                  by = c('song_name' = 'song_name')),
                new_song_T[, c(1, 3)],
                by = c('song_name' = 'song_name'))

# start with an empty dataframe, this will be appended to
temp_genre_T <- data.frame()

# create a character vector of summary statistics for each genre
# these will be placed in the label strip of each plot
# create an empty variable for the summary statistics text
genre_summary_statitics_text <- c()

# iterate through columns 3-17 in temp_song_T
# these are the columns for each of the genres
for (genreColNumber in seq(3, 17, 1)) {
  
  # create a subset table with the necessary columns
  temp_query_T <-
    temp_song_T[which(                          # is a subset table of the temp_song_T table
      temp_song_T[, genreColNumber] == 1),      # if the value = 1 then the song is that genre
      c(genreColNumber, 18, 19)]                # three columns, genre, song decade, song popularity
  
  # store the genre in all caps, the min, the mean, and the max
  # these variables will be used in a summary statistics character vector
  temp_genre_name        <- colnames(temp_song_T) [genreColNumber]
  temp_genre_pop_min     <- round(min(temp_query_T$song_popularity), 0)
  temp_genre_pop_mean    <- round(mean(temp_query_T$song_popularity), 0)
  temp_genre_pop_max     <- round(max(temp_query_T$song_popularity), 0)
  
  # summary statistics character vector for the current genre
  summary_statistics <- paste(       ' | ',
    str_to_upper(temp_genre_name),   ' | ',
    'min =', temp_genre_pop_min,     ' | ',
    'mean =', temp_genre_pop_mean,   ' | ',
    'max =', temp_genre_pop_max,     ' | ')
    
  # append the summary statistics character vector to the main
  genre_summary_statitics_text <- c(genre_summary_statitics_text, summary_statistics)

  # aggregate by song decade and avg song popularity
  temp_query_T <-
    sqldf('
      select
        song_released,
        round(avg(song_popularity), 2) as avg_song_popularity
      from
        temp_query_T
      group by
        song_released')
  
  # add another column for the song genre
  temp_query_T$genre <- colnames(temp_song_T) [genreColNumber]
  
  # append to the main dataframe
  temp_genre_T <- rbind(temp_genre_T, temp_query_T)
  
}

# create a character vector that can be passed as
# an argument in the facet_wrap function
genre_plot_titles <- c(
  
  alt       = genre_summary_statitics_text[1],
  rock      = genre_summary_statitics_text[2],
  metal     = genre_summary_statitics_text[3],
  punk      = genre_summary_statitics_text[4],
  pop       = genre_summary_statitics_text[5],
  hiphop    = genre_summary_statitics_text[6],
  rb        = genre_summary_statitics_text[7],
  rap       = genre_summary_statitics_text[8],
  jazz      = genre_summary_statitics_text[9],
  blues     = genre_summary_statitics_text[10], 
  folk      = genre_summary_statitics_text[11],
  country   = genre_summary_statitics_text[12],
  elect     = genre_summary_statitics_text[13], 
  other     = genre_summary_statitics_text[14], 
  Unknown   = genre_summary_statitics_text[15]
  
)
  
  # plot the avg song popularity by song decade
  ggplot(temp_genre_T,
    
    # song decade on the x axis
    # avg song popularity on the y axis
    aes(
      x = song_released,
      y = avg_song_popularity,
      fill = song_released)) +
    
    # reduce the width of the bars for looks
    geom_col(
      width = 0.5) +
    
    # set the limits between 0 and 100
    # song popularity is always between 0 and 100
    # every visual should display the same scale
    scale_y_continuous(
      limits=c(0,75)) +
    
    scale_fill_manual(
      values = c("dodgerblue4", "darkorange1", "gray76", "lightblue3", "darkorange3", 'gray47')) +
    
    # add title, x axis label, y axis label
    ggtitle('\n Avg Song Popularity by Genre and by Song Decade \n') +
    xlab('\n Song Decade \n') +
    ylab('\n Avg Song Popularity \n') +
    
    # adjust the legend
    theme(legend.position=c(.75, 1.06),
          legend.direction="horizontal") +
    
    # use facet wrap to partition by song genre
    # set layout to 5 rows by 3 columns
    facet_wrap(
      ~ genre,
      nrow = 5,
      ncol = 3, 
      labeller = labeller(
        genre = genre_plot_titles)
    )
  
# clean up the environment
rm(genre_plot_titles, genre_summary_statitics_text, genreColNumber, summary_statistics, temp_genre_T,
   temp_genre_name, temp_genre_pop_max, temp_genre_pop_min, temp_genre_pop_mean, temp_query_T)

```

**Observations**  
- rap is most popular genre (52 avg) and jazz is the least popular (38 avg)  
- some genres have decreased in popularity over time, such as alt and metal  
- some genres have increased in popularity over time, such as country and pop

##### Correlation of Song Attributes

The song attributes were included as part of the original Kaggle datasets.
These will be explored next.

```{r Delineator 19}

#---------------------------------------------------------------------------#

```

```{r Song Attribute Correlation Matrix, fig.height=3, fig.width=3}

# transform the song attributes with min max
temp_song_T_transformed <- min_max_transform(temp_song_T, seq(19, 32, 1))

# truncate the column names
colnames(temp_song_T_transformed) <- substr(colnames(temp_song_T_transformed), 1, 4)

# fix some of the column names manually
colnames(temp_song_T_transformed) [c(19, 20, 28, 32)] <- c('pop', 'dur', 'mode', 'vlnc')

# compute the correlation matrix
temp_cor <- round(cor(temp_song_T_transformed[, 19:32], temp_song_T_transformed[, 19:32]), 2)

# plot the correlation matrix
corrplot(temp_cor,
         method = 'ellipse',
         outline = TRUE)

```

\newpage

**Observations**  
- there are not any very strong correlations with song popularity  
- song acousticness has a strong negative correlation with song energy  
- song energy, loudness, and audio valence have a strong positive correlation

```{r Delineator 30}

#---------------------------------------------------------------------------#

```

##### Song Attribute Importance

Using the boruta package, the importance of each song attribute for predicting
song popularity will be tested.

```{r Song Attribute Boruta, include = FALSE}

# take a sample to reduce the amount of data
set.seed(1000)
temp.song.T.transformed.sampled <- temp_song_T_transformed[
  sample(seq(1, nrow(temp_song_T_transformed), 1), 2500, replace = FALSE), ]

# use boruta package to determine significance
set.seed(1000)
song.boruta <- Boruta(pop ~ .,
                      data = temp.song.T.transformed.sampled[, 19:32],
                      doTrace = 2,
                      pValue = 0.9,
                      maxRuns = 15)

```

```{r Song Attribute Boruta Results, echo = FALSE}

# print out the results of the boruta test
print(song.boruta)

# clean up the environment
rm(song.boruta, temp_cor, temp.song.T.transformed.sampled)

```

```{r Song Attribute Feature Selection}

# remove the variables that are not significant
temp_song_T_transformed <- temp_song_T_transformed[, which(
  ! colnames(temp_song_T_transformed) %in% c('key', 'live', 'mode', 'time'))]

```

**Observations**  
- key and audio mode are not significant for predicting song popularity  
- liveness and time signature are tentative for predicting song popularity  
- based on the results, all four of these columns will be dropped

```{r Delineator 20}

#---------------------------------------------------------------------------#

```

##### Exploration of Song Attributes

Each remaining song attribute will be explored in more detail along with
any data preprocessing that takes place.

```{r Song Single Question}

# what effect does a song being a single have on song popularity?

```

```{r Song Single Answer, echo = FALSE, fig.height=2.5, fig.width=6}

# change the value of song single that is NA to 0
temp_song_T[which(
  is.na(temp_song_T$song_single)),
  which(colnames(temp_song_T) == 'song_single')] <- 0

# query the table to answer the question
temp_query <- 

  sqldf('
    select
      song_released as song_rel,
      song_single as single,
      avg(song_popularity) as pop
    from
      temp_song_T
    group by
      song_released,
      song_single'
  )

# show a chart that answers the question
ggplot() +
  geom_bar(
    data = temp_query,
    aes(
      x = song_rel,
      y = pop,
      fill = as.factor(single)),
    position = "dodge",
    stat = "identity",
    width = 0.5
    ) +
  ggtitle('\n Song Single and Song Popularity \n') +
  ylab('\n Song Popularity \n') +
  xlab('\n Song Decade \n') + 
  
  # adjust the legend
  theme(legend.position=c(.75, 1.13),
        legend.direction="horizontal") +
  
  guides(fill=guide_legend("Single?"))

```

```{r Delineator 21}

#---------------------------------------------------------------------------#

```

**Observations**  
- there is a positive correlation between song single and song popularity  
- this is not surprising because singles are usually more popular songs  
- interestingly, there is little to no difference seen in the NA group

```{r Advanced Box Plot Function, echo=FALSE}

# create a function that can be used for generating the boxplot
advanced_boxplot <- function(song_attr_values,
                             song_attr_name,
                             song_year_values,
                             ylimhigh,
                             xheadercenter,
                             ablinehigh,
                             yupperheader) {
  
  # create a new dataframe with the two columns provided
  temp_dataframe <- data.frame(cbind(
    'SongAttr' = song_attr_values,
    'SongYear' = song_year_values))

  # create an empty variable for the summary text
  # this will be used as a header in the box plot
  boxplot_main <- c()
  
  # iterate through each box plot part in the summary function
  # i.e. min, Q1, median, mean Q3, max, etc.
  for (boxplot_part in names(summary(song_attr_values)) [1:6]) {
    
    # capture the numerical value of the current box plot part
    boxplot_value <- summary(song_attr_values) [boxplot_part]
    
    # concatenate the name and the numerical value seperated by a colon
    boxplot_full <- paste0(boxplot_part, ': ', round(boxplot_value, 1))
    
    # append the concatenated text to the main variable
    boxplot_main <- paste(boxplot_main, boxplot_full, '  |  ')
  
  }
  
  # create the box plot
  plot <- boxplot(
    as.numeric(temp_dataframe$SongAttr) ~ temp_dataframe$SongYear,
    main = paste('Song', song_attr_name, 'by Song Decade'),
    ylim = c((as.numeric(min(temp_dataframe$SongAttr))*1.1), ylimhigh),
    xlab = '',
    ylab = paste('Song', song_attr_name),
    las = 2)
  
  # add a solid horizontal line for the upper header
  abline(h = ablinehigh)
  
  # add a dashed horizontal line for the overall median
  abline(h = (summary(temp_dataframe$SongAttr) ['Median']), lty = 2)
  
  # add the text for the upper header
  text(
    x = xheadercenter,
    y = yupperheader,
    paste(boxplot_main),
    cex = .9
  )

  return(plot)

}

```

```{r Delineator 22}

#---------------------------------------------------------------------------#

```

```{r Song Popularity Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

popularityplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$pop.1,
                 song_attr_name = 'Popularity',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1.05,
                 yupperheader = 1.15)

hist(temp_song_T_transformed$pop.1,
     breaks = 50,
     main = 'Song Popularity Histogram',
     xlab = 'Song Popularity')

rm(popularityplot)

```

**Observations**  
- the distribution of song popularity is normal but slightly left skewed  
- this means that there are more less popular songs than there are popular songs  
- the majority of songs fall within a song popularity of 0.4 to 0.6  
- there is an unusual number of very unpopular songs in the 1990s and 2000s  

```{r Delineator 100}

#---------------------------------------------------------------------------#

```

```{r Retrieve Song Popularity Outliers for 1990s, echo = FALSE}

x.Q1 <- summary(temp_song_T_transformed[which(temp_song_T_transformed$song.2 == '1990s'),
                            which(colnames(temp_song_T_transformed) == 'pop.1')]) [2] %>% as.numeric()

x.Q3 <- summary(temp_song_T_transformed[which(temp_song_T_transformed$song.2 == '1990s'),
                            which(colnames(temp_song_T_transformed) == 'pop.1')]) [5] %>% as.numeric()

x.IQR <- x.Q3 - x.Q1

x.Outliers <- temp_song_T_transformed[which(temp_song_T_transformed$pop.1 < x.Q1 - (1.5 * x.IQR) &
                                            temp_song_T_transformed$song.2 == '1990s'), ]

x.Outliers <- left_join(x.Outliers, song_info, by = c('song' = 'song_name')) %>%
              left_join(master_song_T[, c(1, 2, 3, 5)], by = c('song' = 'song_name'))

```

```{r Show Song Popularity Outliers for 1990s}

# a sample of some of the outliers from 1990s

kable(head(x.Outliers[, c(1, 30, 31, 2, 33, 34)], 5))

```

```{r Delineator 101}

#---------------------------------------------------------------------------#

```

```{r Retrieve Song Popularity Outliers for 2000s, echo = FALSE}

x.Q1 <- summary(temp_song_T_transformed[which(temp_song_T_transformed$song.2 == '2000s'),
                            which(colnames(temp_song_T_transformed) == 'pop.1')]) [2] %>% as.numeric()

x.Q3 <- summary(temp_song_T_transformed[which(temp_song_T_transformed$song.2 == '2000s'),
                            which(colnames(temp_song_T_transformed) == 'pop.1')]) [5] %>% as.numeric()

x.IQR <- x.Q3 - x.Q1

x.Outliers <- temp_song_T_transformed[which(temp_song_T_transformed$pop.1 < x.Q1 - (1.5 * x.IQR) &
                                            temp_song_T_transformed$song.2 == '2000s'), ]

x.Outliers <- left_join(x.Outliers, song_info, by = c('song' = 'song_name')) %>%
              left_join(master_song_T[, c(1, 2, 3, 5)], by = c('song' = 'song_name'))

```

```{r Show Song Popularity Outliers for 2000s}

# a sample of some of the outliers from 2000s

kable(head(x.Outliers[, c(1, 30, 31, 2, 33, 34)], 5))

```

```{r Delineator 102}

#---------------------------------------------------------------------------#

```

**Observations**  
- some of these songs should not have such low song popularity  
- for example, "Ill Be There For You/Youre All I Need To Get By" was a
single by Method Man and sold over 800,000 copies.  
- these songs will be removed due to the potential adverse effect
that they could have on the prediction models later on.

```{r Delineator 103}

#---------------------------------------------------------------------------#

```

```{r Remove Outlier Songs Song Popularity, echo = FALSE}

# create a copy of the dataframe
# this will stay with none of the outliers removed
# might come in handy later for analysis purposes
x.copy <- temp_song_T_transformed

# cycle through twice
for (x.cycle in seq(1, 2, 1)) {
  
# iterate through each song decade
for (x.songdecade in c('1970s', '1980s', '1990s', '2000s', '2010s', 'NA')) {

# store first and third quartiles
x.Q1 <- as.numeric(summary(temp_song_T_transformed[which(temp_song_T_transformed$song.2 == x.songdecade), 19]) [2])
x.Q3 <- as.numeric(summary(temp_song_T_transformed[which(temp_song_T_transformed$song.2 == x.songdecade), 19]) [5])

# store upper and lower outliers
x.LowerOutlier <- x.Q1 - (1.5 * (x.Q3 - x.Q1))
x.UpperOutlier <- x.Q3 + (1.5 * (x.Q3 - x.Q1))

# remove the outliers
temp_song_T_transformed <- temp_song_T_transformed[which(
  
  # is a lower outlier for the current decade
  ! (temp_song_T_transformed$song.2 == x.songdecade &
  temp_song_T_transformed$pop.1 < x.LowerOutlier) &
    
  # or is an upper outlier for the current decade
  ! (temp_song_T_transformed$song.2 == x.songdecade &
  temp_song_T_transformed$pop.1  > x.UpperOutlier)), ]

}
  
}

# print how many were removed
cat('Number of Songs Removed:',
    nrow(x.copy) - nrow(temp_song_T_transformed))

```

```{r Cleaned Song Popularity Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

popularityplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$pop.1,
                 song_attr_name = 'Popularity',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1.05,
                 yupperheader = 1.15)

hist(temp_song_T_transformed$pop.1,
     breaks = 50,
     main = 'Song Popularity Histogram',
     xlab = 'Song Popularity')

rm(popularityplot)

```

**Observations**  
- the distribution of song popularity after removing outliers is more normal  
- there are still some strange pockets of low popularity but will leave for now

```{r Delineator 23}

#---------------------------------------------------------------------------#

```

```{r Song Duration Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

durationplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$dur,
                 song_attr_name = 'Duration',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1.05,
                 yupperheader = 1.15)

hist(temp_song_T_transformed$dur,
     breaks = 50,
     main = 'Song Duration Histogram',
     xlab = 'Song Duration')

rm(durationplot)

```

```{r Top 10 Shortest Songs}

# a sample of some of the shortest songs
kable(
  head(
    sqldf('
      select
        song_name,
        (sum(song_duration_ms) / 1000) as song_duration_seconds
      from
        temp_song_T
      group by
        song_name
      order by
        song_duration_ms asc'
    )
  )
)

```

```{r Top 10 Longest Songs}

# a sample of some of the longest songs
kable(
  head(
    sqldf('
      select
        song_name,
        (sum(song_duration_ms) / 1000) as song_duration_seconds
      from
        temp_song_T
      group by
        song_name
      order by
        song_duration_ms desc'
    )
  )
)

```

```{r Delineator 104}

#---------------------------------------------------------------------------#

```

**Observations**  
- there is little difference in avg song duration between decades  
- there are some outlier songs in terms of their song duration  
- some are either intros, outros, transitions, or not real songs  
- or they could just be really short or really long songs  
- these songs can be removed since the main focus is hit songs

```{r Delineator 105}

#---------------------------------------------------------------------------#

```

```{r Remove Outlier Songs Song Duration, echo = FALSE}

# store first and third quartiles
Q1 <- as.numeric(summary(temp_song_T_transformed$dur) [2])
Q3 <- as.numeric(summary(temp_song_T_transformed$dur) [5])

# store upper and lower outliers
LowerOutlier <- Q1 - (1.5 * (Q3 - Q1))
UpperOutlier <- Q3 + (1.5 * (Q3 - Q1))

# remove the outliers
temp_song_T_transformed <- temp_song_T_transformed[which(
  temp_song_T_transformed$dur > LowerOutlier &
  temp_song_T_transformed$dur < UpperOutlier), ]

# print how many were removed
cat('Number of Songs Removed:',
    nrow(temp_song_T) - nrow(temp_song_T_transformed))

# clean up the environment
rm(Q1, Q3, LowerOutlier, UpperOutlier)

```

```{r Cleaned Song Duration Analysis, echo = FALSE, fig.height=5, fig.width=14}

# reapply the min max transformation
temp_song_T_transformed <- min_max_transform(temp_song_T_transformed, 20)

par(mfrow = c(1, 2))

durationplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$dur,
                 song_attr_name = 'Duration',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1.05,
                 yupperheader = 1.15)

hist(temp_song_T_transformed$dur,
     breaks = 50,
     main = 'Song Duration Histogram',
     xlab = 'Song Duration')

rm(durationplot)

```

**Observations**  
- the distribution of song duration after removing outliers is normal  
- the new range of song duration is between about 1.5 min to 5.5 min  
- the majority of songs fall tend to be around 3 to 4 minutes  

```{r Delineator 24}

#---------------------------------------------------------------------------#

```

```{r Song Acousticness Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

acouplot <-      advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$acou,
                 song_attr_name = 'Acousticness',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1,
                 yupperheader = 1.1)

hist(temp_song_T_transformed$acou,
     breaks = 50,
     main = 'Song Acousticness Histogram',
     xlab = 'Song Acousticness')

rm(acouplot)

```

**Observations**  
- the distribution of song acousticness is skewed  
- this could adversely affect song popularity prediction later on  

```{r Delineator 107}

#---------------------------------------------------------------------------#

```

```{r Song Acousticness Skewness, echo=FALSE, fig.height=3, fig.width=3}

# show the actual counts of the song acousticness
as.data.frame(table(round(temp_song_T_transformed$acou, 1)))

```

**Observations**  
- the song acousticness does not seem to directly affect song popularity  
- to simplify this attribute, a song can have 3 levels of acousticness  
- no acousticness (0), some acousticness (0.5), acousticness (1.0)

```{r Clean Song Acousticness, include = FALSE}

# transform the song acousticness into three bins
temp_song_T_transformed$acou <-
  ifelse(temp_song_T_transformed$acou < 0.25, 0,
    ifelse(temp_song_T_transformed$acou >= 0.25 &
           temp_song_T_transformed$acou < 0.75, 0.5,
      ifelse(temp_song_T_transformed$acou >= 0.75, 1, 0)))

```

```{r Cleaned Song Acousticness Analysis}

# the frequency distribution after transformation
table(temp_song_T_transformed$acou)

```

```{r Delineator 25}

#---------------------------------------------------------------------------#

```

```{r Song Danceability Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

danceplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$danc,
                 song_attr_name = 'Danceability',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1,
                 yupperheader = 1.1)

hist(temp_song_T_transformed$danc,
     breaks = 50,
     main = 'Song Danceability Histogram',
     xlab = 'Song Danceability')

rm(danceplot)

```

**Observations**  
- the danceability attribute has a normal distribution  
- no preprocessing is necessary at this time

```{r Delineator 26}

#---------------------------------------------------------------------------#

```

```{r Song Energy Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

energyplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$ener,
                 song_attr_name = 'Energy',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1,
                 yupperheader = 1.1)

hist(temp_song_T_transformed$ener,
     breaks = 50,
     main = 'Song Energy Histogram',
     xlab = 'Song Energy')

rm(energyplot)

```

**Observations**  
- the danceability attribute has a skewed distribution  
- newer generation songs tend to have higher energy  
- no preprocessing is necessary at this time

```{r Delineator 27}

#---------------------------------------------------------------------------#

```

```{r Song Instrumentalness Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

instrplot <-     advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$inst,
                 song_attr_name = 'Instrumentalness',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1,
                 yupperheader = 1.1)

hist(temp_song_T_transformed$inst,
     breaks = 50,
     main = 'Song Instrumentalness Histogram',
     xlab = 'Song Instrumentalness')

rm(instrplot)

```

**Observations**  
- the song instrumentalness attribute is highly skewed  
- the large majority of songs do not have any instrumentalness  
- the same transformation from song acousticness will be applied

```{r Delineator 108}

#---------------------------------------------------------------------------#

```

```{r Song Instrumentalness Skewness, echo = FALSE}

# show the actual counts of the song acousticness
as.data.frame(table(round(temp_song_T_transformed$inst, 1)))

```

```{r Clean Song Instrumentalness, include = FALSE}

# transform the song instrumentalness into three bins
temp_song_T_transformed$inst <-
  ifelse(temp_song_T_transformed$inst < 0.25, 0,
    ifelse(temp_song_T_transformed$inst >= 0.25 &
           temp_song_T_transformed$inst < 0.75, 0.5,
      ifelse(temp_song_T_transformed$inst >= 0.75, 1, 0)))

```

```{r Cleaned Song Instrumentalness Analysis}

# the frequency distribution after transformation
table(temp_song_T_transformed$inst)

```

```{r Delineator 110}

#---------------------------------------------------------------------------#

```

```{r Song Loudness Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

loudplot <-      advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$loud,
                 song_attr_name = 'Loudness',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1,
                 yupperheader = 1.1)

hist(temp_song_T_transformed$loud,
     breaks = 50,
     main = 'Song Loudness Histogram',
     xlab = 'Song Loudness')

rm(loudplot)

```

**Observations**  
- the loudness attribute has a skewed distribution  
- newer generation songs tend to have higher loudness  
- no preprocessing is necessary at this time

```{r Delineator 29}

#---------------------------------------------------------------------------#

```

```{r Song Speechiness Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

speechplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$spee,
                 song_attr_name = 'Speechiness',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1,
                 yupperheader = 1.1)

hist(temp_song_T_transformed$spee,
     breaks = 50,
     main = 'Song Speechiness Histogram',
     xlab = 'Song Speechiness')

rm(speechplot)

```

```{r Delineator 111}

#---------------------------------------------------------------------------#

```

```{r Song Speechiness Skewness, echo = FALSE}

# show the actual counts of the song Speechiness
as.data.frame(table(round(temp_song_T_transformed$spee, 1)))

```

```{r Clean Song Speechiness, include = FALSE}

# transform the song instrumentalness into three bins
temp_song_T_transformed$spee <-
  ifelse(temp_song_T_transformed$spee < 0.25, 0,
    ifelse(temp_song_T_transformed$spee >= 0.25 &
           temp_song_T_transformed$spee < 0.75, 0.5,
      ifelse(temp_song_T_transformed$spee >= 0.75, 1, 0)))

```

\newpage

```{r Cleaned Song Speechiness Analysis}

# frequency distribution after transformation
table(temp_song_T_transformed$spee)

```

```{r Delineator 33}

#---------------------------------------------------------------------------#

```

```{r Song Tempo Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

tempoplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$temp,
                 song_attr_name = 'Tempo',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1.05,
                 yupperheader = 1.15)

hist(temp_song_T_transformed$temp,
     breaks = 50,
     main = 'Song Tempo Histogram',
     xlab = 'Song Tempo')

rm(tempoplot)

```

```{r Remove Outlier Songs Song Tempo, echo = FALSE}

# create a copy of the dataframe
# this will stay with none of the outliers removed
# might come in handy later for analysis purposes

x.copy2 <- temp_song_T_transformed

# iterate through each song decade
for (x.songdecade in c('1970s', '1980s', '1990s', '2000s', '2010s', 'NA')) {

# store first and third quartiles
x.Q1 <- as.numeric(summary(temp_song_T_transformed[which(temp_song_T_transformed$song.2 == x.songdecade), 27]) [2])
x.Q3 <- as.numeric(summary(temp_song_T_transformed[which(temp_song_T_transformed$song.2 == x.songdecade), 27]) [5])

# store upper and lower outliers
x.LowerOutlier <- x.Q1 - (1.5 * (x.Q3 - x.Q1))
x.UpperOutlier <- x.Q3 + (1.5 * (x.Q3 - x.Q1))

# remove the outliers
temp_song_T_transformed <- temp_song_T_transformed[which(
  
  # is a lower outlier for the current decade
  ! (temp_song_T_transformed$song.2 == x.songdecade &
  temp_song_T_transformed$temp < x.LowerOutlier) &
    
  # or is an upper outlier for the current decade
  ! (temp_song_T_transformed$song.2 == x.songdecade &
  temp_song_T_transformed$temp  > x.UpperOutlier)), ]

}

# print how many were removed
cat('Number of Songs Removed:',
    nrow(x.copy2) - nrow(temp_song_T_transformed))

#reapply the min max transformation
temp_song_T_transformed <- min_max_transform(temp_song_T_transformed, 27)

```

```{r Cleaned Song Tempo Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

tempoplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$temp,
                 song_attr_name = 'Tempo',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1.05,
                 yupperheader = 1.15)

hist(temp_song_T_transformed$temp,
     breaks = 50,
     main = 'Song Tempo Histogram',
     xlab = 'Song Tempo')

rm(tempoplot)

```

```{r Delineator 35}

#---------------------------------------------------------------------------#

```

```{r Song Audio Valence Analysis, echo=FALSE, fig.height=5, fig.width=14}

par(mfrow = c(1, 2))

valenceplot <- advanced_boxplot(
                 song_attr_values = temp_song_T_transformed$vlnc,
                 song_attr_name = 'Audio_Valence',
                 song_year_values = temp_song_T_transformed$song.2,
                 ylimhigh = 1.2,
                 xheadercenter = 3.5,
                 ablinehigh = 1.05,
                 yupperheader = 1.15)

hist(temp_song_T_transformed$vlnc,
     breaks = 50,
     main = 'Song Audio Valence Histogram',
     xlab = 'Song Audio Valence')

rm(valenceplot)

```

```{r Delineator 36}

#---------------------------------------------------------------------------#

```

##### Song Genre and Song Attribute Correlation Matrix

The following pearson correlation matrix shows the correlations between each of
the song genres and song attributes.

The darker the blue color means higher positive correlation (closer to 1), and
the darker the orange color means lower negative correlation (closer to -1).

```{r Join Song Attributes with Known Genres, include = FALSE}

# join in the song attributes and the songs with known genres

song_known_genres <- left_join(
  
  song_known_genres[, c(1, 3:16)],     # left side
  song_data,                           # right side
  by = c('song_name' = 'song_name'))   # by column

```

```{r Genre and Song Attribute Correlation Matrix, echo=FALSE, fig.height=5, fig.width=10}

# create a temporary variable with truncated names to fit on the document
temp_song_known_genres  <- song_known_genres  # first create a copy

# truncate the column names with a max length of 4 characters
colnames(temp_song_known_genres) <-
  substr(colnames(temp_song_known_genres), 1, 4)

# make some manual adjustments to some of the column names
# after being truncated there were some duplicate column names

colnames(temp_song_known_genres) [16] <- 'pplr'   # manually adjust song_popularity name
colnames(temp_song_known_genres) [17] <- 'durn'   # manually adjust song_duration name
colnames(temp_song_known_genres) [25] <- 'mode'   # manually adjust audio mode name
colnames(temp_song_known_genres) [29] <- 'vlnc'   # manually adjust audio valence name

# show a pearson correlation matrix between song attributes and song genres
song_genre_corr <- round(cor(
  temp_song_known_genres[, c(2:15)],
  temp_song_known_genres[, c(16:21, 23:24, 26:27, 29)]), 2)

# scaled the correlation matrix multiply by 2 to show colors better
# removed the legend to avoid that causing any confusion

corrplot(song_genre_corr * 2, method = 'color', outline = TRUE, cl.pos = "n")

# note that a few of the song attributes are ommittied so the table fits in the document
# key, audio mode, and time signature are the columns that are not shown here
# these variables also did not have any interesting correlations to show

# clean up the environment
rm(temp_query, song_genre_corr)

```

**Observations**  
- there are some interesting correlations between song attributes and song genres  
- example 1: alt / rock / metal / punk are less popular than pop / hiphop / rap / r&b  
- example 2: jazz / blues / folk / country tend to have more acousticness  
- example 3: metal / punk tend to have higher energy than the other genres  
- example 4: hiphop / rap have significantly more speechiness than the other genres  
- example 5: other has significantly more audio valence than the other genres  
- this is encouraging because it means it might be possible to effectively infer song genres

\newpage

##### Song Genre Clustering

Based on the song attributes, an unsupervised clustering model will be
implemented to create clusters of songs.

These clusters of songs could represent genres or they could represent
other similarities between the songs.

```{r Song Genre Clustering, echo = FALSE, fig.height=8, fig.width=10}

# kmeans clustering

# compute k-means
create_song_clusters <- function(data, numk) {
    set.seed(175)
    km.song <- kmeans(data, numk, iter.max=30)
    plot <- fviz_cluster(km.song, data = data, geom = c("point"), ellipse.type = "euclid")
    return(plot)
}

songallclusterdata <- data.frame()
for (k in c(5, 7, 9, 11, 13, 15)) {
  songclusterviz <- create_song_clusters(temp_song_T_transformed[, 19:28], k)
  songclusterdata <- songclusterviz$data
  songclusterdata$numK <- paste('K =', k)
  songallclusterdata <- rbind(songallclusterdata, songclusterdata)
}

ggplot() +
  geom_point(
    data = songallclusterdata,
    aes(
      x = x,
      y = y,
      colour = cluster
    )
  ) +
  facet_wrap(
    ~ numK,
    nrow = 3,
    ncol = 2, 
    # labeller = labeller(
    #   genre = genre_plot_titles)
    )

# clean up the environment
rm(songclusterdata, songallclusterdata, songclusterviz, k)

```

**Observations**  
- there must be some level of disctinction based on the song attributes  
- it is possible that these clusters could represent genres  
- for example, genres have their distinct sounds but do overlap

\newpage

##### Association Rules mining

Association rules mining will be performed primarily in the interest of
understanding what correlations exist with song popularity.

```{r Association Rules Mining Data Preparation, include = FALSE}

#create a new dataframe with the data
song_main <- left_join(
  
  new_song_T[
    which(new_song_T$song_name %in% temp_song_T_transformed$song),
    which(colnames(new_song_T) %in% c('song_name', 'song_released'))],
  
  song_data[
  which(! colnames(song_data) %in% c('key', 'liveness', 'audio_mode', 'time_signature'))],
  
  by = c('song_name' = 'song_name'))

##### discretization

#discretize the song attribute columns based on summary statistics
#i.e. five bins - lowoutlier, lowoutlier-Q1, Q1-Q2, Q2-Q3, upperoutlier
#this way each song attribute is scaled and stays relative to itself
#a potential benefit of doing it this way is that outliers are seperated

for (x.column.x in seq(3, 12, 1)) {
  x.tempcolumn.x <- song_main[, x.column.x]
  x.q1.x <- summary(x.tempcolumn.x) [2] %>% as.numeric()
  x.q2.x <- summary(x.tempcolumn.x) [3] %>% as.numeric()
  x.q3.x <- summary(x.tempcolumn.x) [4] %>% as.numeric()
  x.iqr.x <- x.q3.x - x.q1.x
  x.outlow.x <- x.q1.x - (1.5 * x.iqr.x)
  x.outhigh.x <- x.q3.x + (1.5 * x.iqr.x)
  x.newcolumn.x <- c()
  for (x.value.x in seq(1, length(x.tempcolumn.x), 1)) {
    x.newvalue.x <- x.tempcolumn.x [x.value.x]
    x.newvalue.x <-
      ifelse(x.newvalue.x <= x.outlow.x, 'vlow',
        ifelse(x.newvalue.x > x.outlow.x & x.newvalue.x <= x.q1.x, 'low',
          ifelse(x.newvalue.x > x.q1.x & x.newvalue.x <= x.q2.x, 'med',
            ifelse(x.newvalue.x > x.q2.x & x.newvalue.x <= x.q3.x, 'med',
              ifelse(x.newvalue.x > x.q3.x & x.newvalue.x <= x.outhigh.x, 'high',
                ifelse(x.newvalue.x > x.outhigh.x, 'vhigh', 'undef'))))))
  x.newcolumn.x <- c(x.newcolumn.x, x.newvalue.x)
  }
  song_main[, x.column.x] <- x.newcolumn.x
}

#manually rename the columns
colnames(song_main) [3:12] <- c('poplr', 'dur', 'acou', 'dance', 'ener', 'instr', 'loud', 'speech', 'tempo', 'vlnc')

#clean up the environment
rm(x.column.x, x.iqr.x, x.newcolumn.x, x.newvalue.x, x.outhigh.x, x.outlow.x,
   x.q1.x, x.q2.x, x.q3.x, x.tempcolumn.x, x.value.x)

#create a copy
song_main_copy <- song_main

```

```{r Association Rules Mining Generate Rules, include = FALSE}

##### what makes a song popular?

for (x.songdecade.x in unique(song_main$song_released)) {
  
  #filter on the current song decade
  #binarize the song attribute columns
  x.tempsong.x <- data.frame(cbind(
    song_main[song_main$song_released == x.songdecade.x, 3:ncol(song_main)] %>% binarize()))
  
  #change columns to factors for association rules mining
  x.tempsong.x <- x.tempsong.x %>% mutate_all(as.numeric)
  
  # now the dataset can be converted into a transaction object
  x.songtransactions.x <- as(as.matrix(x.tempsong.x), "transactions")
  
  #create association rules for high and very high popularity
  x.songrules.x <-
    apriori(
    x.songtransactions.x,
    parameter = list(
      supp = .02,
      conf = .01,
      minlen = 3),
    appearance = list(
      rhs = c("poplr__vhigh", "poplr__high"),
      default = "lhs")
    )
  
  #store the rules in the global environment
  if (x.songdecade.x == '1970s') {
        x.songrules1970s.x <<- x.songrules.x} else if (
      x.songdecade.x == '1980s') {
        x.songrules1980s.x <<- x.songrules.x} else if (
      x.songdecade.x == '1990s') {
        x.songrules1990s.x <<- x.songrules.x} else if (
      x.songdecade.x == '2000s') {
        x.songrules2000s.x <<- x.songrules.x} else if (
      x.songdecade.x == '2010s') {
        x.songrules2010s.x <<- x.songrules.x} else if (
      x.songdecade.x == 'NA') {
        x.songrulesNA.x <<- x.songrules.x}
  
}
  
```

```{r Delineator 37}

#---------------------------------------------------------------------------#

```

```{r Print Association Rules Function, include = FALSE}

print_association_rules <- function(assocationrulesobject) {
  
  # create a data frame from the rules and sorted versions of the dataframe
  associationrulesdf   <- as(assocationrulesobject, "data.frame")
  sortedbysupport      <- sqldf('select * from associationrulesdf order by support desc')
  sortedbyconfidence   <- sqldf('select * from associationrulesdf order by confidence desc')
  sortedbylift         <- sqldf('select * from associationrulesdf order by lift desc')
  
  # round off the measures to 2 decimal points to save room on the document
  sortedbysupport[, 2:5] <- sortedbysupport[, 2:5] %>% mutate_all(round, 2)
  sortedbyconfidence[, 2:5] <- sortedbyconfidence[, 2:5] %>% mutate_all(round, 2)
  sortedbylift[, 2:5] <- sortedbylift[, 2:5] %>% mutate_all(round, 2)
  
  # get the top 3 results from each sorted version of the rules
  x.sortedbysupport.x <- head(sortedbysupport, 3)
  x.sortedbyconfidence.x <- head(sortedbyconfidence, 3)
  x.sortedbylift.x <- head(sortedbylift, 3)
  
  # print out the results sorted by support
  cat('Top 3 Rules Sorted by Support \n \n')
  for (x.rownumber.x in seq(1, 3, 1)) {
    cat('Rule Number', x.rownumber.x, '\n', 
         x.sortedbysupport.x[x.rownumber.x, 1], '\n')
    print.data.frame(x.sortedbysupport.x[x.rownumber.x, 2:6])
    cat('\n \n')
  } 
  
  # print out the results sorted by confidence
  cat('Top 3 Rules Sorted by Confidence \n \n')
  for (x.rownumber.x in seq(1, 3, 1)) {
    cat('Rule Number', x.rownumber.x, '\n', 
         x.sortedbyconfidence.x[x.rownumber.x, 1], '\n')
    print.data.frame(x.sortedbyconfidence.x[x.rownumber.x, 2:6])
    cat('\n \n')
  } 
  
  # print out the results sorted by lift
  cat('Top 3 Rules Sorted by Lift \n \n')
  for (x.rownumber.x in seq(1, 3, 1)) {
    cat('Rule Number', x.rownumber.x, '\n', 
         x.sortedbylift.x[x.rownumber.x, 1], '\n')
    print.data.frame(x.sortedbylift.x[x.rownumber.x, 2:6])
    cat('\n \n')
  }
  
}
  
```

```{r Associative Rules Mining Results 1970s, echo = FALSE}

# print a header with the decade
cat0('What Makes a Song Popular in the 1970s?  \n \n')
print_association_rules(x.songrules1970s.x)

```

**Observations**  
- songs with medium instrumentalness were the most popular  
- low to medium speechiness is common in popular songs  
- low audio valence was associated with high confidence

```{r Delineator 38}

#---------------------------------------------------------------------------#

```

```{r Associative Rules Mining Results 1980s, echo = FALSE}

# print a header with the decade
cat0('What Makes a Song Popular in the 1980s?  \n \n')
print_association_rules(x.songrules1980s.x)

```

**Observations**  
- some of the same qualities from the 1970s appear in the 1980s  
- medium instrumentalness and low to medium speechiness

```{r Delineator 39}

#---------------------------------------------------------------------------#

```

```{r Associative Rules Mining Results 1990s, echo = FALSE}

# print a header with the decade
cat0('What Makes a Song Popular in the 1990s?  \n \n')
print_association_rules(x.songrules1990s.x)

```

**Observations**  
- there seems to be a shift in which song attributes correlate with song popularity  
- for example, medium acousticness and high audio valence were not seen until now  

```{r Delineator 40}

#---------------------------------------------------------------------------#

```

```{r Associative Rules Mining Results 2000s, echo = FALSE}

# print a header with the decade
cat0('What Makes a Song Popular in the 2000s?  \n \n')
print_association_rules(x.songrules2000s.x)

```

**Observations**  
- not much stands out in this decade compared to the others  
- it seems like more of a blend of the previous decades

```{r Delineator 43}

#---------------------------------------------------------------------------#

```

```{r Associative Rules Mining Results 2010s, echo = FALSE}

# print a header with the decade
cat0('What Makes a Song Popular in the 2010s?  \n \n')
print_association_rules(x.songrules2010s.x)

```

**Observations**  
- low instrumentalness not medium is associated with high popularity  
- high danceability appears and is associated with high confidence + lift

```{r Delineator 45}

#---------------------------------------------------------------------------#

```

```{r Associative Rules Mining Results NA, echo = FALSE}

# print a header with the decade
cat0('What Makes a Song Popular in the unknown years?  \n \n')
print_association_rules(x.songrulesNA.x)

```

**Observations**  
- very high acousticness appears in several rules  
- this is odd because most songs had low acousticness  
- there is definitely something different about these songs

```{r Associative Rules Mining Clean Up Environment, include = FALSE}

rm(x.songrules.x, x.songrules1970s.x, x.songrules1980s.x, x.songrules1990s.x,
   x.songrules2000s.x, x.songrules2010s.x, x.songrulesNA.x, x.songtransactions.x,
   x.tempsong.x, x.songdecade.x, x.Outliers, x.cycle, x.IQR, x.LowerOutlier,
   x.Q1, x.Q3, x.songdecade, x.UpperOutlier)

```

```{r Delineator 46}

#---------------------------------------------------------------------------#

```

\newpage

##### Perform KNN Trials to determine a K Value

An iteration of trials is performed using a KNN model with different K values
on data for which the genre is known.

Since it is possible for a song to have multiple genres, iterate through each
song and each genre one by one, and determine if the song fits in that genre.

First the data is split 80 / 20 training and testing and then the KNN model is
implemented, the accuracy is measured based on the percent correct of the test data.

```{r Trial KNN model Song Genre, include = FALSE}

# this is going to be tricky because songs can have multiple genres
# approach: run a k nearest for every individual song genre
# i.e. run a binary k nearest on all songs for if it is alt, rock, metal, etc.
# at the end, the results will be re-packed into the multi genre format
# i.e. the results from each k nearest will become a binary col in the table

# test out different k values for different genres
# see how much the results are different from eachother
# this will help decide on what value for k to use

# ---------------------------------------------------------------------------- #

# standardize all of the song attributes using a min max transformation
# this is because knn model needs standardized attributes to be effective

# create a seperate version as the standardized version
genre_standardized  <- song_known_genres[
  which(song_known_genres$song_name %in% temp_song_T_transformed$song), 
  which(! colnames(song_known_genres) %in% c('key', 'liveness', 'audio_mode', 'time_signature'))]

# iterate through each of the columns of the song attributes, cols 16 - 29
for (songAttrCol in seq(16, 25, 1)) {
  
  # store the values for the column for the current iteration
  temp_column <- unlist(as.numeric(genre_standardized[, songAttrCol]))
  
  # store the min and max values for the column 
  temp_min    <- min(temp_column); temp_max    <- max(temp_column)
  
  # transform the column based on the min max transformation
  new_column  <- (temp_column - temp_min) / (temp_max - temp_min)
  
  # append the transformed column to the main dataframe
  # if it is the first iteration, create a new variable, otherwise append
  ifelse(songAttrCol == 16,
    genre_temp_standardize <- data.frame('X1' = new_column),
    genre_temp_standardize <- cbind(genre_temp_standardize, new_column))
  
}

  # change the column names of the standardized dataframe accordingly
  colnames(genre_temp_standardize) <- colnames(genre_standardized) [16:25]
  
  # replace the values in the main dataframe with the standardized values
  genre_standardized[, 16:25] <- genre_temp_standardize
  
# ---------------------------------------------------------------------------- #

# start out by creating an empty dataframe variable
# this will be appended to with the results from each trial
genreKNNres <- data.frame()

# iterate through k values 3-11 by 2 for each genre
# using odd number k values to prevent ties
for (genreColNumber in seq(2, 15, 1)) {
  for (genreKvalue in seq(3, 21, 2)) {
    
    set.seed(520) # set seed then partition the data into testing and training data
    genre_partition   <- createDataPartition(genre_standardized[, genreColNumber], p = 0.8, list = FALSE)
    genre_training    <- genre_standardized[genre_partition, c(genreColNumber, 16:25)]
    genre_testing     <- genre_standardized[-genre_partition, c(genreColNumber, 16:25)]
    
    # process the knn model for the current iteration
    genreKNNpred <- knn(train = genre_training[, 2:ncol(genre_training)],  # training data (minus the label)
                    test = genre_testing[, 2:ncol(genre_testing)],         # testing data (minus the label)
                    cl = genre_training[, 1],                              # labels from the training data
                    k = genreKvalue)                                       # k value for the current iteration
    
    # store the accuracy percentage from the model
    TempKNNRes <- round(mean(genreKNNpred == genre_testing[, 1]) * 100, 2)
    
    # put the accuracy percentage in a temporary dataframe
    TempKNNTable <- data.frame(cbind(
      'Genre'      = colnames(genre_standardized) [genreColNumber],
      'NumK'       = genreKvalue,
      'Accuracy'   = TempKNNRes))
    
    # append the temporary dataframe to the main dataframe
    genreKNNres <- rbind(genreKNNres, TempKNNTable)
    
    # print progress updates after each model is completed
    if (genreKvalue == 21) {
    cat('model complete | Genre:',
        colnames(genre_standardized) [genreColNumber], '\n')}
    
  }
}

# change the numK and accuracy column to numeric data type
genreKNNres$NumK       <- as.numeric(genreKNNres$NumK)       # NumK to numeric        
genreKNNres$Accuracy   <- as.numeric(genreKNNres$Accuracy)   # Accuracy to numeric

# restructure the table to be more readable for printing out

genreKNNdisplay <- dcast(genreKNNres, Genre ~ NumK, mean)                       # cast the k values to columns
colnames(genreKNNdisplay) [-1] <- paste0('K', colnames(genreKNNdisplay)) [-1]   # add K to the columns
genreKNNdisplay$Avg <- rowMeans(genreKNNdisplay[, -1])                          # add the row averages

```

```{r Show KNN Results, echo = FALSE}

# overall KNN model accuracy by number of k nearest neighbors
cat('Overall Accuracy by Number of K Nearest Neighbors \n \n')
round(colMeans(genreKNNdisplay[, -1]), 0)

# KNN model accuracy by number of k nearest neighbors and genre
kable(genreKNNdisplay, digits = 0)

```

**Observations**  
- able to effectively determine song genre with 88% on avg accuracy  
- K values of 9-21 all produced very similar accuracies  
- moving forward with a conservative middle of the pack k value of 11  
- do not want to risk selecting too many k nearest neighbors  

\newpage

##### Infer Missing Song Genres with KNN model K = 11

All of the missing song genres are inferred using a KNN model with a K value of
11. The training data used is all of the data with known genres.

```{r Infer Missing Song Genres with KNN, include = FALSE}

# ---------------------------------------------------------------------------- #

# make sure that any songs that were removed are removed here
song_unknown_genres <- song_unknown_genres[
  which(song_unknown_genres$song_name %in% temp_song_T_transformed$song),
  which(! colnames(song_unknown_genres) %in% c('key', 'liveness', 'audio_mode', 'time_signature'))]
song_known_genres <- song_known_genres[
  which(song_known_genres$song_name %in% temp_song_T_transformed$song),
  which(! colnames(song_known_genres) %in% c('key', 'liveness', 'audio_mode', 'time_signature'))]

# add in all of the genre columns to the song_unkown_genres dataframe
# use the reconcile tables function that was introduced earlier
song_unknown_genres <- reconcile_tables(song_unknown_genres, song_known_genres[, 2:15]) [, -2]

# join in all of the song attributes from the song_data table
song_unknown_genres <- left_join(song_unknown_genres, song_data, by = c('song_name' = 'song_name'))
song_unknown_genres <- song_unknown_genres[,
  which(! colnames(song_unknown_genres) %in% c('key', 'liveness', 'audio_mode', 'time_signature'))]

# trim down the dataframe back to the correct rows
# some NA rows were added incidentally due to the reconcile tables function
song_unknown_genres <- song_unknown_genres[-which(is.na(song_unknown_genres$song_name)), ]

# standardize all of the song attributes using a min max transformation
# this is because knn model needs standardized attributes to be effective

# create a seperate version as the standardized version
genre_inference  <- song_unknown_genres   # first create a copy

# iterate through each of the columns of the song attributes, cols 16 - 29
genre_inference <- min_max_transform(genre_inference, seq(16, 25, 1))
  
# ---------------------------------------------------------------------------- #

# iterate through each genre column and make an inference
for (genreColNumber in seq(2, 15, 1)) {
    
    # the training data are the known genres
    # the testing data are the unknown genres
    genre_training    <- genre_standardized[, c(genreColNumber, 16:25)]
    genre_testing     <- genre_inference[, c(genreColNumber, 16:25)]
    
    # process the knn model for the current iteration
    genreKNNpred <- knn(train = genre_training[, 2:ncol(genre_training)],  # training data (minus the label)
                    test = genre_testing[, 2:ncol(genre_testing)],         # testing data (minus the label)
                    cl = genre_training[, 1],                              # labels from the training data
                    k = 11)                                                # k value selected as 11 from trials
    
    # if its the first iteration, create new variable, otherwise append
    # this variable will contain all of the inferred genre columns
    ifelse(genreColNumber == 2,
      genreKNNInference <- data.frame('X1' = genreKNNpred),
      genreKNNInference <- cbind(genreKNNInference, genreKNNpred))
    
    # print progress updates after each model is completed
    cat('model complete | Genre:',
        colnames(genre_standardized) [genreColNumber], '\n')
    
}
  
# change the column names of the genre inference dataframe accordingly
colnames(genreKNNInference) <- colnames(song_unknown_genres) [2:15]

# insert the inferred genres into the song_unkown_genres dataframe
song_unknown_genres[, 2:15] <- genreKNNInference

# finally, create the final dataframe with all of the genres for each song
song_genre <- rbind(song_known_genres[, 1:15], song_unknown_genres[, 1:15])

```

```{r Genre Inference Results, echo = FALSE}

temp_inference_results <- 
  
    sqldf('
      select
        sum(alt) as alt,
        sum(rock) as rock,
        sum(metal) as metal,
        sum(punk) as punk,
        sum(pop) as pop,
        sum(hiphop) as hiphop,
        sum(rb) as rb,
        sum(rap) as rap,
        sum(jazz) as jazz,
        sum(blues) as blues,
        sum(folk) as folk,
        sum(elect) as elect,
        sum(other) as other
      from
        song_unknown_genres'
    )

temp_inference_results

# clean up the environment

rm(genre_inference, genre_partition, genre_standardized, genre_temp_standardize,
   genre_testing, genre_training, genreKNNdisplay, genreKNNInference, genreKNNres,
   song_known_genres, song_unknown_genres, song_temp_genres,
   TempKNNTable, genreColNumber, genreKNNpred, genreKvalue, new_column,
   songAttrCol, temp_column, temp_max, temp_min, TempKNNRes, temp_inference_results)

```

##### Perform KNN Trials to determine a K Value for Song Decade

An iteration of trials is performed using a KNN model with different K values
on data for which the song decade is known.

First the data is split 80 / 20 training and testing and then the KNN model is
implemented, the accuracy is measured based on the percent correct of the test data.

```{r Trial KNN model Song Decade, include = FALSE}

# test out different k values for different genres
# see how much the results are different from eachother
# this will help decide on what value for k to use

# create a seperate dataframe with the known years
song_known_years <- temp_song_T_transformed[
  -which(temp_song_T_transformed$song.2 == 'NA')
  , c(18, 3:17, 19:28)]

# create a seperate dataframe with the unknown years
song_unknown_years <- temp_song_T_transformed[
  which(temp_song_T_transformed$song.2 == 'NA')
  , c(18, 3:17, 19:28)]

# ---------------------------------------------------------------------------- #

# start out by creating an empty dataframe variable
# this will be appended to with the results from each trial
x.yearKNNres.x <- data.frame()

# iterate through k values 3-11 by 2 for each year
# using odd number k values to prevent ties
for (x.yearKvalue.x in seq(3, 21, 2)) {
    
    set.seed(520) # set seed then partition the data into testing and training data
    x.year_partition.x   <- createDataPartition(song_known_years[, 1], p = 0.8, list = FALSE)
    x.year_training.x    <- song_known_years[x.year_partition.x, ]
    x.year_testing.x     <- song_known_years[-x.year_partition.x, ]
    
    # process the knn model for the current iteration
    x.yearKNNpred.x <- knn(train = x.year_training.x[, 2:ncol(x.year_training.x)],  # training data (minus the label)
                           test = x.year_testing.x[, 2:ncol(x.year_testing.x)],     # testing data (minus the label)
                           cl = x.year_training.x[, 1],                             # labels from the training data
                           k = x.yearKvalue.x)                                      # k value for the current iteration
    
    # store the accuracy percentage from the model
    x.TempKNNRes.x <- round(mean(x.yearKNNpred.x == x.year_testing.x[, 1]) * 100, 2)
    
    # put the accuracy percentage in a temporary dataframe
    x.TempKNNTable.x <- data.frame(cbind(
      'Year'       = colnames(song_known_years) [1],
      'NumK'       = x.yearKvalue.x,
      'Accuracy'   = x.TempKNNRes.x))
    
    # append the temporary dataframe to the main dataframe
    x.yearKNNres.x <- rbind(x.yearKNNres.x, x.TempKNNTable.x)
    
    # print progress updates after each model is completed
    if (x.yearKvalue.x == 21) {
    cat('model complete | Year:',
        colnames(song_known_years) [1], '\n')}
    
  }

```

```{r Show KNN Results Song Decade, echo = FALSE}

# overall KNN model accuracy by number of k nearest neighbors
cat('Overall Accuracy by Number of K Nearest Neighbors \n \n')
kable(x.yearKNNres.x[, -1])

# clean up the environment
rm(x.year_partition.x, x.year_testing.x, x.year_training.x, x.yearKNNres.x,
   x.yearKNNpred.x, x.yearKvalue.x, x.TempKNNTable.x, x.TempKNNRes.x,
   song_main_copy, song_known_years, song_unknown_years, temp_song_known_genres,
   temp_song_T)

```

**Observations**  
- the accuracy for inferring the song genre is very good  
- the accuracy for inferring the song decade is not good  
- it might be better to model with the missing song decades

\newpage

# Data Modeling

Now that the data has been explored and cleaned, several models will be implemented
to predict the song popularity and the results will be compared.

The song popularity will be predicted via both classification and regression.
Song popularity will need to be binned prior to classification models.

The models that will be implemented include, decision tree, naive bayes,
k nearest neighbors, support vector machine, and neural net.

The objective is to find a model that is suitable for accurately predicting the
song popularity.

```{r Delineator 115}

#---------------------------------------------------------------------------#

```

##### Sampling

Due to the distribution of song genres being skewed, a sample will be taken
with an even amount of songs from each genre.

```{r Sampling, include = FALSE}

# a sample of n = 100 is taken from each genre
# i.e. 100 songs alt, 100 songs rock, ... 100 songs rap, etc.
# since there is overlap it wont add up exactly to 1400

# add a column for the row indexes
row.names(song_genre) <- NULL
song_genre_sample <- song_genre
song_genre_sample$rownum <- row.names(song_genre_sample)
song_genre_sample[, -1] <- song_genre_sample[, -1] %>% mutate_all(as.numeric)
song_genre_sample[complete.cases(song_genre_sample), ]

# create an empty list that will be used to store row indexes
x.song_genre_sample.x <- data.frame()

# sample alternative - first 100 songs
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$alt == 1), which(colnames(song_genre_sample) == 'rownum')], 100, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[x.sample.x, ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$alt == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)])

# sample rock - 35 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$rock == 1), which(colnames(song_genre_sample) == 'rownum')], 35, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$rock == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample metal - 91 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$metal == 1), which(colnames(song_genre_sample) == 'rownum')], 91, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$metal == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample punk - 85 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$punk == 1), which(colnames(song_genre_sample) == 'rownum')], 85, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$punk == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample pop - 17 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$pop == 1), which(colnames(song_genre_sample) == 'rownum')], 17, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$pop == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample hiphop - 73 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$hiphop == 1), which(colnames(song_genre_sample) == 'rownum')], 73, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$hiphop == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample rb - 75 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$rb == 1), which(colnames(song_genre_sample) == 'rownum')], 75, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$rb == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample rap - 54 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$rap == 1), which(colnames(song_genre_sample) == 'rownum')], 54, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$rap == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample jazz - 90 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$jazz == 1), which(colnames(song_genre_sample) == 'rownum')], 90, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$jazz == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample blues - 82 songs needed
# only 37 songs available
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$blues == 1), which(colnames(song_genre_sample) == 'rownum')], 37, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$blues == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample folk - 66 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$folk == 1), which(colnames(song_genre_sample) == 'rownum')], 66, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$folk == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample country - 65 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$country == 1), which(colnames(song_genre_sample) == 'rownum')], 65, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$country == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample electronic - 52 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$elect == 1), which(colnames(song_genre_sample) == 'rownum')], 52, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$elect == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# sample other - 100 songs needed
set.seed(15)
x.sample.x <- sample(song_genre_sample[which(song_genre_sample$other == 1), which(colnames(song_genre_sample) == 'rownum')], 100, replace = FALSE)
x.song_genre_sample.x <- rbind(x.song_genre_sample.x, song_genre_sample[which(song_genre_sample$rownum %in% x.sample.x), ])
song_genre_sample <- song_genre_sample[-which(song_genre_sample$other == 1), ]
colSums(x.song_genre_sample.x[, -c(1, 16)], na.rm = TRUE)

# store the songs from the sampling
song_sampled <- x.song_genre_sample.x[, 1]

# clean up the environment
rm(x.sample.x)
  
```

```{r Prepare Several Datasets for Modeling, include = FALSE}

# dataset 1 - discrete song popularity and discrete song attributes
data1 <-    (data.frame(cbind('song_name' = song_sampled)) %>%
            left_join(song_main[, c(1, 3)], by = c('song_name' = 'song_name')) %>%
            left_join(song_genre[, -16], by = c('song_name' = 'song_name')) %>%
            left_join(song_main[, -c(2, 3)], by = c('song_name' = 'song_name'))) [, -1]

            # split into testing and training
            set.seed(650)
            partition1 <- createDataPartition(data1[, 1], p = 0.80, list = FALSE)
            train1 <- data1[partition1, ]
            test1 <- data1[-partition1, ]

# dataset 2 - discrete song popularity and continuous song attributes
data2 <-    (data.frame(cbind('song_name' = song_sampled)) %>%
            left_join(song_main[, c(1, 3)], by = c('song_name' = 'song_name')) %>%
            left_join(song_genre[, -16], by = c('song_name' = 'song_name')) %>%
            left_join(temp_song_T_transformed[, c(1, 20:28)], by = c('song_name' = 'song'))) [, -1]

            # split into testing and training
            set.seed(650)
            partition2 <- createDataPartition(data2[, 1], p = 0.80, list = FALSE)
            train2 <- data2[partition2, ]
            test2 <- data2[-partition2, ]

# dataset 3 - continous song popularity and discrete song attributes
data3 <-    (data.frame(cbind('song_name' = song_sampled)) %>%
            left_join(temp_song_T_transformed[, c(1, 19)], by = c('song_name' = 'song')) %>%
            left_join(song_genre[, -16], by = c('song_name' = 'song_name')) %>%
            left_join(song_main[, -c(2, 3)], by = c('song_name' = 'song_name'))) [, -1]

            # split into testing and training
            set.seed(650)
            partition3 <- createDataPartition(data3[, 1], p = 0.80, list = FALSE)
            train3 <- data3[partition3, ]
            test3 <- data3[-partition3, ]

# dataset 4 - continous song popularity and continous song attributes
data4 <-    (data.frame(cbind('song_name' = song_sampled)) %>%
            left_join(temp_song_T_transformed[, c(1, 19)], by = c('song_name' = 'song')) %>%
            left_join(song_genre[, -16], by = c('song_name' = 'song_name')) %>%
            left_join(temp_song_T_transformed[, c(1, 20:28)], by = c('song_name' = 'song'))) [, -1]

            # split into testing and training
            set.seed(650)
            partition4 <- createDataPartition(data4[, 1], p = 0.80, list = FALSE)
            train4 <- data4[partition4, ]
            test4 <- data4[-partition4, ]

```

```{r Show the Sampled Data, echo = FALSE, fig.height=5, fig.width=10}

# join in the song generation with the song genres dataframe

x.song_genre_sample.x <- left_join(x.song_genre_sample.x[, -16], new_song_T[, c(1, 2)], by = c('song_name' = 'song_name'))

# create a temporary variable for the distribution of genres by song generation

temp_analysis <- 
  
sqldf('
select
  song_released,
  sum(alt) as alt,
  sum(rock) as rock,
  sum(metal) as metal,
  sum(punk) as punk,
  sum(pop) as pop,
  sum(hiphop) as hiphop,
  sum(rb) as rb,
  sum(rap) as rap,
  sum(jazz) as jazz,
  sum(blues) as blues,
  sum(folk) as folk,
  sum(elect) as elect,
  sum(other) as other
from `x.song_genre_sample.x`
group by song_released')

# melt the data frame for plotting a stacked bar chart

temp_analysis <- melt(temp_analysis, variable.name = 'Genre', value.name = 'Count')

# what is the distribution of the song genres - by generation?
# keep in mind a song can have more than one genre

ggplot(temp_analysis, aes(reorder(Genre, -Count), Count)) +
  geom_col(position = 'stack', aes(fill = song_released), colour = 'black') +
  ggtitle('\n Distribution of Genres by Song Generation for Sampled Data \n ') +
  ylab('\n Count \n') + xlab('\n Song Genre \n') +
  scale_fill_grey(start = 0, end = .9) +

  # some more customizations to the plot
  theme(
        
        # change the font size of x axis and y axis labels to 11
        axis.text.x = element_text(size = 9), 
        axis.text.y = element_text(size = 9),
        
        # change the x axis and y axis titles to size 12, black, and boldface
        axis.title.x = element_text(size = 10, color = "black", face = "bold"),
        axis.title.y = element_text(size = 10, color = "black", face = "bold"),
        
        # get rid of the default background
        panel.background = element_rect(fill = NA)
        
        )

```

**Observations**  
- did not get full 100 songs from song genre = blues due to lack of data  
- this makes sense becuase blues is not a very popular genre  
- moving forward regardless of the reduced sample size for the blues genre

```{r Delineator 116}

#---------------------------------------------------------------------------#

```

\newpage

```{r Delineator 120}

#---------------------------------------------------------------------------#

```

##### Decision Tree Model 1

Decision tree classification model using the rpart package.

```{r Decision Tree Model 1, echo = FALSE}

# fit the decision tree model using the rpart package

z.fit <- rpart(poplr ~ ., data = train1, method = "class")
z.pred <- predict(z.fit, test1[which(test1$vlnc != 'vlow'), -1], type = "class")
z.res <- data.frame('actual' = test1[which(test1$vlnc != 'vlow'), 1], 'predicted' = z.pred)
z.accuracy <- mean(z.res$actual == z.res$predicted, na.rm = TRUE)

# show the results from the decision tree model
cat('Decision Tree Model 1 - Accuracy =', round(z.accuracy * 100, 2), '%', '\n \n')

fancyRpartPlot(z.fit)
confusionMatrix(table(z.res$actual, z.res$predicted))$table

```

<p>&nbsp;</p>

**Observations**  
- the model did not perform well  
- possible causes - noise, discretization method, attributes, model

```{r Delineator 117}

#---------------------------------------------------------------------------#

```

\newpage

```{r Delineator 125}

#---------------------------------------------------------------------------#

```

##### Decision Tree Model 2

Decision tree classification model using the C5.0 package.

```{r Decision Tree Model 2, echo = FALSE}

# model requires a factor outcome

train1[, 1] <- as.factor(train1[, 1])
train2[, 1] <- as.factor(train2[, 1])

# fit the decision tree model using the C50 package

z.fit <- C5.0(poplr ~ ., data = train2)
z.pred <- predict(z.fit, test2[, -1])
z.res <- data.frame('actual' = test1$poplr, 'predicted' = z.pred)
z.accuracy <- mean(z.res$actual == z.res$predicted, na.rm = TRUE)

# show the results from the decision tree model

overall <- as.data.frame(confusionMatrix(table(z.res$actual, z.res$predicted))$overall)
colnames(overall) <- 'Metric'
kable(overall)

```

<p>&nbsp;</p>

```{r Decision Tree 2 Confusion Matrix, echo = FALSE}

cat('\n Decision Tree 2 Confusion Matrix \n')

kable(confusionMatrix(table(z.res$actual, z.res$predicted))$table)

```

<p>&nbsp;</p>

```{r Decision Tree 2 Precision and Recall, echo = FALSE}

cat('\n Decision Tree 2 Precision and Recall \n')

(confusionMatrix(table(z.res$actual, z.res$predicted))$byClass) [, c(5, 6)]

```

<p>&nbsp;</p>

**Observations**  
- the model did not perform well  
- decision tree is probably not effecive for this prediction task

```{r Delineator 135}

#---------------------------------------------------------------------------#

```

```{r Delineator 140}

#---------------------------------------------------------------------------#

```

##### Naive Bayes Model 1

Naive Bayes classification model using the e1071 package.

```{r Naive Bayes Model 1, echo = FALSE}

# fit the naive bayes model

z.fit <- naiveBayes(poplr ~ ., data = train1) # accuracy = 23.4%
z.pred <- predict(z.fit, test1[, -1])
z.res <- data.frame('actual' = test1$poplr, 'predicted' = z.pred)
z.accuracy <- mean(z.res$actual == z.res$predicted, na.rm = TRUE)

# show the results from the naive bayes model

overall <- as.data.frame(confusionMatrix(table(z.res$actual, z.res$predicted))$overall)
colnames(overall) <- 'Metric'
kable(overall)

```

<p>&nbsp;</p>

**Observations**  
- results about the same as decision tree  
- possible causes - noise, discretization method, attributes, model

```{r Delineator 300}

#---------------------------------------------------------------------------#

```

##### Naive Bayes Model 2

Same as previous but with various discretization methods.

```{r Naive Bayes Model 2, include = FALSE}

z.nbresults <- c()

# iterate from 5 to 20 bins
# this is the number of bins that the song attributes will be
# discretized into each time
for (z.breaks in seq(5, 20, 1)) {
  for (z.method in c("interval", "frequency")) {

# make a fresh copy of the data
data4new <- data4

# discretize the song attributes by the corresponding
# method and number of breaks
data4new[, c(1, 16:24)] <- data4new[, c(1, 16:24)] %>%
  mutate_all(discretize, method = z.method, breaks = z.breaks)

# split into testing and training data
set.seed(470)
partition4new <- createDataPartition(data4new[, 1], p = 0.80, list = FALSE)
train4new <- data4new[partition4new, ]
test4new <- data4new[-partition4new, ]

# fit the naive bayes model

z.fit <- naiveBayes(pop.1 ~ ., data = train4new)
z.pred <- predict(z.fit, test4new[, -1])
z.res <- data.frame('actual' = test4new$pop.1, 'predicted' = z.pred)
z.accuracy <- mean(z.res$actual == z.res$predicted, na.rm = TRUE)

# append the results to the main list
z.nbresults <- c(z.nbresults, z.accuracy)

  }
}

```

<p>&nbsp;</p>

```{r Naive Bayes Model 2 Results, echo = FALSE}

cat('Results from Different Number of Bins \n \n',
    paste0(round(z.nbresults * 100, 2), '%  '))

```

<p>&nbsp;</p>

**Observations**  
- changing the bins did not improve the result  
- naive bayes does not seem to be effective either  

```{r Delineator 301}

#---------------------------------------------------------------------------#

```

\newpage

##### K Nearest Neighbors

K nearest neighbors model with a K value of 21.

```{r K Nearest Neighbors, echo = FALSE}

# change the song genres to numeric

train2[, 1] <- as.factor(train2[, 1])
test2[, 1] <- as.factor(test2[, 1])
train2[, 2:15] <- train2[, 2:15] %>% mutate_all(as.numeric)
test2[, 2:15] <- test2[, 2:15] %>% mutate_all(as.numeric)

# process the knn model for the current iteration
z.knn <- knn(train = train2[, 2:ncol(train2)],                      # training data (minus the label)
                       test = test2[, 2:ncol(test2)],               # testing data (minus the label)
                       cl = train2[, 1],                            # labels from the training data
                       k = 21)                                      # k value for the current iteration

# store the accuracy percentage from the model
z.res <- data.frame('actual' = test2$poplr, 'predicted' = z.knn)
z.accuracy <- mean(z.res$actual == z.res$predicted, na.rm = TRUE)

# show the results from the random forest model

overall <- as.data.frame(confusionMatrix(table(z.res$actual, z.res$predicted))$overall)
colnames(overall) <- 'Metric'
kable(overall)

```

<p>&nbsp;</p>

```{r KNN Confusion Matrix, echo = FALSE}

cat('\n KNN Confusion Matrix \n')

kable(confusionMatrix(table(z.res$actual, z.res$predicted))$table)

```

<p>&nbsp;</p>

```{r KNN Precision and Recall, echo = FALSE}

cat('\n KNN Precision and Recall \n')

(confusionMatrix(table(z.res$actual, z.res$predicted))$byClass) [, c(5, 6)]

```

<p>&nbsp;</p>

**Observations**  
- the knn model did not perform well  
- none of the classification models have produced good results  

\newpage

##### SVM Model

A regression SVM model will be implemented to predict song popularity.

```{r SVM Model, echo=FALSE}

# change the song genres to numeric

train4[, 2:15] <- train4[, 2:15] %>% mutate_all(as.numeric)
test4[, 2:15] <- test4[, 2:15] %>% mutate_all(as.numeric)

# fit the SVM model

z.fit <- svm(pop.1 ~ ., data = train4, kernel = "polynomial", cost = 1)
z.pred <- predict(z.fit, test4[, -1])
z.res <- data.frame('actual' = test4$pop.1, 'predicted' = z.pred)
z.res$error <- z.res$predicted - z.res$actual
z.res$abserror <- abs(z.res$predicted - z.res$actual)
z.accuracy <- 1 - (sum(z.res$abserror) / sum(z.res$actual))
z.avgerror <- mean(z.res$error)
z.avgabserror <- mean(z.res$abserror)
z.biaspct <- sum(z.res$error) / sum(z.res$actual)

# show the results from the SVM model

cat(' Accuracy =', paste0(round(z.accuracy * 100, 2), '%'), '\n',
    'Avg Abs Error =', round(z.avgabserror * 100, 2), '\n',
    'Prediction Bias =', paste0(round(z.biaspct * 100, 2), '%'), '\n')

```

```{r SVM Model Plot, echo=FALSE, fig.height=3, fig.width=3}

plot(z.res[, c(1, 2)],
     xlab('Actual Song Popularity'),
     ylab('Predicted Song Popularity'))

```

```{r SVM Model Appendix}

# kernel = polynomial, cost = 1, degree = 3 -----> accuracy 66%
# kernel = polynomial, cost = 5, degree = 3 -----> accuracy 59%
# kernel = polynomial, cost = 10, degree = 3 -----> accuracy 55%
# kernel = polynomial, cost = 1, degree = 4 -----> accuracy 66%
# kernel = polynomial, cost = 1, degree = 5 -----> accuracy 64%
# kernel = polynomial, cost = 1, degree = 6 -----> accuracy 62%
# type = nu-regression, kernel = linear, cost = 1 -----> accuracy 64%
# type = nu-regression, kernel = linear, cost = 25 -----> accuracy 62%
# type = nu-regression, kernel = linear, cost = 50 -----> accuracy 66%
# type = nu-regression, kernel = linear, ost = 500 -----> accuracy 66%
# type = nu-regression, kernel = polynomial, cost = 5 -----> accuracy 60%

```

<p>&nbsp;</p>

**Observations**  
-  the polynomial kernel with a cost of 1 and degree of 3 gave best results  
-  accuracy was 66% when predicting a continuous song popularity  
-  the SVM model has better results than the decision tree and naive bayes

\newpage

##### Random Forest Model

A regression random forest model will be implemented to predict song popularity.

```{r Random Forest Model, echo = FALSE}

# fit the random forest model

# z.fit <- randomForest(pop.1 ~ ., data = train4, ntree = 10) # accuracy 65.8%
# z.fit <- randomForest(pop.1 ~ ., data = train4, ntree = 50) # accuracy 67%
# z.fit <- randomForest(pop.1 ~ ., data = train4, ntree = 100) # accuracy 67.2%
z.fit <- randomForest(pop.1 ~ ., data = train4, ntree = 100) # accuracy 67.2%
z.pred <- predict(z.fit, test4[, -1], type = c("response"))
z.res <- data.frame('actual' = test4$pop.1, 'predicted' = z.pred)
z.res$error <- z.res$predicted - z.res$actual
z.res$abserror <- abs(z.res$predicted - z.res$actual)
z.accuracy <- 1 - (sum(z.res$abserror) / sum(z.res$actual))
z.avgerror <- mean(z.res$error)
z.avgabserror <- mean(z.res$abserror)
z.biaspct <- sum(z.res$error) / sum(z.res$actual)

# show the results from the random forest model

cat(' Accuracy =', paste0(round(z.accuracy * 100, 2), '%'), '\n',
    'Avg Abs Error =', round(z.avgabserror * 100, 2), '\n',
    'Prediction Bias =', paste0(round(z.biaspct * 100, 2), '%'), '\n')

```

```{r Random Forest Model Plot, echo=FALSE, fig.height=3, fig.width=3}

plot(z.res[, c(1, 2)],
     xlab('Actual Song Popularity'),
     ylab('Predicted Song Popularity'))

```

**Observations**  
- the random forest with 100 trees gave best results  
- after 100 trees the accuracy begins to flatten  
- the random forest was a little better than the SVM  

\newpage

```{r Neural Network Model 1, echo = FALSE}

# # assemble the dataframe for the svm model
# 
# z.data <- (data.frame(cbind('song_name' = song_sampled)) %>%
#           left_join(song_main[, c(1, 3)], by = c('song_name' = 'song_name')) %>%
#           left_join(song_genre, by = c('song_name' = 'song_name')) %>%
#           left_join(temp_song_T_transformed[, c(1, 20:28)], by = c('song_name' = 'song'))) [, -1]
# 
# # change the song genres to numeric
# 
# z.data[, 1] <- as.factor(z.data[, 1])
# z.data[, 2:15] <- z.data[, 2:15] %>% mutate_all(as.numeric)
# z.data <- z.data[complete.cases(z.data), ]
# z.data <- data.frame(cbind(z.data %>% select(poplr) %>% binarize(), z.data[, 2:ncol(z.data)]))
# 
# # split the dataframe into training and testing
# 
# set.seed(650)
# z.partition <- createDataPartition(z.data[, 1], p = 0.80, list = FALSE)
# z.train <- z.data[z.partition, ]
# z.test <- z.data[-z.partition, ]
# 
# # Set up formula
# n <- names(z.train)
# f <- as.formula(paste("poplr__high +
#                       poplr__low +
#                       poplr__med +
#                       poplr__vhigh +
#                       poplr__vlow ~",
#                       paste(n[!n %in% c("poplr__high",
#                                         "poplr__low",
#                                         "poplr__med",
#                                         "poplr__vhigh",
#                                         "poplr__vlow")],
#                             collapse = " + ")))
# 
# # fit the neural network model
# 
# z.fit <- neuralnet(f,
#                    data = z.train,
#                    hidden = 3,
#                    act.fct = "logistic",
#                    linear.output = FALSE)
# 
# z.pred <- predict(z.fit, z.test[, -1])
# 
# # Accuracy (training set)
# original_values <- max.col(z.test[, 1:5])
# pr.nn_2 <- max.col(z.pred)
# mean(pr.nn_2 == original_values)

```

##### Neural Net

A regression neural net will be implemented to predict song popularity.

```{r Neural Network Model, echo = FALSE}

# Set up formula
n <- names(train4)
f <- as.formula(paste("pop.1 ~",
                      paste(n[!n %in% c("pop.1")],
                            collapse = " + ")))

# fit the neural network model

# z.fit <- neuralnet(f, data = train4, linear.output = TRUE) # accuracy = 66.9%
z.fit <- neuralnet(f, data = train4, hidden = 2, linear.output = TRUE) # accuracy = 61.4%
z.pred <- predict(z.fit, test4[, -1])
z.res <- data.frame('actual' = test4$pop.1, 'predicted' = z.pred)
z.res$error <- z.res$predicted - z.res$actual
z.res$abserror <- abs(z.res$predicted - z.res$actual)
z.accuracy <- 1 - (sum(z.res$abserror) / sum(z.res$actual))
z.avgerror <- mean(z.res$error)
z.avgabserror <- mean(z.res$abserror)
z.biaspct <- sum(z.res$error) / sum(z.res$actual)

# show the results from the neural net model

cat(' Accuracy =', paste0(round(z.accuracy * 100, 2), '%'), '\n',
    'Avg Abs Error =', round(z.avgabserror * 100, 2), '\n',
    'Prediction Bias =', paste0(round(z.biaspct * 100, 2), '%'), '\n')

```

```{r Neural Net Model Plot 2, echo=FALSE, fig.height=3, fig.width=3}

plot(z.res[, c(1, 2)],
     xlab('Actual Song Popularity'),
     ylab('Predicted Song Popularity'))

```

```{r Neural Net Model Plot 1, echo=FALSE, fig.height=3, fig.width=3}

plot(z.fit, rep = "best")

```

<p>&nbsp;</p>

**Observations**  
- the neural net produced similar results to SVM and random forest  
- adding additional hidden layers did not increase the accuracy  

\newpage

##### Linear Model

A regression linear model will be implemented to predict song popularity.

```{r Linear Regression Model, echo = FALSE}

z.fit <- lm(pop.1 ~ ., data = train4)
z.pred <- predict(z.fit, test4[, -1])
z.res <- data.frame('actual' = test4$pop.1, 'predicted' = z.pred)
z.res$error <- z.res$predicted - z.res$actual
z.res$abserror <- abs(z.res$predicted - z.res$actual)
z.accuracy <- 1 - (sum(z.res$abserror) / sum(z.res$actual))
z.avgerror <- mean(z.res$error)
z.avgabserror <- mean(z.res$abserror)
z.biaspct <- sum(z.res$error) / sum(z.res$actual)

# show the results from the SVM model

cat(' Accuracy =', paste0(round(z.accuracy * 100, 2), '%'), '\n',
    'Avg Abs Error =', round(z.avgabserror * 100, 2), '\n',
    'Prediction Bias =', paste0(round(z.biaspct * 100, 2), '%'), '\n')

```

```{r Store Linear Model Plot, echo=FALSE, fig.height=7, fig.width=7}

par(mfrow = c(2, 2))
plot(z.fit)

```

**Observations**  
- the linear model performed similar to the other models  
- the regression models had similar performance

\newpage

# Project Conclusion

Through the analysis of the two original datasets from Kaggle along with the
two datasets retrieved from Wikipedia, the conclusion will answer the original
questions stated in the introduction. 

**How have popular songs change overtime?**

As noted in the introduction, music has evolved over time. Song attributes such
as acousticness, danceability, loudness, tempo, etc. have changed with culture
and generations. The findings in this project reflect that song popularity has
increased positively over time, with the highest frequency of popular songs in
the 2010s. Over time the types songs that are popular have also changed. For example,
alternative and Metal started off popular but over time have declined in popularity.
Some genres have increased in popularity over time, such as Country and Pop. In the
1970s and 1980s, songs with low to medium instrumentalness and low to medium
speechiness correlated with higher popularity. In the 1990s and 2000s, attributes 
such as acousticness and audio valence correlate with popular songs. Additionally,
in the 2000s, there was a shift to a new song attribute, danceability, which was 
found to be correlated with song popularity.

**Are there certain attributes that correlate with popular songs?**

The findings indicate that there are certain attributes that correlate with
popular songs. One of which was whether or not the song was a single. A single is
typically a song that is released earlier and separately from an album, and then
later, appear on an album. A single is important because it gives the consumer a
taste of the artists new music and helps promote the album because consumers
have gotten familiar to the music from the single song release. The data shows
that there is a positive correlation between a song single and song popularity.
Songs that come out as singles generally have 5%-10% greater song popularity.
The data also shows that popular songs are largely defined by the generation, and
what a particular generation is listening to and liking at the moment. In 1970 and
1980s, popular possessed qualities such as medium instrumentalness, high tempo,
low audio valence and medium speechiness. In the 1990s- 2010s, theres a shift
in song attributes related to popular songs with attributes such as medium to
high acousticness, high energy, high tempo, low speechiness, and high danceability.
From a general music knowledge, these findings make sense to see a shift
in song attributes of popular songs between decades. Music groups like boy bands
like *NSYNC or Backstreet Boys that promote songs that have high energy, high
danceability originated in the mid 1990s and these groups are believed to be two
of the most popular bands with many popular songs then and still now.

**Can the popularity of a song be predicted based on its attributes?**

The overall conclusion is yes, to a certain extend song popularity can be predicted.
It became evident, however, after applying several different prediction techniques
predicting song popularity is a regression problem and not a classification problem.
The classification models - decision tree, naive bayes, and k nearest neighbors -
did not perform well. Furthermore, even after removing outliers, tuning model
parameters, and discretizing the data in multiple ways, the classification models
still did not perform well. On the other hand, the regression models - support vector
machine, random forest, neural network, and general linear models - did perform well.
These models were able to predict song popularity with about 70% accuracy.

This is all taking into consideration major challenges with data quality. The data
was cleaned up as much as possible, including removing outliers, transforming
the song attributes, inferring missing song genres, and removing non-significant
variables, but yet was far from perfect. When taking that into consideration, it is
actually pretty impressive that the song popularity was able to be predicted as
accurately as it was. Not to mention, there were large biases in the types of songs
in the dataset due to the listening preferences and playlists of the creator of
the dataset, but this was mitigated prior to the predictive modeling by sampling
the data. The quality of the prediction can only be as good as the quality of the
data that goes into it.

**Final thoughts and next steps**

The results from this project were promising. This project shows that it is possible
to predict the popularity of a song. If more resources were invested to
improve the quality of the data, or to collect more data such as additional song
attributes, the accuracy of predicting the song popularity could only increase and it
may very well prove to be worthwhile for music production company to do. This project
only scratches the surface of what can potentially be done in the music industry,
but it does provide a foundation for analytics in music that can continue to built on.

```{r Appendix Song Info Web Scraping Code, include = FALSE, eval = FALSE}

#-------------------------------------------------------------------------------#
### initialization ###

require(dplyr)
require(rvest)
require(stringr)
require(sqldf)

#-------------------------------------------------------------------------------#
### web scraping function defined below ###

# test links - used these to build out the web scraping logic

# song_wiki_link <- 'https://en.wikipedia.org/wiki/Saturnz_Barz'
# song_wiki_link <- 'https://en.wikipedia.org/wiki/Hate_to_Say_I_Told_You_So'
# song_wiki_link <- 'https://en.wikipedia.org/wiki/Heaven%27s_Just_a_Sin_Away'
# song_wiki_link <- 'https://en.wikipedia.org/wiki/Good_Vibrations'
# song_wiki_link <- 'https://en.wikipedia.org/wiki/Love_Wins_(song)'

#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#

song_wiki_scraper <- function(song_wiki_link, song_info_part) {
  
# the song part info arg is an input as either 'Genre', 'Label', 'Songwriter', 'Producer'

#---------| read the web page based on the link that was input |----------------#

song_wiki_page <- read_html(song_wiki_link)

#------| return a list of all of the pieces of the infobox table |--------------#

iteration1  <-
song_wiki_page            %>%
html_nodes('.vevent')     %>%
html_children()           %>%
html_children()

#------| filter the list down just to the piece that contains the input |-------#

iteration2 <-
  iteration1[which(
    grepl(song_info_part, iteration1))]

#------| special case: if iteration2 has more than 1 element |------------------#

# does iteration2 contain more than one element?
testlength <- ifelse(
  length(iteration2) > 1,
    TRUE, FALSE)

# if yes, are the elements the same or are they different?
testunique <- ifelse(
  testlength == TRUE &
  length(unique(as.character(iteration2))) == 1,
    TRUE, FALSE)

# if they are the same, just take the first element
iteration2 <- if (testunique == TRUE) {
  iteration2[1]} else {iteration2}

# if they are not the same, does one of them have a seperated hlist?
# if yes, take that one, otherwise just take the first element
iteration2 <- if (testunique == FALSE & testlength == TRUE) {
  if (max(TRUE %in% grepl('hlist-separated', iteration2))) {
    iteration2[which(grepl('hlist-separated', iteration2))]} else {
      iteration2[1]}} else {
        iteration2}

#------| check if iteration2 is nested with more than one |---------------------#

# if there is a seperated hlist, then it is nested
testnested <- ifelse(
  max(grepl('hlist-separated', iteration2)), TRUE, FALSE)

# special case: sometimes it is nested without a seperated hlist
testnested2 <- ifelse(
  testnested == FALSE &
  length(unique(unlist(str_locate_all(iteration2 %>% html_text(), '\\n')))) > 1,
    TRUE, FALSE)

#-----------------------| drill down into the next level |----------------------#

iteration3 <- 
  iteration2 %>%
  html_children()

#-----------------| ilter it down to just the piece that we need |--------------#

iteration4 <-
  iteration3[which(
    grepl('hlist', iteration3))]

#----------| parse the final result dependent on several logical tests|---------#

iteration5 <-
  
  # if it was hidden nested, drill down 2 layers then get text
  
  if (testnested2 == TRUE)  {
         iteration4       %>%
         html_children()  %>%
         html_children()  %>%
         html_text()        }   else if (
  
  # if it was not nested, no drill down necessary just get text
           
      testnested == FALSE) {
         iteration4       %>%
         html_text()        }   else    {

  # if it was nested but not hidden nested, drill down 3 layers then get text
           
        iteration4        %>%
        html_children()   %>%
        html_children()   %>%
        html_children()   %>%
        html_text()         }

#-------------------| clean up the result as needed |---------------------------#

# get rid of (s) where it exists. no effect if it doesn't exist.
# for example, sometimes there is songwriter or songwriter(s)

iteration5 <-
  trimws(gsub('\\(s\\)', '', iteration5))

#-------------------------------------------------------------------------------#

return(iteration5)
}

#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#

#-------------------------------------------------------------------------------#
# initialize empty song master. Each record will be appended here.

song_master <- data.frame()

#-------------------------------------------------------------------------------#
# initialize the for loop. Currently testing on 50 iterations.

for (song_artist_ID in paste(song_artist$song_name, song_artist$artist_name)) {

#-------------------------------------------------------------------------------#
# initialize variables. Everything starts off as NA - values assigned later

song_wiki_link    <- NA
song_single       <- NA
song_isalbum      <- NA
song_isartist     <- NA
song_released     <- NA
song_genre        <- NA
song_label        <- NA
song_songwriter   <- NA
song_producer     <- NA

#-------------------------------------------------------------------------------#
# dynamically create the google search link

song_search_text <- paste(song_artist_ID, 'Wikipedia')
song_search_text <- gsub(' ', '+', song_search_text) %>% trimws()
song_google_link <- paste0("https://www.google.com/search?q=", song_search_text)

#-------------------------------------------------------------------------------#
# attempt to retrieve wikipedia link from google search

# read the first page of google search
skiptonext <- FALSE
tryCatch({
  
google_search_page <- read_html(song_google_link)

# this is a list of links on google page
google_search_result <-
  (google_search_page    %>%
  html_nodes('body')     %>%
  html_children()) [2]   %>%
  html_children()        %>%
  html_nodes('a')        %>%
  html_attr('href')},

error = function(e) {skiptonext <<- TRUE})

if (skiptonext) { #----------------------
song_sample$song_wiki_link    <- NA
song_sample$song_single       <- NA
song_sample$song_isalbum      <- NA
song_sample$song_isartist     <- NA
song_sample$song_released     <- NA
song_sample$song_genre        <- NA
song_sample$song_label        <- NA
song_sample$song_songwriter   <- NA
song_sample$song_producer     <- NA

print(paste0('song skipped: ', song_artist_ID, ' #',
   which(paste(song_artist$song_name, song_artist$artist_name) == song_artist_ID)))}

if (skiptonext) {next} #----------------

# parse the list of links for the wiki link
# if the wiki link is there, this is reliable
# if its not there, might return incorrect link
# if its incorrect, program will realize it later
google_search_result <- 
  google_search_result[which(
    ! grepl('search?', google_search_result) &
    ! grepl('maps'   , google_search_result) &
      grepl('url?'   , google_search_result) &
      grepl('wiki'   , google_search_result))] [1]

# the link is buried between two characters
# the first character is the first '=' sign of the string
parseresultstart <-
  unique(unlist(str_locate_all(google_search_result, '='))) [1] + 1

# the second character is the first '&' sign of the string
parseresultend <-
  unique(unlist(str_locate_all(google_search_result, '&'))) [1] - 1

# extract the link by the text between the '=' and the '&'
song_wiki_link <-
  substr(google_search_result, parseresultstart, parseresultend)

# special case: if there is an apostrophe in the link,
# it gets encoded as '%27'. The '%', however, is
# also encoded as '%25'. Decode '%25' to '%' for it to work.
song_wiki_link <-
  gsub('\\%25', '\\%', song_wiki_link)

song_sample <- data.frame(cbind('song_wiki_link' = song_wiki_link,
                                'song_id' = song_artist_ID))

#-------------------------------------------------------------------------------#
# read the page, if error, everything NA and skip to next loop iteration

skiptonext <- FALSE
tryCatch({
  song_wiki_page <- 
    read_html(song_wiki_link)},
error = function(e) {
  skiptonext <<- TRUE})

if (skiptonext) { #----------------------
song_sample$song_wiki_link    <- NA
song_sample$song_single       <- NA
song_sample$song_isalbum      <- NA
song_sample$song_isartist     <- NA
song_sample$song_released     <- NA
song_sample$song_genre        <- NA
song_sample$song_label        <- NA
song_sample$song_songwriter   <- NA
song_sample$song_producer     <- NA

print(paste0('song skipped: ', song_artist_ID, ' #',
   which(paste(song_artist$song_name, song_artist$artist_name) == song_artist_ID)))}

if (skiptonext) {next} #----------------

#-------------------------------------------------------------------------------#
# attempt to scrape the info box, 

tryCatch(                   {
  song_table_infobox        <-
    song_wiki_page          %>%
    html_nodes('.vevent')   %>%
    html_table()            %>%
    .[[1]]                  },
error = function(e)         {
  song_table_infobox   <<- NA
  song_single          <<- NA
  song_isalbum         <<- NA})

#-------------------------------------------------------------------------------#
# if it has the word Single it is a single

tryCatch({
  song_single <- ifelse(
    max(TRUE %in%
      grepl('Single', song_table_infobox)),
  1, 0) },
error = function(e) {
  song_single <<- NA})

#-------------------------------------------------------------------------------#
# if it has the word Album it is an album

tryCatch({
  song_isalbum <- ifelse(
    max(TRUE %in%
      grepl('Album', song_table_infobox)),
    1, 0) },
error = function(e) {
  song_isalbum <<- NA})

#-------------------------------------------------------------------------------#
# attempt to scrape .plainlist node then check it for certain text
# words like 'Origin', 'Born', indicate artist homepage and not a song

tryCatch({
  song_artist_test <-
    as.character(
    (song_wiki_page %>%
    html_nodes('.plainlist')) [1])},
error = function(e) {
  song_artist_test <<- NA})

tryCatch({
song_isartist <- ifelse(
  max(TRUE %in%
    grepl('Origin',         song_artist_test)   |
    grepl('Born',           song_artist_test)   |
    grepl('Occupation',     song_artist_test)   |
    grepl('Years active',   song_artist_test)   |
    grepl('Website',        song_artist_test)  ),
  1, 0)},
error = function(e) {
  song_isartist <<- NA})

#-------------------------------------------------------------------------------#
# retrieve the song release date based on the infobox node

tryCatch({
  song_released <- as.character(
    song_table_infobox[which(
      song_table_infobox[,1] == 'Released'), 2]) },
error = function(e) {
  song_released <<- NA})

#-------------------------------------------------------------------------------#
# move forward with scraping song data

tryCatch({song_genre        <- song_wiki_scraper(song_wiki_link, 'Genre')},
  error = function(e) {song_genre        <<- NA})

tryCatch({song_label        <- song_wiki_scraper(song_wiki_link , 'Label')},
  error = function(e) {song_label        <<- NA})

tryCatch({song_songwriter   <- song_wiki_scraper(song_wiki_link , 'Songwriter')},
  error = function(e) {song_songwriter   <<- NA})

tryCatch({song_producer     <- song_wiki_scraper(song_wiki_link , 'Producer')},
  error = function(e) {song_producer     <<- NA})

#-------------------------------------------------------------------------------#
# piece it all together into the song sample dataframe

song_sample$song_wiki_link    <- song_wiki_link
song_sample$song_single       <- song_single
song_sample$song_isalbum      <- song_isalbum
song_sample$song_isartist     <- song_isartist
song_sample$song_released     <- song_released
song_sample$song_genre        <- list(song_genre)
song_sample$song_label        <- list(song_label)
song_sample$song_songwriter   <- list(song_songwriter)
song_sample$song_producer     <- list(song_producer)

#-------------------------------------------------------------------------------#
# append the data to the main dataframe and print progress update

song_master <- rbind(song_master, song_sample)

print(paste0('song recorded: ', song_artist_ID, ' #',
   which(paste(song_artist$song_name, song_artist$artist_name) == song_artist_ID)))

}

#-------------------------------------------------------------------------------#

```

```{r Appendix Artist Web Scraping Code, eval = FALSE, include = FALSE}

#-------------------------------------------------------------------------------#
### initialization ###

require(dplyr)
require(rvest)
require(stringr)
require(sqldf)

#-------------------------------------------------------------------------------#
### web scraping function defined below ###

# test links - used these to build out the web scraping logic

# artist_wiki_link <- 'https://en.wikipedia.org/wiki/ASAP_Ferg'
# artist_wiki_link <- 'https://en.wikipedia.org/wiki/Chris_Webby'
# artist_wiki_link <- 'https://en.wikipedia.org/wiki/Kream'
# artist_wiki_link <- 'https://en.wikipedia.org/wiki/Nina_Sky_(album)'
# artist_wiki_link <- 'https://en.wikipedia.org/wiki/Psychotic_Reaction_(album)'

#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#

artist_wiki_scraper <- function(artist_wiki_link, artist_info_part) {

#---------| read the web page based on the link that was input |----------------#

artist_wiki_page <- read_html(artist_wiki_link)

#------| return a list of all of the pieces of the infobox table |--------------#

iteration1  <-
artist_wiki_page          %>%
html_nodes('.plainlist')  %>%
html_children()           %>%
html_children()

#------| filter the list down just to the piece that contains the input |-------#

iteration2 <-
  iteration1[which(
    grepl(artist_info_part, iteration1))]
  
#------| special case: if iteration2 has more than 1 element |------------------#

# does iteration2 contain more than one element?
testlength <- ifelse(
  length(iteration2) > 1,
    TRUE, FALSE)

# if yes, are the elements the same or are they different?
testunique <- ifelse(
  testlength == TRUE &
  length(unique(as.character(iteration2))) == 1,
    TRUE, FALSE)

# if they are the same, just take the first element
iteration2 <- if (testunique == TRUE) {
  iteration2[1]} else {iteration2}

# if they are not the same, does one of them have a seperated hlist?
# if yes, take that one, otherwise just take the first element
iteration2 <- if (testunique == FALSE & testlength == TRUE) {
  if (max(TRUE %in% grepl('hlist-separated', iteration2))) {
    iteration2[which(grepl('hlist-separated', iteration2))]} else {
      iteration2[1]}} else {
        iteration2}

#------| check if iteration2 is nested with more than one |---------------------#

# if there is a seperated hlist, then it is nested
tryCatch({
testnested <- ifelse(
  max(grepl('hlist-separated', iteration2)), TRUE, FALSE)}, warning = function(e) {
    testnested <<- FALSE})

# special case: sometimes it is nested without a seperated hlist
testnested2 <- ifelse(
  testnested == FALSE &
  length(unique(unlist(str_locate_all(iteration2 %>% html_text(), '\\n')))) > 1,
    TRUE, FALSE)

#-----------------------| drill down into the next level |----------------------#

iteration3 <- 
  iteration2 %>%
  html_children()

#-----------------| ilter it down to just the piece that we need |--------------#

iteration4 <-
  iteration3[which(
    grepl('hlist', iteration3))]

#----------| parse the final result dependent on several logical tests|---------#

iteration5 <-
  
  # if it was hidden nested, drill down 2 layers then get text
  
  if (testnested2 == TRUE)  {
         iteration4       %>%
         html_children()  %>%
         html_children()  %>%
         html_text()        }   else if (
  
  # if it was not nested, no drill down necessary just get text
           
      testnested == FALSE) {
         iteration4       %>%
         html_text()        }   else    {

  # if it was nested but not hidden nested, drill down 3 layers then get text
           
        iteration4        %>%
        html_children()   %>%
        html_children()   %>%
        html_children()   %>%
        html_text()         }

#-------------------| clean up the result as needed |---------------------------#

# get rid of (s) where it exists. no effect if it doesn't exist.
# for example, sometimes there is songwriter or songwriter(s)

iteration5 <-
  trimws(gsub('\\(s\\)', '', iteration5))

#-------------------------------------------------------------------------------#

return(iteration5)
}

#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#

#-------------------------------------------------------------------------------#
# initialize empty artist master. Each record will be appended here.

artist_T <- unique(as.data.frame(cbind('song_artist' = song_artist[,2])))

artist_master <- data.frame()

#-------------------------------------------------------------------------------#
# initialize the for loop. Currently testing on 50 iterations.

for (artist_ID in artist_T$song_artist) {

#-------------------------------------------------------------------------------#
# initialize variables. Everything starts off as NA - values assigned later

artist_wiki_link    <- NA
artist_born         <- NA
artist_origin       <- NA
artist_yearsactive  <- NA
artist_genre        <- NA
artist_label        <- NA

#-------------------------------------------------------------------------------#
# dynamically create the google search link

artist_search_text <- paste(artist_ID, 'Wikipedia')
artist_search_text <- gsub(' ', '+', artist_search_text) %>% trimws()
artist_google_link <- paste0("https://www.google.com/search?q=", artist_search_text)

#-------------------------------------------------------------------------------#
# attempt to retrieve wikipedia link from google search

# read the first page of google search
skiptonext <- FALSE
tryCatch({
  
google_search_page <- read_html(artist_google_link)

# this is a list of links on google page
google_search_result <-
  (google_search_page    %>%
  html_nodes('body')     %>%
  html_children()) [2]   %>%
  html_children()        %>%
  html_nodes('a')        %>%
  html_attr('href')},

error = function(e) {skiptonext <<- TRUE})

if (skiptonext) { #----------------------
artist_sample$artist_wiki_link    <- NA
artist_sample$artist_born         <- NA
artist_sample$artist_origin       <- NA
artist_sample$artist_yearsactive  <- NA
artist_sample$genre               <- NA
artist_sample$label               <- NA

print(paste0('artist skipped: ', artist_ID, ' #',
   which(artist_T$song_artist == artist_ID)))}

if (skiptonext) {next} #----------------

# parse the list of links for the wiki link
# if the wiki link is there, this is reliable
# if its not there, might return incorrect link
# if its incorrect, program will realize it later
skiptonext <- FALSE
google_search_result <- 
  google_search_result[which(
    ! grepl('search?', google_search_result) &
    ! grepl('maps'   , google_search_result) &
      grepl('url?'   , google_search_result) &
      grepl('wiki'   , google_search_result))] [1]
if (is.na(google_search_result)) {skiptonext <<- TRUE}
if (skiptonext) {next}

# the link is buried between two characters
# the first character is the first '=' sign of the string
parseresultstart <-
  unique(unlist(str_locate_all(google_search_result, '='))) [1] + 1

# the second character is the first '&' sign of the string
parseresultend <-
  unique(unlist(str_locate_all(google_search_result, '&'))) [1] - 1

# extract the link by the text between the '=' and the '&'
artist_wiki_link <-
  substr(google_search_result, parseresultstart, parseresultend)

# special case: if there is an apostrophe in the link,
# it gets encoded as '%27'. The '%', however, is
# also encoded as '%25'. Decode '%25' to '%' for it to work.
artist_wiki_link <-
  gsub('\\%25', '\\%', artist_wiki_link)

artist_sample <- data.frame(cbind('artist_wiki_link' = artist_wiki_link,
                                'artist_id' = artist_ID))

#-------------------------------------------------------------------------------#
# read the page, if error, everything NA and skip to next loop iteration

skiptonext <- FALSE
tryCatch({
  artist_wiki_page <- 
    read_html(artist_wiki_link)},
error = function(e) {
  skiptonext <<- TRUE})

if (skiptonext) { #----------------------
artist_sample$artist_wiki_link    <- NA
artist_sample$artist_born         <- NA
artist_sample$artist_origin       <- NA
artist_sample$artist_yearsactive  <- NA
artist_sample$artist_genre        <- NA
artist_sample$artist_label        <- NA

print(paste0('artist skipped: ', artist_ID, ' #',
   which(artist_T$song_artist == artist_ID)))}

if (skiptonext) {next} #----------------

#-------------------------------------------------------------------------------#
# attempt to scrape the info box, 

infoboxskip <- FALSE
tryCatch(                   {
  artist_table_infobox        <-
    (artist_wiki_page          %>%
    html_nodes('.plainlist')) [1] %>%
    html_table() %>%
    .[[1]]                  },
error = function(e)         {
  artist_table_infobox   <<- NA
  infoboxskip <<- TRUE})

if (infoboxskip) {
tryCatch(                   {
  artist_table_infobox        <-
    (artist_wiki_page          %>%
    html_nodes('.vcard')) [1] %>%
    html_table() %>%
    .[[1]]                  },
error = function(e)         {
  artist_table_infobox   <<- NA})}
    
#-------------------------------------------------------------------------------#
# retrieve information from the infobox node

# standardize to lower case
tryCatch({
lowercase_artist_info <- c()
for (element in artist_table_infobox[,1]) {
newelement <- tolower(element)
lowercase_artist_info <- c(lowercase_artist_info, newelement)}
new_artist_table_infobox <- data.frame(cbind('X1' = lowercase_artist_info, 'X2' = artist_table_infobox[,2]))
}, error = function(e) {new_artist_table_infobox <<- NA})

# when the artist was born
tryCatch({
  artist_born <- as.character(
    new_artist_table_infobox[which(
      new_artist_table_infobox[,1] == 'born'), 2]) },
error = function(e) {
  artist_born <<- NA})

# where the artist was born
tryCatch({
  artist_origin <- as.character(
    new_artist_table_infobox[which(
      new_artist_table_infobox[,1] == 'origin'), 2]) },
error = function(e) {
  artist_origin <<- NA})

# years the artist has been active
tryCatch({
  artist_yearsactive <- as.character(
    new_artist_table_infobox[which(grepl('years',
      new_artist_table_infobox[,1])), 2]) },
error = function(e) {
  artist_yearsactive <<- NA})

tryCatch({artist_genre        <- artist_wiki_scraper(artist_wiki_link, 'Genre')},
  error = function(e) {artist_genre        <<- NA})

if (length(artist_genre) == 0) {
  
    tryCatch({
    artist_genre <- as.character(
      new_artist_table_infobox[which(grepl('genre',
        new_artist_table_infobox[,1])), 2]) },
  error = function(e) {
    artist_genre <<- NA})}

tryCatch({artist_label        <- artist_wiki_scraper(artist_wiki_link, 'Label')},
  error = function(e) {artist_label        <<- NA})

if (length(artist_label) == 0) {
  
    tryCatch({
    artist_label <- as.character(
      new_artist_table_infobox[which(grepl('label',
        new_artist_table_infobox[,1])), 2]) },
  error = function(e) {
    artist_label <<- NA})}

#-------------------------------------------------------------------------------#
# piece it all together into the song sample dataframe

tryCatch({artist_sample$artist_wiki_link    <- artist_wiki_link}, error = function(e) {artist_sample$artist_wiki_link    <<- NA})
tryCatch({artist_sample$artist_born     <- artist_born }, error = function(e) {artist_sample$artist_born     <<- NA})
tryCatch({artist_sample$artist_origin    <- artist_origin}, error = function(e) {artist_sample$artist_origin    <<- NA})
tryCatch({artist_sample$artist_yearsactive    <- artist_yearsactive}, error = function(e) {artist_sample$artist_yearsactive    <<- NA})
tryCatch({artist_sample$artist_genre    <- list(artist_genre)}, error = function(e) {artist_sample$artist_genre    <<- NA})
tryCatch({artist_sample$artist_label    <- list(artist_label)}, error = function(e) {artist_sample$artist_label    <<- NA})

#-------------------------------------------------------------------------------#
# append the data to the main dataframe and print progress update

artist_master <- rbind(artist_master, artist_sample)

print(paste0('artist recorded: ', artist_ID, ' #',
   which(artist_T$song_artist == artist_ID)))

}

```

```{r Appendix Song and Artist Info Clean Up, eval = FALSE, include = FALSE}

#-------------------------------------------------------------------------------#
### initialization ###

require(dplyr)
require(rvest)
require(stringr)
require(sqldf)

#-------------------------------------------------------------------------------#

# clean up the environment

rm(google_search_page, song_wiki_page, google_search_result, parseresultend,
parseresultstart, skiptonext, song_artist_test, song_genre, song_google_link,
song_isalbum, song_isartist, song_label, song_producer, song_released, song_search_text,
song_single, song_songwriter, song_table_infobox, song_wiki_link, song_sample, song_artist_ID)

#-------------------------------------------------------------------------------#
# eliminate anything that came out in brackets/parentheses

song_master$song_released <- gsub("\\(.*?)", "", song_master$song_released)
song_master$song_released <- gsub("\\[.*?]", "", song_master$song_released)

#-------------------------------------------------------------------------------#
# create a function to extract the month from the date

date_parse_month <- function(x, month3dig, month2dig) {
  
  tryCatch({
    month <- if (grepl(month3dig, x)) {month2dig} else {NA}},
  error = function(e) {month <<- NA})
  
return(month)
}

#-------------------------------------------------------------------------------#
# create a seperate column for whether the month is recognized
# this is because some songs had a second or third release

song_master$song_release_jan <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Jan', month2dig = '01'))
song_master$song_release_feb <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Feb', month2dig = '02'))
song_master$song_release_mar <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Mar', month2dig = '03'))
song_master$song_release_apr <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Apr', month2dig = '04'))
song_master$song_release_may <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'May', month2dig = '05'))
song_master$song_release_jun <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Jun', month2dig = '06'))
song_master$song_release_jul <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Jul', month2dig = '07'))
song_master$song_release_aug <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Aug', month2dig = '08'))
song_master$song_release_sep <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Sep', month2dig = '09'))
song_master$song_release_oct <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Oct', month2dig = '10'))
song_master$song_release_nov <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Nov', month2dig = '11'))
song_master$song_release_dec <- unlist(lapply(song_master$song_released, FUN = date_parse_month, month3dig = 'Dec', month2dig = '12'))

#-------------------------------------------------------------------------------#
# add a column for which records have multiple months and which do not
monthcount <- c()
for (rownumber in seq(1, dim(song_master) [1] )) {
  count <- 0
  for (element in song_master[rownumber, 11:(dim(song_master) [2] )]) {
    count <- count + ! is.na(element)
  }
  monthcount <- c(monthcount, count)
}
song_master$monthcount <- monthcount   

# show the count of records for each count of months
# the large majority have one. 3 and 4 were very uncommon.
table(song_master$monthcount)

#-------------------------------------------------------------------------------#
# extract the month that we want to use for the song release month
# for the ones with more than one date, take the first one
# this will be considered the 'original' song release date

monthextract <- c()
for (rownumber in seq(1, dim(song_master) [1] )) {
  targetcolumnposition <- which(! is.na(song_master[rownumber, 11:((dim(song_master) [2] ) - 1)])) [1]
  targetcolumnposition <- ifelse(is.na(targetcolumnposition), NA, targetcolumnposition)
  targetcolumnvalue <- ifelse(is.na(targetcolumnposition), NA, song_master[rownumber, targetcolumnposition+10])
  monthextract <- c(monthextract, targetcolumnvalue)}
song_master$song_release_month <- monthextract
  
# show how many songs there are for each month
table(song_master$song_release_month)

# remove all of the columns that are no longer needed
song_master <- song_master[, -c(11:23)]

#-------------------------------------------------------------------------------#
# extract the year and date from the song release date

dayyearnumbers <- c()
for (rownumber in seq(1, dim(song_master) [1] )) {
  for (element in song_master[rownumber, 6]) {
    separatedtext <- as.numeric(str_split(element, "") %>% .[[1]])
    seperatedtext <- which(! is.na(separatedtext))
    newseparatedtext <- c()
      for (textelement in separatedtext) {
        newseparatedtext <- paste0(newseparatedtext, textelement)}
        newseparatedtext <- trimws(gsub('NA', '', newseparatedtext))
        dayyearnumbers <- c(dayyearnumbers, newseparatedtext)
  }
}

song_master$song_release_year <- dayyearnumbers

song_master$song_release_year <- substr(
  song_master$song_release_year,
  nchar(song_master$song_release_year) - 3,
  nchar(song_master$song_release_year))

song_master[which(song_master$song_release_year == ''),
            which(colnames(song_master) == 'song_release_year')] <- NA

song_master$song_release_month <- as.numeric(song_master$song_release_month)
song_master$song_release_year <- as.numeric(song_master$song_release_year)

#-------------------------------------------------------------------------------#
# explore the song genres

# create a seperate table where the work will be completed
song_genre_extract <- song_master[, which(colnames(song_master) %in% c('song_id', 'song_genre'))]

# create a unique list of all of the genres
unique_genre_list <- c()

  for (rownumber in seq(1, dim(song_genre_extract) [1] )) {
    for (element in song_genre_extract[rownumber, 2]) {
      unique_genre_list <- c(unique_genre_list, element)}}

#get rid of brackets and make it unique
unique_genre_list <- gsub("\\[.*?]", "", unique_genre_list)
unique_genre_list <- tolower(unique_genre_list)
unique_genre_list <- unique(unique_genre_list)

# further extract each individual genre
# each genre should be a single word
new_unique_genre_list <- c()

  for (rownumber in seq(1, length(unique_genre_list))) {
    for (element in unique_genre_list[rownumber]) {
      atomicelement <- str_split(element, ' ') %>% .[[1]]
      new_unique_genre_list <- c(new_unique_genre_list, atomicelement)}}

new_unique_genre_list <- unique(new_unique_genre_list)

#-------------------------------------------------------------------------------#
# extract the song genres
# the parent genre will be selected out of the following
# alternative, rock, metal, punk, pop, r&b, rap, jazz, blues, folk, country, electronic, other

parent_genre_list <- c('alt', 'rock', 'metal', 'punk', 'pop', 'r&b',
                       'rap', 'jazz', 'blues', 'folk', 'country', 'elect', 'other')

# create a data frame for the parent genres
genre_dataframe <- data.frame()

  for (rownumber in seq(1, dim(song_genre_extract) [1] )) {
    for (element in song_genre_extract[rownumber, 2]) {
      newelement <- gsub("\\[.*?]", "", element)
      newelement <- tolower(newelement)
      genre_list <- c()
      for (genre in parent_genre_list) {
        check_genre <- tryCatch({ifelse(max(grepl(genre, newelement)), 1, 0)},
                       warning = function(e) {check_genre <<- 0})
        genre_list <- c(genre_list, check_genre)}
      genre_dataframe <- rbind(genre_dataframe, genre_list)}}
        
song_genre_extract <- cbind(song_genre_extract, genre_dataframe)
colnames(song_genre_extract) [3:15] <- parent_genre_list

# fill in the values for the other column
# if none of the parent genres were found, then it is considered to be other

song_genre_extract$song_genre <- ifelse(song_genre_extract$song_genre == 'character(0)', NA, song_genre_extract$song_genre)
song_genre_extract$other <- ifelse(rowSums(song_genre_extract[,3:14]) == 0, 1, 0)
song_genre_extract$other <- ifelse(is.na(song_genre_extract$song_genre), 0, song_genre_extract$other)

song_genre <- song_genre_extract
rm(song_genre_extract)

#-------------------------------------------------------------------------------#
# create the song label dataframe
# some songs have multiple labels, take only the top 3 labels

song_label              <- song_master[, which(colnames(song_master) %in% c('song_id', 'song_label'))]
song_label$song_label   <- ifelse(song_label$song_label == 'character(0)', NA, song_label$song_label)

song_label_dataframe <- data.frame()
  for (rownumber in seq(1, dim(song_label) [1] )) {
    song_label_row <- c()
    for (element in song_label[rownumber, 2]) {
      song_label_row <- c(song_label_row, element)
      song_label1 <- song_label_row[1]
      song_label2 <- song_label_row[2]
      song_label3 <- song_label_row[3]
      song_label_matrix <- data.frame(cbind(
        'label1' = song_label1,
        'label2' = song_label2,
        'label3' = song_label3))
      song_label_dataframe <- rbind(song_label_dataframe, song_label_matrix)}}

# standardize the text. get rid of anything in (), [], and change to lower case
song_label_dataframe$label1 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_label_dataframe$label1)))
song_label_dataframe$label2 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_label_dataframe$label2)))
song_label_dataframe$label3 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_label_dataframe$label3)))

# append to main
song_label <- cbind(song_label, song_label_dataframe)

# create a temporary dataframe to correct the records still separated by comma
temp_label_dataframe <- song_label[which(grepl(',', song_label$label1)),]  # these are the records seperated by comma
temp_label_dataframe$label1 <- gsub(' ', '', temp_label_dataframe$label1)  # remove spaces this will help with the next step

# create a for loop to extract the comma separated labels
temp_label1 <- c()
temp_label2 <- c()
temp_label3 <- c()
for (temp_label in temp_label_dataframe$label1) {
num_commas <- length(unique(unlist(str_locate_all(temp_label, ','))))
comma_pos1 <- unique(unlist(str_locate_all(temp_label, ','))) [1]
comma_pos2 <- unique(unlist(str_locate_all(temp_label, ','))) [2]
split1 <- substr(temp_label, 1, comma_pos1 - 1)
split2 <- ifelse(num_commas > 1, 
                 substr(temp_label, comma_pos1 + 1, comma_pos2 - 1),
                 substr(temp_label, comma_pos1 + 1, nchar(temp_label)))
split3 <- substr(temp_label, comma_pos2 + 1, nchar(temp_label))
temp_label1 <- c(temp_label1, split1)
temp_label2 <- c(temp_label2, split2)
temp_label3 <- c(temp_label3, split3)
}

# join the labels into the temporary dataframe
temp_label_dataframe <- data.frame(cbind(temp_label_dataframe, temp_label1, temp_label2, temp_label3))
temp_label_dataframe <- temp_label_dataframe[, c(1, 6, 7, 8)]

# join the new labels into the main dataframe
song_label <- left_join(song_label, temp_label_dataframe, on = 'song_id')
song_label$label1 <- ifelse(is.na(song_label$temp_label1), song_label$label1, song_label$temp_label1)
song_label$label2 <- ifelse(is.na(song_label$temp_label2), song_label$label2, song_label$temp_label2)
song_label$label3 <- ifelse(is.na(song_label$temp_label3), song_label$label3, song_label$temp_label3)
song_label <- song_label[, - which(colnames(song_label) %in% c('temp_label1', 'temp_label2', 'temp_label3'))]

# clean up the environment
rm(temp_label_dataframe, comma_pos1, comma_pos2, num_commas, split1, split2, split3,
temp_label1, temp_label2, temp_label3, temp_label)

#-------------------------------------------------------------------------------#
# create the song songwriter dataframe
# some songs have multiple songwriters, take only the top 3 songwriters

song_songwriter                   <- song_master[, which(colnames(song_master) %in% c('song_id', 'song_songwriter'))]
song_songwriter$song_songwriter   <- ifelse(song_songwriter$song_songwriter == 'character(0)', NA, song_songwriter$song_songwriter)

song_songwriter_dataframe <- data.frame()
  for (rownumber in seq(1, dim(song_songwriter) [1] )) {
    song_songwriter_row <- c()
    for (element in song_songwriter[rownumber, 2]) {
      song_songwriter_row <- c(song_songwriter_row, element)
      song_songwriter1 <- song_songwriter_row[1]
      song_songwriter2 <- song_songwriter_row[2]
      song_songwriter3 <- song_songwriter_row[3]
      song_songwriter_matrix <- data.frame(cbind(
        'songwriter1' = song_songwriter1,
        'songwriter2' = song_songwriter2,
        'songwriter3' = song_songwriter3))
      song_songwriter_dataframe <- rbind(song_songwriter_dataframe, song_songwriter_matrix)}}

# standardize the text. get rid of anything in (), [], and change to lower case
song_songwriter_dataframe$songwriter1 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_songwriter_dataframe$songwriter1)))
song_songwriter_dataframe$songwriter2 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_songwriter_dataframe$songwriter2)))
song_songwriter_dataframe$songwriter3 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_songwriter_dataframe$songwriter3)))

# append to main
song_songwriter <- cbind(song_songwriter, song_songwriter_dataframe)

# create a temporary dataframe to correct the records still separated by comma
temp_songwriter_dataframe <- song_songwriter[which(grepl(',', song_songwriter$songwriter1)),]  # these are the records seperated by comma
temp_songwriter_dataframe$songwriter1 <- gsub(' ', '', temp_songwriter_dataframe$songwriter1)  # remove spaces this will help with the next step

# create a for loop to extract the comma separated songwriters
temp_songwriter1 <- c()
temp_songwriter2 <- c()
temp_songwriter3 <- c()
for (temp_songwriter in temp_songwriter_dataframe$songwriter1) {
num_commas <- length(unique(unlist(str_locate_all(temp_songwriter, ','))))
comma_pos1 <- unique(unlist(str_locate_all(temp_songwriter, ','))) [1]
comma_pos2 <- unique(unlist(str_locate_all(temp_songwriter, ','))) [2]
comma_pos3 <- unique(unlist(str_locate_all(temp_songwriter, ','))) [3]
split1 <- substr(temp_songwriter, 1, comma_pos1 - 1)
split2 <- ifelse(num_commas > 1, 
                 substr(temp_songwriter, comma_pos1 + 1, comma_pos2 - 1),
                 substr(temp_songwriter, comma_pos1 + 1, nchar(temp_songwriter)))
split3 <- ifelse(num_commas > 2,
                 substr(temp_songwriter, comma_pos2 + 1, comma_pos3 - 1),
                 substr(temp_songwriter, comma_pos2 + 1, nchar(temp_songwriter)))
temp_songwriter1 <- c(temp_songwriter1, split1)
temp_songwriter2 <- c(temp_songwriter2, split2)
temp_songwriter3 <- c(temp_songwriter3, split3)
}

# join the songwriters into the temporary dataframe
temp_songwriter_dataframe <- data.frame(cbind(temp_songwriter_dataframe, temp_songwriter1, temp_songwriter2, temp_songwriter3))
temp_songwriter_dataframe <- temp_songwriter_dataframe[, c(1, 6, 7, 8)]

# join the new songwriters into the main dataframe
song_songwriter <- left_join(song_songwriter, temp_songwriter_dataframe, on = 'song_id')
song_songwriter$songwriter1 <- ifelse(is.na(song_songwriter$temp_songwriter1), song_songwriter$songwriter1, song_songwriter$temp_songwriter1)
song_songwriter$songwriter2 <- ifelse(is.na(song_songwriter$temp_songwriter2), song_songwriter$songwriter2, song_songwriter$temp_songwriter2)
song_songwriter$songwriter3 <- ifelse(is.na(song_songwriter$temp_songwriter3), song_songwriter$songwriter3, song_songwriter$temp_songwriter3)
song_songwriter <- song_songwriter[, - which(colnames(song_songwriter) %in% c('temp_songwriter1', 'temp_songwriter2', 'temp_songwriter3'))]

# clean up the environment
rm(temp_songwriter_dataframe, comma_pos1, comma_pos2, num_commas, split1, split2, split3,
temp_songwriter1, temp_songwriter2, temp_songwriter3, temp_songwriter, comma_pos3)

#-------------------------------------------------------------------------------#
# create the song producer dataframe
# some songs have multiple producers, take only the top 3 producers

song_producer                 <- song_master[, which(colnames(song_master) %in% c('song_id', 'song_producer'))]
song_producer$song_producer   <- ifelse(song_producer$song_producer == 'character(0)', NA, song_producer$song_producer)

song_producer_dataframe <- data.frame()
  for (rownumber in seq(1, dim(song_producer) [1] )) {
    song_producer_row <- c()
    for (element in song_producer[rownumber, 2]) {
      song_producer_row <- c(song_producer_row, element)
      song_producer1 <- song_producer_row[1]
      song_producer2 <- song_producer_row[2]
      song_producer3 <- song_producer_row[3]
      song_producer_matrix <- data.frame(cbind(
        'producer1' = song_producer1,
        'producer2' = song_producer2,
        'producer3' = song_producer3))
      song_producer_dataframe <- rbind(song_producer_dataframe, song_producer_matrix)}}

# standardize the text. get rid of anything in (), [], and change to lower case
song_producer_dataframe$producer1 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_producer_dataframe$producer1)))
song_producer_dataframe$producer2 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_producer_dataframe$producer2)))
song_producer_dataframe$producer3 <- tolower(gsub("\\[.*?]", "", gsub("\\(.*?)", "", song_producer_dataframe$producer3)))

# append to main
song_producer <- cbind(song_producer, song_producer_dataframe)

# create a temporary dataframe to correct the records still separated by comma
temp_producer_dataframe <- song_producer[which(grepl(',', song_producer$producer1)),]  # these are the records seperated by comma
temp_producer_dataframe$producer1 <- gsub(' ', '', temp_producer_dataframe$producer1)  # remove spaces this will help with the next step

# create a for loop to extract the comma separated producers
temp_producer1 <- c()
temp_producer2 <- c()
temp_producer3 <- c()
for (temp_producer in temp_producer_dataframe$producer1) {
num_commas <- length(unique(unlist(str_locate_all(temp_producer, ','))))
comma_pos1 <- unique(unlist(str_locate_all(temp_producer, ','))) [1]
comma_pos2 <- unique(unlist(str_locate_all(temp_producer, ','))) [2]
comma_pos3 <- unique(unlist(str_locate_all(temp_producer, ','))) [3]
split1 <- substr(temp_producer, 1, comma_pos1 - 1)
split2 <- ifelse(num_commas > 1, 
                 substr(temp_producer, comma_pos1 + 1, comma_pos2 - 1),
                 substr(temp_producer, comma_pos1 + 1, nchar(temp_producer)))
split3 <- ifelse(num_commas > 2,
                 substr(temp_producer, comma_pos2 + 1, comma_pos3 - 1),
                 substr(temp_producer, comma_pos2 + 1, nchar(temp_producer)))
temp_producer1 <- c(temp_producer1, split1)
temp_producer2 <- c(temp_producer2, split2)
temp_producer3 <- c(temp_producer3, split3)
}

# join the producers into the temporary dataframe
temp_producer_dataframe <- data.frame(cbind(temp_producer_dataframe, temp_producer1, temp_producer2, temp_producer3))
temp_producer_dataframe <- temp_producer_dataframe[, c(1, 6, 7, 8)]

# join the new producers into the main dataframe
song_producer <- left_join(song_producer, temp_producer_dataframe, on = 'song_id')
song_producer$producer1 <- ifelse(is.na(song_producer$temp_producer1), song_producer$producer1, song_producer$temp_producer1)
song_producer$producer2 <- ifelse(is.na(song_producer$temp_producer2), song_producer$producer2, song_producer$temp_producer2)
song_producer$producer3 <- ifelse(is.na(song_producer$temp_producer3), song_producer$producer3, song_producer$temp_producer3)
song_producer <- song_producer[, - which(colnames(song_producer) %in% c('temp_producer1', 'temp_producer2', 'temp_producer3'))]

# clean up the environment
rm(temp_producer_dataframe, comma_pos1, comma_pos2, num_commas, split1, split2, split3,
temp_producer1, temp_producer2, temp_producer3, temp_producer, comma_pos3)

#-------------------------------------------------------------------------------#
# put all the pieces back together in a new song master

song_master2 <- song_master[, c(2, 1, 3, 11, 12)]

# clean up the environment
rm(date_cleaning, genre_dataframe, song_label_dataframe, song_label_matrix, song_producer_dataframe,
song_producer_matrix, song_songwriter_dataframe, song_songwriter_matrix, addelement, atomicelement,
check_genre, checkelement, count, dayyearnumbers, element, genre, genre_list, getelement, month,
monthcount, monthcounter, monthextract, new_unique_genre_list, newelement, newseparatedtext,
nextelement, parent_genre_list, rownumber, seperatedtext, seperatedtext, song_label_row, song_label1,
song_label2, song_label3, song_label4, song_label5, song_producer_row, song_producer1, song_producer2,
song_producer3, song_songwriter_row, song_songwriter1, song_songwriter2, song_songwriter3, song_songwriter_row,
song_targetcolumnposition, targetcolumnvalue, test, testing, textelement, unique_genre_list)

#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
#-------------------------------------------------------------------------------#
### T_song_genre ###

# change the album genre column names so that when it is joined
# there is not ambiguity with the song genre column names
colnames(album_genre) [3:15] <- paste0(colnames(album_genre) [3:15], '1')

# join the album names to to the song genre
song_genre <- unique(left_join(song_genre, data.frame(cbind('song_id' = paste(song_album[,1], song_album[,2]), 'album_id' =  paste(song_album[,2], song_album[,3]))), on = 'song_id'))

# join the album genre to the song genre
song_genre <- unique(left_join(song_genre, album_genre[, c(1, 3:15)], on = 'album_id'))

# converge the values from the album genre where needed
song_genre$alt <- ifelse( ! is.na(song_genre$alt1), song_genre$alt1, song_genre$alt)
song_genre$rock <- ifelse( ! is.na(song_genre$rock1), song_genre$rock1, song_genre$rock)
song_genre$metal <- ifelse( ! is.na(song_genre$metal1), song_genre$metal1, song_genre$metal)
song_genre$punk <- ifelse( ! is.na(song_genre$punk1), song_genre$punk1, song_genre$punk)
song_genre$pop <- ifelse( ! is.na(song_genre$pop1), song_genre$pop1, song_genre$pop)
song_genre$`r&b` <- ifelse( ! is.na(song_genre$`r&b1`), song_genre$`r&b1`, song_genre$`r&b`)
song_genre$rap <- ifelse( ! is.na(song_genre$rap1), song_genre$rap1, song_genre$rap)
song_genre$jazz <- ifelse( ! is.na(song_genre$jazz1), song_genre$jazz1, song_genre$jazz)
song_genre$blues <- ifelse( ! is.na(song_genre$blues1), song_genre$blues1, song_genre$blues)
song_genre$folk <- ifelse( ! is.na(song_genre$folk1), song_genre$folk1, song_genre$folk)
song_genre$country <- ifelse( ! is.na(song_genre$country1), song_genre$country1, song_genre$country)
song_genre$elect <- ifelse( ! is.na(song_genre$elect1), song_genre$elect1, song_genre$elect)
song_genre$other <- ifelse( ! is.na(song_genre$other1), song_genre$other1, song_genre$other)

# drop the columns that were used for lookup
T_song_genre <- unique(song_genre[, c(1, 3:15)])

#-------------------------------------------------------------------------------#
### T_song_label ###

# change the album genre column names so that when it is joined
# there is not ambiguity with the song genre column names
colnames(album_label) [3:5] <- paste0(colnames(album_label) [3:5], '1')

# join the album names to to the song genre
song_label <- unique(left_join(song_label, data.frame(cbind('song_id' = paste(song_album[,1], song_album[,2]), 'album_id' =  paste(song_album[,2], song_album[,3]))), on = 'song_id'))

# join the album genre to the song genre
song_label <- unique(left_join(song_label, album_label[, c(1, 3:5)], on = 'album_id'))

# converge the values from the album genre where needed
song_label$label1 <- ifelse( ! is.na(song_label$label11), song_label$label11, song_label$label1)
song_label$label2 <- ifelse( ! is.na(song_label$label21), song_label$label21, song_label$label2)
song_label$label3 <- ifelse( ! is.na(song_label$label31), song_label$label31, song_label$label3)

# drop the columns that were used for lookup
T_song_label <- unique(song_label[, c(1, 3:5)])

#-------------------------------------------------------------------------------#
### T_song_producer ###

# change the album genre column names so that when it is joined
# there is not ambiguity with the song genre column names
colnames(album_producer) [3:5] <- paste0(colnames(album_producer) [3:5], '1')

# join the album names to to the song genre
song_producer <- unique(left_join(song_producer, data.frame(cbind('song_id' = paste(song_album[,1], song_album[,2]), 'album_id' =  paste(song_album[,2], song_album[,3]))), on = 'song_id'))

# join the album genre to the song genre
song_producer <- unique(left_join(song_producer, album_producer[, c(1, 3:5)], on = 'album_id'))

# converge the values from the album genre where needed
song_producer$producer1 <- ifelse( ! is.na(song_producer$producer11), song_producer$producer11, song_producer$producer1)
song_producer$producer2 <- ifelse( ! is.na(song_producer$producer21), song_producer$producer21, song_producer$producer2)
song_producer$producer3 <- ifelse( ! is.na(song_producer$producer31), song_producer$producer31, song_producer$producer3)

# drop the columns that were used for lookup
T_song_producer <- unique(song_producer[, c(1, 3:5)])

#-------------------------------------------------------------------------------#
### T_song_master ###

# create the data frame from the song master
T_song_master <- song_master2

# join the album names to to the song genre
T_song_master <- unique(left_join(T_song_master, data.frame(cbind('song_id' = paste(song_album[,1], song_album[,2]), 'album_id' =  paste(song_album[,2], song_album[,3]))), on = 'song_id'))

# join the album genre to the song genre
T_song_master <- unique(left_join(T_song_master, album_master[, c(2, 7:8)], on = 'album_id'))

# change album release dates to numberic to align with song release dates
T_song_master$album_release_month <- as.numeric(T_song_master$album_release_month)
T_song_master$album_release_year <- as.numeric(T_song_master$album_release_year)

# converge the values from the album genre where needed
T_song_master$song_release_month <- ifelse( ! is.na(T_song_master$album_release_month), T_song_master$album_release_month, T_song_master$song_release_month)
T_song_master$song_release_year <- ifelse( ! is.na(T_song_master$album_release_year), T_song_master$album_release_year, T_song_master$song_release_year)

# drop the columns that were used for lookup
T_song_master <- unique(T_song_master[, c(1:2, 3:5)])

#-------------------------------------------------------------------------------#
### T_song_songwriter ###

# create the data frame from the song master
T_song_songwriter <- song_songwriter[, c(1, 3:5)]

#-------------------------------------------------------------------------------#
### filter down the results to minimize the number of nulls ###

# remove NA's
T_song_master <- T_song_master[which( ! is.na(T_song_master$song_release_year)),] 

```